\subsection{Spells (Q)}

\subsection{Spells (R)}

\DndSpellHeader{Ray of Enfeeblement\label{spell:ray-of-enfeeblement}}
{3 AET}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 1 minute}

A black beam of enervating energy springs from your finger toward a creature within range. Make a ranged spell attack against the target. On a hit, the target takes 3d8 necrotic damage and deals only half damage with weapon attacks until the spell ends.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage increases by 1d8 for every 2 additional aether spent.


\DndSpellHeader{Ray of Frost\label{spell:ray-of-frost}}
{cantrip}
{1 action}
{60 feet}
{V, S}
{Instantaneous}

A frigid beam of blue-white light streaks toward a creature within range. Make a ranged spell attack against the target. On a hit, it takes 1d8 cold damage, and its speed is reduced by 10 feet until the start of your next turn.

The spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Reincarnate\label{spell:reincarnate}}
{12 AET}
{1 hour}
{Touch}
{V, S, M (rare oils and unguents worth at least 1,000 gp, which the spell consumes)}
{Instantaneous}

You touch a dead humanoid or a piece of a dead humanoid. Provided that the creature has been dead no longer than 10 days, the spell forms a new adult body for it and then calls the soul to enter that body.

If the target's soul isn't free or willing to do so, the spell fails. The creature is aware of who is casting the spell and can refuse if it wishes.

The magic fashions a new body for the creature to inhabit, which likely causes the creature's race to change. The GM rolls a d100 and consults the following table to determine what form the creature takes when restored to life, or the GM chooses a form.

\begin{DndTable}{XX}
\textbf{d100}  & \textbf{Lineage}             \\
1-45 & \nameref{lineage:human} \\
46-55 & \nameref{lineage:dwarf} \\
56-65 & \nameref{lineage:ihmisi} \\
66-75 & \nameref{lineage:orc} \\
76-85 & \nameref{lineage:gwerin} \\
86-90 & \nameref{lineage:halfling} \\
91-95 & \nameref{lineage:dragonborn} \\
96-100 & \nameref{lineage:jazuu}
\end{DndTable}

The reincarnated creature recalls its former life and experiences. It retains the capabilities it had in its original form, except it exchanges its original lineage for the new one and changes its lineage traits accordingly. Its culture does not change.


\DndSpellHeader{Remove Curse\label{spell:remove-curse}}
{5 AET}
{1 action}
{Touch}
{V, S}
{Instantaneous}

At your touch, all non-legendary curses affecting one creature or object end. If the object is a cursed magic item, its curse remains, but the spell breaks its owner's attunement to the object so it can be removed or discarded.


\DndSpellHeader{Resilient Sphere\label{spell:resilient-sphere}}
{8 AET}
{1 action}
{30 feet}
{V, S, M (a hemispherical piece of clear crystal and a matching hemispherical piece of gum arabic)}
{Concentration, up to 1 minute}

A sphere of shimmering force encloses a creature or object of Large size or smaller within range. An unwilling creature must make a Dexterity saving throw. On a failed save, the creature is enclosed for the duration.

Nothing---not physical objects, energy, or other spell effects—--can pass through the barrier, in or out, though a creature in the sphere can breathe there. The sphere is immune to all damage from outside the sphere, and a creature or object inside can't be damaged by attacks or effects originating from outside, nor can a creature inside the sphere damage anything outside it. Unwilling targets inside can attack the sphere as an object with AC equal to your spell save DC, damage threshold 10, and 16 HP. When it reaches 0 HP, the spell ends and the globe vanishes.

The sphere is weightless and just large enough to contain the creature or object inside. An enclosed creature can use its action to push against the sphere's walls and thus roll the sphere at up to half the creature's speed. Similarly, the globe can be picked up and moved by other creatures. While enclosed, creatures take no damage from falling.


\DndSpellHeader{Resistance\label{spell:resistance}}
{Cantrip}
{1 action}
{Touch}
{V, S, M (a miniature cloak)}
{Concentration, up to 1 minute}

You touch one willing creature. Once before the spell ends, the target can roll a d4 and add the number rolled to one saving throw of its choice. It can roll the die before or after making the saving throw. The spell then ends. Remember that no more than one bonus die can be used on any d20 roll.

\DndSpellHeader{Retribution\label{spell:retribution}}
{2 AET}
{1 action}
{30 feet}
{S (a pointed finger)}
{Instantaneous}

One target you can see within range must make a Dexterity saving throw. On a failed save, the creature takes radiant damage equal to twice the difference between your maximum health and your current health (maximum 10). On a success, the creature takes half as much damage.

Alternatively, you can choose an ally within the spells' range to act as the source of the damage---the target takes damage (on a failed save) equal to twice the difference between the source's maximum health and their current health (up to the same limit).

In either case, temporary hit points do not count as health (maximum or otherwise).

\subparagraph*{Overcast} When you cast this spell and spend more than 2 AET, the maximum damage dealt increases by 10 for every 2 AET you spend.

\DndSpellHeader{Revivify\label{spell:revivify}}
{5 AET}
{1 action}
{Touch}
{V, S, M (diamonds worth 300 gp, which the spell consumes)}
{Instantaneous}

You touch a creature that has died within the last minute. That creature returns to life with 1 hit point. This spell can't return to life a creature that has died of old age, nor can it restore any missing body parts.

\subsection{Spells (S)}


\DndSpellHeader{Sacred Flame\label{spell:sacred-flame}}
{Cantrip}
{1 action}
{60 feet}
{V, S}
{Instantaneous}

Flame-like radiance descends on a creature that you can see within range. The target must succeed on a Dexterity saving throw or take 1d8 radiant damage. The target gains no benefit from cover for this saving throw and you do not need a clear path to the target.

The spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Sanctuary\label{spell:sanctuary}}
{2 AET}
{1 bonus action}
{30 feet}
{V, S, M (a small silver mirror)}
{1 minute}

You ward a creature within range against attack. Until the spell ends, any creature who targets the warded creature with an attack or a harmful spell must first make a Wisdom saving throw. On a failed save, the creature must choose a new target or lose the attack or spell. This spell doesn't protect the warded creature from area effects, such as the explosion of a fireball.

If the warded creature makes an attack, deals damage, or casts a spell that affects an enemy creature, this spell ends.


\DndSpellHeader{Scorching Ray\label{spell:scorching-ray}}
{3 AET}
{1 action}
{120 feet}
{V, S}
{Instantaneous}

You create three rays of fire and hurl them at targets within range. You can hurl them at one target or several.

Make a ranged spell attack for each ray. On a hit, the target takes 2d6 fire damage.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, you create one additional ray for every 3 additional AET you spend.


\DndSpellHeader{See Invisibility\label{spell:see-invisibility}}
{3 AET}
{1 action}
{Self}
{V, S, M (a pinch of talc and a small sprinkling of powdered silver)}
{1 hour}

For the duration, you see invisible creatures and objects as if they were visible, and you can see into the Border Shadow. Ethereal creatures and objects appear ghostly and translucent. You ignore the invisible condition.


\DndSpellHeader{Shatter\label{spell:shatter}}
{3 AET}
{1 action}
{60 feet}
{V, S, M (a chip of mica)}
{Instantaneous}

A sudden loud ringing noise, painfully intense, erupts from a point of your choice within range. Each creature in a 10-foot-radius sphere centered on that point must make a Constitution saving throw. A creature takes 3d8 thunder damage on a failed save, or half as much damage on a successful one. A creature made of inorganic material such as stone, crystal, or metal has disadvantage on this saving throw.

A non-magical object that isn't being worn or carried also takes the damage if it's in the spell's area. Objects made of crystal, glass, or stone have vulnerability to this damage.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Shield\label{spell:shield}}
{2 AET}
{1 action}
{Self}
{V, S}
{10 minutes}

An invisible barrier of magical force appears and protects you. This barrier has 10 hit points. Whenever you would take damage and the barrier is active, the barrier takes the damage instead, reducing its hit points. If the barrier absorbs all the damage, you do not need to make Concentration saves if you are concentrating on a spell or ability. You take any damage in excess of the barrier's hit points, which provokes Concentration saves as normal. When the barrier is reduced to zero hit points, the spell ends. Because this barrier takes damage instead of you, it does not benefit from any resistances or immunities you may have. Any overflow damage is, however, affected.

In addition, you are immune to the \nameref{spell:magic-missile} spell as long as the barrier is active.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the barrier's hit points increase by 5 for every additional AET spent.


\DndSpellHeader{Shield of Faith\label{spell:shield-of-faith}}
{2 AET}
{1 bonus action}
{60 feet}
{V, S, M (a small parchment with a bit of holy text written on it)}
{Concentration, up to 10 minutes}

A shimmering field appears and surrounds a creature of your choice within range, granting it the ability to use the \nameref{action:deflect} action without expending STA.


\DndSpellHeader{Shillelagh\label{spell:shillelagh}}
{Cantrip}
{1 bonus action}
{Touch}
{V, S, M (mistletoe, a shamrock leaf, and a club or quarterstaff)}
{1 minute}

The wood of a club or quarterstaff you are holding is imbued with nature's power. For the duration, you can use your spellcasting ability instead of Strength for the attack and damage rolls of melee attacks using that weapon, and the weapon's damage die becomes a d8. The weapon also becomes masterwork, if it isn't already. The spell ends if you cast it again or if you let go of the weapon.


\DndSpellHeader{Shocking Grasp\label{spell:shocking-grasp}}
{Cantrip}
{1 action}
{Touch}
{V, S}
{Instantaneous}

Lightning springs from your hand to deliver a shock to a creature you try to touch. Make a melee spell attack against the target. You have advantage on the attack roll if the target is wearing armor dominantly made of metal. On a hit, the target takes 1d8 lightning damage, and it can't take reactions until the start of its next turn.

The spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Silence\label{spell:silence}}
{3 AET}
{1 action}
{120 feet}
{V, S}
{Concentration, up to 10 minutes}

For the duration, no sound can be created within or pass through a 20-foot-radius sphere centered on a point you choose within range. Any creature or object entirely inside the sphere is immune to thunder damage, and creatures are deafened while entirely inside it. Casting a spell that includes a verbal component is impossible there.


\DndSpellHeader{Silent Image\label{spell:silent-image}}
{2 AET}
{1 action}
{60 feet}
{V, S, M (a bit of fleece)}
{Concentration, up to 10 minutes}

You create the image of an object, a creature, or some other visible phenomenon that is no larger than a 15-foot cube (ie no more than 15 feet in any dimension). The image appears at a spot within range and lasts for the duration. The image is purely visual; it isn't accompanied by sound, smell, or other sensory effects. It reflects light as normal for the object created.

You can use your action to cause the image to move to any spot within range. As the image changes location, you can alter its appearance so that its movements appear natural for the image. For example, if you create an image of a creature and move it, you can alter the image so that it appears to be walking.

Physical interaction with the image reveals it to be an illusion, because things can pass through it. A creature that uses its action to examine the image can determine that it is an illusion with a successful Intelligence (Investigation) check against your spell save DC. If a creature discerns the illusion for what it is, the creature can see through the image.


\DndSpellHeader{Sleep\label{spell:sleep}}
{2 AET}
{1 action}
{90 feet}
{V, S, M (a pinch of fine sand, rose petals, or a cricket)}
{Concentration, up to 1 minute}

This spell sends creatures into a magical slumber. Up to 5 creatures within 20 feet of a point you choose within range must make a Constitution saving throw. Creatures who fail their saving throw fall unconscious until the spell ends, the sleeper takes damage, or someone uses an action to shake or slap the sleeper awake.

Undead and creatures immune to being charmed aren't affected by this spell.

\subparagraph*{Overcast} When you cast this spell and expend more than 2 AET, the number of affected creatures increases by one for every 3 additional AET spent.


\DndSpellHeader{Sleet Storm\label{spell:sleet-storm}}
{5 AET}
{1 action}
{150 feet}
{V, S, M (a pinch of dust and a few drops of water)}
{Concentration, up to 1 minute}

Until the spell ends, freezing rain and sleet fall in a 20-foot-tall cylinder with a 40-foot radius centered on a point you choose within range. The area is heavily obscured, and exposed flames in the area are doused.

The ground in the area is covered with slick ice, making it difficult terrain. When a creature enters the spell's area for the first time on a turn or starts its turn there, it must make a Dexterity saving throw. On a failed save, it falls prone.

If a creature is concentrating in the spell's area, the creature must make a successful Constitution saving throw against your spell save DC or lose concentration.


\DndSpellHeader{Slow\label{spell:slow}}
{5 AET}
{1 action}
{120 feet}
{V, S, M (a drop of molasses)}
{Concentration, up to 1 minute}

You alter time around up to six creatures of your choice in a 40-foot cube within range. Each target must succeed on a Wisdom saving throw or be affected by this spell for the duration.

An affected target is \nameref{condition:shaken} and \nameref{condition:staggered}. On its turn, it can use either an action or a bonus action, not both. Regardless of the creature's abilities or magic items, it can't make more than one melee or ranged attack during its turn.

If the creature attempts to cast a spell with a casting time of 1 action, roll a d20. On an 11 or higher, the spell doesn't take effect until the creature's next turn, and the creature must use its action on that turn to complete the spell. If it can't, the spell is wasted.

A creature affected by this spell makes another Wisdom saving throw at the end of its turn. On a successful save, the effect ends for it.


\DndSpellHeader{Spike Growth\label{spell:spike-growth}}
{3 AET}
{1 action}
{150 feet}
{V, S, M (seven sharp thorns or seven small twigs, each sharpened to a point)}
{Concentration, up to 10 minutes}

The ground in a 20-foot radius centered on a point within range twists and sprouts hard spikes and thorns. The area becomes difficult terrain for the duration. When a creature moves into or within the area, it takes 1d4 piercing damage for every 5 feet it travels, up to a total per round of 16d4.

The transformation of the ground is camouflaged to look natural. Any creature that can't see the area at the time the spell is cast must make a Wisdom (Perception) check against your spell save DC to recognize the terrain as hazardous before entering it.


\DndSpellHeader{Spirit Guardians\label{spell:spirit-guardians}}
{8 AET}
{1 action}
{Self (15-foot radius)}
{V, S, M (a holy symbol)}
{Concentration, up to 10 minutes}

You call forth spirits to protect you. They flit around you to a distance of 15 feet for the duration. Their appearance is up to you.

When you cast this spell, you can designate any number of creatures you can see to be unaffected by it. Affected creatures treat the area as difficult terrain, and when the creature enters the area for the first time on a turn or starts its turn there, it must make a Dexterity saving throw. On a failed save, the creature takes 3d8 radiant damage or 3d8 necrotic damage (your choice when you cast the spell). On a successful save, the creature takes half as much damage.

\subparagraph*{Overcast} When you cast this spell using more than 5 AET, the damage increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Spiritual Weapon\label{spell:spiritual-weapon}}
{2 AET}
{1 bonus action}
{60 feet}
{V, S}
{1 minute}

You create a floating, spectral weapon within range that lasts for the duration or until you cast this spell again. When you cast the spell, you can make a melee spell attack against a creature within 5 feet of the weapon. On a hit, the target takes bludgeoning, piercing, or slashing (your choice) damage equal to 1d10 + your spellcasting ability modifier.

As a bonus action on your turn, you can move the weapon up to 20 feet and repeat the attack against a creature within 5 feet of it.

The weapon can take whatever form you choose. The damage type should be appropriate for the weapon.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d10 for every two additional AET.


\DndSpellHeader{Stinking Cloud\label{spell:stinking-cloud}}
{5 AET}
{1 action}
{90 feet}
{V, S, M (a rotten egg or several skunk cabbage leaves)}
{Concentration, up to 1 minute}

You create a 20-foot-radius sphere of yellow, nauseating gas centered on a point within range. The cloud spreads around corners, and its area is heavily obscured. The cloud lingers in the air for the duration.

Each creature that is completely within the cloud at the start of its turn must make a Constitution saving throw against poison. On a failed save, the creature spends its action that turn retching and reeling. Creatures that don't need to breathe or are immune to poison automatically succeed on this saving throw.

A moderate wind (at least 10 miles per hour) disperses the cloud after 4 rounds. A strong wind (at least 20 miles per hour) disperses it after 1 round.


\DndSpellHeader{Stone Shape\label{spell:stone-shape}}
{8 AET}
{1 action}
{Touch}
{V, S, M (soft clay, which must be worked into roughly the desired shape of the stone object)}
{Instantaneous}

You touch a stone object of Medium size or smaller or a section of stone no more than 5 feet in any dimension and form it into any shape that suits your purpose. So, for example, you could shape a large rock into a weapon, idol, or coffer, or make a small passage through a wall, as long as the wall is less than 5 feet thick. You could also shape a stone door or its frame to seal the door shut. The object you create can have up to two hinges and a latch, but finer mechanical detail isn't possible.


\DndSpellHeader{Stoneskin\label{spell:stoneskin}}
{8 AET}
{1 action}
{Touch}
{V, S, M (diamond dust worth 100 gp, which the spell consumes)}
{1 hour}

This spell turns the flesh of a willing creature you touch as hard as stone. Until the spell ends, the target has resistance to bludgeoning, piercing, and slashing damage, but their speed is reduced by 10 ft and they have disadvantage on Dexterity ability checks.


\DndSpellHeader{Suggestion\label{spell:suggestion}}
{3 AET}
{1 action}
{30 feet}
{V, M (a snake's tongue and either a bit of honeycomb or a drop of sweet oil)}
{Concentration, up to 8 hours}

You suggest a course of activity (limited to a sentence or two) and magically influence a creature you can see within range that can hear and understand you. Creatures that can't be charmed are immune to this effect. The suggestion must be worded in such a manner as to make the course of action sound reasonable, as judged by the GM. Asking the creature to stab itself, throw itself onto a spear, immolate itself, or do some other obviously harmful act ends the spell.

The target must make a Wisdom saving throw. On a failed save, it pursues the course of action you described to the best of its ability. The suggested course of action can continue for the entire duration. If the suggested activity can be completed in a shorter time, the spell ends when the subject finishes what it was asked to do.

You can also specify conditions that will trigger a specific course of action during the duration. For example, you could suggest that an orc stab his leader in the back the next time they're in camp together.

If you or any of your companions damage the target or any of its allies that it can see, the spell ends.

\DndSpellHeader{Sunbeam\label{spell:sunbeam}}
{14 AET}
{1 action}
{Self (60-foot line)}
{V, S, M (a magnifying glass)}
{Concentration, up to 1 minute}

A beam of brilliant sunlight flashes out from your hand in a 5-foot-wide, 60-foot-long line. Each creature in the line must make a Constitution saving throw. On a failed save, a creature takes 5d8 radiant damage and is blinded until your next turn. On a successful save, it takes half as much damage and isn't blinded by this spell. Undead and oozes have disadvantage on this saving throw.

You can create a new line of radiance as your action on any turn until the spell ends.

For the duration, a mote of brilliant radiance shines in your hand. It sheds bright light in a 30-foot radius and dim light for an additional 30 feet. This light is sunlight.

\subsection{Spells (T)}


\DndSpellHeader{Telekinesis\label{spell:telekinesis}}
{12 AET}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 10 minutes}

You gain the ability to move or manipulate creatures or objects by thought. When you cast the spell, and as your action each round for the duration, you can exert your will on one creature or object that you can see within range, causing the appropriate effect below. You can affect the same target round after round, or choose a new one at any time. If you switch targets, the prior target is no longer affected by the spell.

\subparagraph*{Creature.} You can try to move a Huge or smaller creature. Make an ability check with your spellcasting ability contested by the creature's Strength check. If you win the contest, you move the creature up to 30 feet in any direction, including upward but not beyond the range of this spell. Until the end of your next turn, the creature is restrained in your telekinetic grip. A creature lifted upward is suspended in mid-air.

On subsequent rounds, you can use your action to attempt to maintain your telekinetic grip on the creature by repeating the contest.

\subparagraph*{Object.} You can try to move an object that weighs up to 1,000 pounds. If the object isn't being worn or carried, you automatically move it up to 30 feet in any direction, but not beyond the range of this spell.

If the object is worn or carried by a creature, you must make an ability check with your spellcasting ability contested by that creature's Strength check. If you succeed, you pull the object away from that creature and can move it up to 30 feet in any direction but not beyond the range of this spell.

You can exert fine control on objects with your telekinetic grip, such as manipulating a simple tool, opening a door or a container, stowing or retrieving an item from an open container, or pouring the contents from a vial.


\DndSpellHeader{Thaumaturgy\label{spell:thaumaturgy}}
{Cantrip}
{1 action}
{30 feet}
{V}
{Up to 1 minute}

You manifest a minor wonder, a sign of supernatural power, within range. You create one of the following magical effects within range:
\begin{itemize}
\item Your voice booms up to three times as loud as normal for 1 minute.
\item You cause flames to flicker, brighten, dim, or change color for 1 minute.
\item You cause harmless tremors in the ground for 1 minute.
\item You create an instantaneous sound that originates from a point of your choice within range, such as a rumble of thunder, the cry of a raven, or ominous whispers.
\item You instantaneously cause an unlocked door or window to fly open or slam shut.
\item You alter the appearance of your eyes for 1 minute. If you cast this spell multiple times, you can have up to three of its 1-minute effects active at a time, and you can dismiss such an effect as an action.
\end{itemize}


\DndSpellHeader{Thunderwave\label{spell:thunderwave}}
{2 AET}
{1 action}
{Self (15-foot cube)}
{V, S}
{Instantaneous}

A wave of thunderous force sweeps out from your outstretched hands. Each creature in a 15-foot cube originating from one side of your space you must make a Strength saving throw. On a failed save, a creature takes 2d8 thunder damage and is pushed 10 feet away from you. On a successful save, the creature takes half as much damage and isn't pushed.

In addition, unsecured objects that are completely within the area of effect are automatically pushed 10 feet away from you by the spell's effect, and the spell emits a thunderous boom audible out to 300 feet.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d8 for every additional AET, up to a maximum of 8d8 (8 AET).


\DndSpellHeader{True Seeing\label{spell:true-seeing}}
{12 AET}
{1 action}
{Touch}
{V, S, M (an ointment for the eyes that costs 25 gp; is made from mushroom powder, saffron, and fat; and is consumed by the spell)}
{1 hour}

This spell gives the willing creature you touch the ability to see things as they actually are. For the duration, the creature has truesight, notices secret doors hidden by magic, and can see into the Border Shadow, all out to a range of 120 feet.


\DndSpellHeader{True Strike\label{spell:true-strike}}
{1 AET}
{1 bonus action}
{60 feet}
{S}
{1 round}

You extend your hand and point a finger at a target in range. Your magic grants you a brief insight into the target's defenses. Your next attack against the creature until the end of your next turn has advantage and scores a critical hit on a 19 or 20. 

\subsection{Spells (U)}

\DndSpellHeader{Unbind\label{spell:unbind}}
{5 AET}
{1 action}
{120 feet}
{V, S}
{Instantaneous}

Choose one creature, object, or magical effect within range. Any spell or ability costing 5 AET or lower on the target ends. For each spell higher cost on the target, make a proficient ability check using your spellcasting ability. The DC equals 10 + half the spell's aether cost. On a successful check, the spell ends.

This spell can affect non-spell magical effects as well. The effective aether cost is determined by the GM or specified in the ability that created it. Some creatures, such as animated objects, have special effects when this spell is cast on them. This has no effect on legendary effects or effects of similar power.

\subparagraph*{Overcast} When you cast this spell using more than 5 AET, you gain +1 on the check for every 3 additional AET spent.

\subsection{Spells (V)}

\DndSpellHeader{Vampiric Touch\label{spell:vampiric-touch}}
{4 AET}
{1 action}
{Self}
{V, S}
{Concentration, up to 1 minute}

The touch of your shadow-wreathed hand can siphon life force from others to heal your wounds. Make a melee spell attack against a creature within your reach. On a hit, the target takes 3d6 necrotic damage, and you regain hit points equal to half the amount of necrotic damage dealt. Until the spell ends, you can make the attack again on each of your turns as an action.

\subparagraph*{Overcast} When you cast this spell using more than 4 AET, the damage increases by 1d6 for every two additional AET spent.

\DndSpellHeader{Vicious Mockery\label{spell:vicious-mockery}}
{0 AET}
{1 action}
{60 ft}
{V}
{Instantaneous}

You unleash a string of insults laced with subtle enchantments at a creature you can see within range. If the target can hear you (though it need not understand you), it must succeed on a Wisdom saving throw or take 1d4 psychic damage and have disadvantage on the next attack roll it makes before the end of its next turn.

This spell's damage increases by 1d4 when you reach 5th level (2d4), 11th level (3d4), and 17th level (4d4).

\subsection{Spells (W)}

\DndSpellHeader{Wall of Fire\label{spell:wall-of-fire}}
{8 AET}
{1 action}
{120 feet}
{V, S, M (a small piece of phosphorus)}
{Concentration, up to 1 minute}

You create a wall of fire on a solid surface within range. You can make the wall up to 60 feet long, 20 feet high, and 1 foot thick, but it must form in a straight line. The wall is opaque and lasts for the duration.

When the wall appears, each creature within its area must make a Dexterity saving throw. On a failed save, a creature takes 5d8 fire damage, or half as much damage on a successful save.

One side of the wall, selected by you when you cast this spell, deals 5d8 fire damage to each creature that ends its turn within 10 feet of that side or inside the wall. A creature takes the same damage when it enters the wall for the first time on a turn or ends its turn there. The other side of the wall deals no damage.

\subparagraph*{Overcast} When you cast this spell using more than 8 AET, the damage increases by 1d8 for every 3 additional AET.


\DndSpellHeader{Wall of Force\label{spell:wall-of-force}}
{15 AET}
{1 action}
{120 feet}
{V, S, M (a pinch of powder made by crushing a clear gemstone)}
{Concentration, up to 10 minutes}

An invisible wall of force springs into existence at a point you choose within range. The wall appears in any orientation you choose, as a horizontal or vertical barrier or at an angle. It can be free floating or resting on a solid surface. You can shape a flat surface made up of ten 10-foot-by-10-foot panels. Each panel must be contiguous with another panel and cannot have any gaps within a panel, but you can choose the orientation of each panel separately. In any form, the wall is 1/4 inch thick. It lasts for the duration. If the wall cuts through a creature's space when it appears, the creature is pushed to one side of the wall (your choice which side).

Nothing can physically pass through the wall---it provides total cover but does not obscure the target. Non-damaging sound and light can pass through wall as normal. The wall also extends into the Border Shadow, blocking ethereal travel through the wall. Each panel of the wall has an AC equal to your spell-save DC, a damage threshold of 10, and 45 HP. When its HP is reduced to 0, that panel collapses.

\subparagraph*{Overcast} When you cast this spell using more than 15 AET, its HP increases by 3 for every additional AET spent.


\DndSpellHeader{Wall of Ice\label{spell:wall-of-ice}}
{12 AET}
{1 action}
{120 feet}
{V, S, M (a small piece of quartz)}
{Concentration, up to 10 minutes}

You create a wall of ice on a solid surface within range. You can form it into a hemispherical dome or a sphere with a radius of up to 10 feet, or you can shape a vertical surface made up of ten 10-foot-square panels. Each panel must be contiguous with another panel along its edges. In any form, the wall is 1 foot thick and lasts for the duration.

If the wall cuts through a creature's space when it appears, the creature within its area is pushed to one side of the wall and must make a Dexterity saving throw. On a failed save, the creature takes 10d6 cold damage, or half as much damage on a successful save.

The wall is an object that can be damaged and thus breached. It has AC equal to 4 + your spell attack modifier and 30 hit points per 10-foot section, and it is vulnerable to fire damage. Reducing a 10-foot section of wall to 0 hit points destroys it and leaves behind a sheet of frigid air in the space the wall occupied. A creature moving through the sheet of frigid air for the first time on a turn must make a Constitution saving throw. That creature takes 5d6 cold damage on a failed save, or half as much damage on a successful one.

\subparagraph*{Overcast} When you cast this spell using a more than 12 AET, the damage the wall deals when it appears increases by 2d6, and the damage from passing through the sheet of frigid air increases by 1d6, for every 3 additional AET.


\DndSpellHeader{Wall of Stone\label{spell:wall-of-stone}}
{12 AET}
{1 action}
{120 feet}
{V, S, M (a small block of granite)}
{Concentration, up to 10 minutes}

A non-magical wall of solid stone springs into existence at a point you choose within range. The wall is 6 inches thick and is composed of ten 10-foot by-10-foot panels. Each panel must be contiguous with at least one other panel. Alternatively, you can create 10-foot-by-20-foot panels that are only 3 inches thick.

If the wall cuts through a creature's space when it appears, the creature is pushed to one side of the wall (your choice). If a creature would be surrounded on all sides by the wall (or the wall and another solid surface), that creature can make a Dexterity saving throw. On a success, it can use its reaction to move up to its speed so that it is no longer enclosed by the wall.

The wall can have any shape you desire, though it can't occupy the same space as a creature or object. The wall doesn't need to be vertical or rest on any firm foundation. It must, however, merge with and be solidly supported by existing stone along one edge of one panel. Thus, you can use this spell to bridge a chasm or create a ramp.

If you create a span greater than 20 feet in length, you must halve the size of each panel to create supports. You can crudely shape the wall to create crenellations, battlements, and so on, but cannot have any gaps or openings within individual panels.

The wall is an object made of stone that can be damaged and thus breached. Each panel has AC 6 + your spell attack modifier and 30 hit points per inch of thickness. Reducing a panel to 0 hit points destroys it and might cause connected panels to collapse at the GM's discretion.

The wall disappears when the spell ends.


\DndSpellHeader{Wall of Thorns\label{spell:wall-of-thorns}}
{7 AET}
{1 action}
{120 feet}
{V, S, M (a handful of thorns)}
{Concentration, up to 10 minutes}

You create a wall of tough, pliable, tangled brush bristling with needle-sharp thorns. The wall appears within range on a solid surface and lasts for the duration. You choose to make the wall up to 60 feet long, 10 feet high, and 5 feet thick or a circle that has a 20-foot diameter and is up to 20 feet high and 5 feet thick. The wall blocks line of sight and provides 3/4 cover.

When the wall appears, each creature within its area must make a Dexterity saving throw. On a failed save, a creature takes 5d8 piercing damage, or half as much damage on a successful save.

A creature can move through the wall, albeit slowly and painfully. For every 1 foot a creature moves through the wall, it must spend 4 feet of movement. Furthermore, the first time a creature enters the wall on a turn or ends its turn there, the creature must make a Dexterity saving throw. It takes 5d8 slashing damage on a failed save, or half as much damage on a successful one.

\subparagraph*{Overcast} When you cast this spell using more than 9 AET, both types of damage increase by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Warding Bond\label{spell:warding-bond}}
{3 AET}
{1 action}
{Touch}
{V, S, M (a pair of platinum rings worth at least 50 gp each, which you and the target must wear for the duration)}
{1 hour}

This spell wards a willing creature you touch and creates a mystic connection between you and the target until the spell ends. While the target is within 60 feet of you, it has resistance to all damage. Also, each time it takes damage, you take the same amount of damage. The resistance to damage on the target is factored in first, then any resistance or immunity you have applies separately to the reduced damage.

The spell ends if you drop to 0 hit points or if you and the target become separated by more than 60 feet. It also ends if the spell is cast again on either of the connected creatures. You can also dismiss the spell as an action.


\DndSpellHeader{Web\label{spell:web}}
{3 AET}
{1 action}
{60 feet}
{V, S, M (a bit of spiderweb)}
{Concentration, up to 1 hour}

You conjure a mass of thick, sticky webbing at a point of your choice within range. The webs fill a 20 foot cube from that point for the duration unless conjured in mid-air without support, in which case it tangles around anything in the area and falls to the ground. The webs are difficult terrain and lightly obscure their area.

If the webs aren't anchored between two solid masses (such as walls or trees) or layered across a floor, wall, or ceiling, the conjured web collapses on itself, and the spell ends at the start of your next turn. Webs layered over a flat surface have a depth of 5 feet.

Each creature that starts its turn in the webs or that enters them during its turn must make a Strength saving throw. On a failed save, the creature is restrained as long as it remains in the webs or until it breaks free.

A creature restrained by the webs can use its action to make a Strength check against your spell save DC. If it succeeds, it is no longer restrained.

The webs are flammable. Any 5-foot cube of webs exposed to fire burns away in 1 round, dealing 2d4 fire damage to any creature that starts its turn in the fire.


\DndSpellHeader{Wind Wall\label{spell:wind-wall}}
{5 AET}
{1 action}
{120 feet}
{V, S, M (a tiny fan and a feather of exotic origin)}
{Concentration, up to 1 minute}

A wall of strong wind rises from the ground at a point you choose within range. You can make the wall up to 50 feet long, 15 feet high, and 1 foot thick. You can shape the wall in any way you choose so long as it makes one continuous path along the ground. The wall lasts for the duration.

When the wall appears, each creature within its area must make a Strength saving throw. A creature takes 3d8 bludgeoning damage on a failed save, or half as much damage on a successful one.

The strong wind keeps fog, smoke, and other gases at bay. Small or smaller flying creatures or objects can't pass through the wall. Loose, lightweight materials brought into the wall fly upward. Arrows, bolts, and other ordinary projectiles launched at targets behind the wall are deflected upward and automatically miss. (Boulders hurled by giants or siege engines, and similar projectiles, are unaffected.) Creatures in gaseous form can't pass through it.

\subsection{Spells (X)}

\subsection{Spells (Y)}
\DndSpellHeader{Yoink\label{spell:yoink}}
{1 AET}
{1 bonus action}
{90 ft}
{V, S, M (a tiny replica grappling hook and thread)}
{Instantaneous}

You extend a tether of sticky force and try to pull a creature or unattended object within range. Unwilling creatures must make a Strength saving throw; creatures more than one size larger than you do so at advantage. Willing creatures, objects, and those that fail their saving throw are pulled up to 15 feet closer to you. You can angle this as long as their final position is closer to you than their initial position. If you target an object, it must be no more than one size larger than you, otherwise the spell fails.

Alternatively, you can use this spell to grapple to a fixed object that can support your weight or a creature you can see of at least two sizes larger than yourself within range. If you do so, you are pulled 15 feet closer to that creature or object.

\subsection{Spells (Z)}