\subsection{Spells (H)}

\DndSpellHeader{Hallucinatory Terrain\label{spell:hallucinatory-terrain}}
{8 AET}
{10 minutes}
{300 feet}
{V, S, M (a stone, a twig, and a bit of green plant)}
{24 hours}

You make natural terrain in a 150-foot cube in range look, sound, and smell like some other sort of natural terrain. Thus, open fields or a road can be made to resemble a swamp, hill, crevasse, or some other difficult or impassable terrain. A pond can be made to seem like a grassy meadow, a precipice like a gentle slope, or a rock-strewn gully like a wide and smooth road. Manufactured structures, equipment, and creatures within the area aren't changed in appearance.

The tactile characteristics of the terrain are unchanged, so creatures entering the area are likely to see through the illusion. If the difference isn't obvious by touch, a creature carefully examining the illusion can attempt an Intelligence (Investigation) check against your spell save DC to disbelieve it. A creature who discerns the illusion for what it is, sees it as a vague image superimposed on the terrain.


\DndSpellHeader{Haste\label{spell:haste}}
{5 AET}
{1 action}
{30 feet}
{V, S, M (a shaving of licorice root)}
{Concentration, up to 1 minute}

Choose a willing creature that you can see within range. Until the spell ends, the target's speed is doubled, it gains a +2 bonus to AC, it has advantage on Dexterity saving throws, and it gains an additional action on each of its turns. That action can be used only to take the Dash, Disengage, Hide, or Use an Object action.

\subparagraph*{Overcast} When you cast this spell using 7 AET, the target can also take the Attack action (making only a single weapon attack). Alternatively, you can affect one additional target with the base effect for every 3 additional AET spent.


\DndSpellHeader{Healing Word\label{spell:healing-word}}
{2 AET}
{1 action}
{60 feet}
{V}
{Instantaneous}

A creature of your choice that you can see within range regains hit points equal to 1d4 + your spellcasting ability modifier. This spell has no effect on undead or constructs.
\subparagraph*{Overcast} When you cast this spell using more than 1 AET, the healing increases by 1d4 for every 2 additional AET spent.


\DndSpellHeader{Heat Metal\label{spell:heat-metal}}
{3 AET}
{1 action}
{60 feet}
{V, S, M (a piece of iron and a flame)}
{Concentration, up to 1 minute}

Choose a manufactured metal object, such as a metal weapon or a suit of heavy or medium metal armor, that you can see within range. You cause the object to glow red-hot. Any creature in physical contact with the object takes 2d8 fire damage when you cast the spell. Until the spell ends, you can use a bonus action on each of your subsequent turns to cause this damage again.

If a creature is holding or wearing the object and takes the damage from it, the creature must succeed on a Constitution saving throw or drop the object if it can. If it doesn't drop the object, it has disadvantage on attack rolls and ability checks until the start of your next turn.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Heroism\label{spell:heroism}}
{2 AET}
{1 action}
{Touch}
{V, S}
{Concentration, up to 1 minute}

A willing creature you touch is imbued with bravery. Until the spell ends, the creature is immune to being frightened and gains temporary hit points equal to your spellcasting ability modifier at the start of each of its turns. When the spell ends, the target loses any remaining temporary hit points from this spell.
\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can target one additional creature for every 3 additional AET spent.


\DndSpellHeader{Hideous Laughter\label{spell:hideous-laughter}}
{2 AET}
{1 action}
{30 feet}
{V, S, M (tiny tarts and a feather that is waved in the air)}
{Concentration, up to 1 minute}

A creature of your choice that you can see within range perceives everything as hilariously funny and falls into fits of laughter if this spell affects it. The target must succeed on a Wisdom saving throw or fall prone and be unable to stand up for the duration. Affected creatures cannot cast spells with verbal components and must make Constitution saving throw against your spell save DC at the start of their turn to maintain concentration on spells or abilities. A creature with an Intelligence score of -3 or less isn't affected.

At the end of each of its turns, and each time it takes damage, the target can make another Wisdom saving throw. The target has advantage on the saving throw if it's triggered by damage. On a success, the spell ends.

\DndSpellHeader{Hypnotic Pattern\label{spell:hypnotic-pattern}}
{5 AET}
{1 action}
{120 feet}
{S, M (a glowing stick of incense or a crystal vial filled with phosphorescent material)}
{Concentration, up to 1 minute}

You create a twisting pattern of colors that weaves through the air inside a 30-foot cube within range. The pattern appears for a moment and vanishes. Each creature in the area who sees the pattern must make a Charisma saving throw. On a failed save, the creature becomes charmed for the duration. While charmed by this spell, the creature has a speed of 0 and must make a Charisma saving throw against your spell save DC if they try to act, losing their action on a failure. Affected creatures can attempt the saving throw at the end of each of their turns, ending the effect on a success. The spell also ends for an affected creature if it takes any damage.

\subsection{Spells (I)}

\DndSpellHeader{Ice Storm\label{spell:ice-storm}}
{7 AET}
{1 action}
{300 feet}
{V, S, M (a pinch of dust and a few drops of water)}
{Instantaneous}

A hail of rock-hard ice pounds to the ground in a 20-foot-radius, 40-foot-high cylinder centered on a point within range. Each creature in the cylinder must make a Dexterity saving throw. A creature takes 4d6 bludgeoning damage and 4d6 cold damage on a failed save, or half as much damage on a successful one.

Hailstones turn the storm's area of effect into difficult terrain until the end of your next turn.
\subparagraph*{Overcast} When you cast this spell using more than 7 AET, the cold and bludgeoning damage each increase by 1d6 for every 3 additional AET spent.


\DndSpellHeader{Inflict Wounds\label{spell:inflict-wounds}}
{2 AET}
{1 action}
{Touch}
{V, S}
{Instantaneous}

Make a melee spell attack against a creature you can reach. On a hit, the target takes 3d10 necrotic damage.
\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d10 for every 2 additional AET spent.


\DndSpellHeader{Insect Plague\label{spell:insect-plague}}
{14 AET}
{1 action}
{300 feet}
{V, S, M (a few grains of sugar, some kernels of grain, and a smear of fat)}
{Concentration, up to 10 minutes}

Swarming, biting locusts fill a 20-foot-radius sphere centered on a point you choose within range. The sphere spreads around corners. The sphere remains for the duration, and its area is lightly obscured. The sphere's area is difficult terrain.

When the area appears, each creature in it must make a Constitution saving throw. A creature takes 4d10 piercing damage on a failed save, or half as much damage on a successful one. A creature must also make this saving throw when it enters the spell's area for the first time on a turn or ends its turn there.
\subparagraph*{Overcast} When you cast this spell using more than 14 AET, the damage increases by 1d10 for every 2 additional AET spent.


\DndSpellHeader{Invisibility\label{spell:invisibility}}
{3 AET}
{1 action}
{Touch}
{V, S, M (an eyelash encased in gum arabic)}
{Concentration, up to 1 hour}

A creature you touch becomes invisible until the spell ends. Anything the target is wearing or carrying when the spell is cast is invisible as long as it is on the target's person. The spell ends for a target that attacks, casts a spell, or deals damage to a creature or object. Objects picked up by an invisible target do not become invisible.
\subparagraph*{Overcast} When you cast this spell using more than 3 AET, you can target one additional creature for every 3 additional AET spent.

\subsection{Spells (J)}


\DndSpellHeader{Jump\label{spell:jump}}
{2 AET}
{1 action}
{Touch}
{V, S, M (a grasshopper's hind leg)}
{1 minute}

You touch a creature. The creature's jump distance is tripled until the spell ends.

\subsection{Spells (K)}

\subsection{Spells (L)}


\DndSpellHeader{Levitate\label{spell:levitate}}
{3 AET}
{1 action}
{60 feet}
{V, S, M (either a small leather loop or a piece of golden wire bent into a cup shape with a long shank on one end)}
{Concentration, up to 10 minutes}

One creature or object of your choice that you can see within range rises vertically, up to 20 feet, and remains suspended there for the duration. The spell can levitate a target that weighs up to 500 pounds. An unwilling creature that succeeds on a Strength saving throw is unaffected.

The target can move only by pushing or pulling against a fixed object or surface within reach (such as a wall or a ceiling), which allows it to move as if it were climbing. You can change the target's altitude by up to 20 feet in either direction on your turn. If you are the target, you can move up or down as part of your movement. Otherwise, you can use your action to move the target, which must remain within the spell's range.

When the spell ends, the target floats gently to the ground if it is still aloft.


\DndSpellHeader{Light\label{spell:light}}
{cantrip}
{1 action}
{Touch}
{V, M (a firefly or phosphorescent moss)}
{1 hour}

You touch one object that is no larger than 10 feet in any dimension. Until the spell ends, the object sheds bright light in a 20-foot radius and dim light for an additional 20 feet. The light can be colored as you like. Completely covering the object with something opaque blocks the light. The spell ends if you cast it again or dismiss it as an action.

If you target an object held or worn by a hostile creature, that creature must succeed on a Dexterity saving throw to avoid the spell.


\DndSpellHeader{Lightning Bolt\label{spell:lightning-bolt}}
{5 AET}
{1 action}
{Self (100-foot line)}
{V, S, M (a bit of fur and a rod of amber, crystal, or glass)}
{Instantaneous}

A stroke of lightning forming a line 100 feet long and 5 feet wide blasts out from you in a direction you choose. Each creature in the line must make a Dexterity saving throw. A creature takes 6d6 lightning damage on a failed save, or half as much damage on a successful one. Targets that fail their saving throw by 5 or more are shaken until the end of their next turn.

The lightning ignites flammable objects in the area that aren't being worn or carried.
\subparagraph*{Overcast.}When you cast this spell using more than 5 AET, the damage increases by 1d6 for every 2 additional aether spent.


\DndSpellHeader{Longstrider\label{spell:longstrider}}
{2 AET}
{1 action}
{Touch}
{V, S, M (a pinch of dirt)}
{1 hour}

You touch a creature. The target's speed increases by 10 feet until the spell ends.
\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can target one additional creature for every 3 additional AET spent.

\subsection{Spells (M)}


\DndSpellHeader{Mage Armor\label{spell:mage-armor}}
{2 AET}
{1 action}
{Touch}
{V, S, M (a piece of cured leather)}
{8 hours}

You touch a willing creature who isn't wearing armor, and a protective magical force surrounds it until the spell ends. The target's base AC becomes 12 + its Dexterity modifier and the target can take the \nameref{action:deflect} action. The spell ends if the target dons armor or if you dismiss the spell as an action.


\DndSpellHeader{Mage Hand\label{spell:mage-hand}}
{cantrip}
{1 action}
{30 feet}
{V, S}
{1 minute}

A spectral, floating hand appears at a point you choose within range. The hand lasts for the duration or until you dismiss it as an action. The hand vanishes if it is ever more than 30 feet away from you or if you cast this spell again.

You can use your action to control the hand. You can use the hand to manipulate an object, open an unlocked door or container, stow or retrieve an item from an open container, or pour the contents out of a vial. You can move the hand up to 30 feet each time you use it.

The hand can't attack, activate magic items, or carry more than 10 pounds.


\DndSpellHeader{Magic Missile\label{spell:magic-missile}}
{2 AET}
{1 action}
{120 feet}
{V, S}
{Instantaneous}

You create three glowing darts of magical force. Each dart hits a creature of your choice that you can see within range. A dart deals 1d4 + 1 piercing damage to its target. The darts all strike simultaneously, and you can direct them to hit one creature or several. Roll for each dart separately; each dart causes a different instance of damage for effects (such as concentration saving throw) that care about such things.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the spell creates one more dart for every 2 additional AET spent.


\DndSpellHeader{Magic Weapon\label{spell:magic-weapon}}
{3 AET}
{1 bonus action}
{Touch}
{V, S}
{Concentration, up to 1 hour}

You touch a non-masterwork weapon. Until the spell ends, that weapon becomes masterwork for the duration, granting a bonus to damage equal to the wielder's proficiency bonus.
\subparagraph*{Overcast} When you cast this spell using more than 3 AET, you can target one additional weapon for every 3 AET you spend.

\DndSpellHeader{Major Image\label{spell:major-image}}
{5 AET}
{1 action}
{120 feet}
{V, S, M (a bit of fleece)}
{Concentration, up to 10 minutes}

You create the image of an object, a creature, or some other visible phenomenon that is no larger than a 20-foot cube. The image appears at a spot that you can see within range and lasts for the duration. It seems completely real, including sounds, smells, and temperature appropriate to the thing depicted. You can't create sufficient heat or cold to cause damage, a sound loud enough to deal thunder damage or deafen a creature, or a smell that might sicken a creature (like a troglodyte's stench), or any effect that would impose any negative condition on a creature. An image of a reflective surface (such as a mirror) reflects only what the viewer would expect it to reflect, but cannot be used to focus light such as to ignite objects.

As long as you are within range of the illusion, you can use your action to cause the image to move to any other spot within range. As the image changes location, you can alter its appearance so that its movements appear natural for the image. For example, if you create an image of a creature and move it, you can alter the image so that it appears to be walking. Similarly, you can cause the illusion to make different sounds at different times, even making it carry on a conversation, for example.

Physical interaction with the image reveals it to be an illusion, because things can pass through it. A creature that uses its action to examine the image can determine that it is an illusion with a successful Intelligence (Investigation) check against your spell save DC. If a creature discerns the illusion for what it is, the creature can see through the image, and its other sensory qualities become faint to the creature.
\subparagraph*{Overcast} When you cast this spell using 15 AET, the spell lasts until dispelled, without requiring your concentration.


\DndSpellHeader{Mass Cure Wounds\label{spell:mass-cure-wounds}}
{9 AET}
{1 action}
{60 feet}
{V, S}
{Instantaneous}

A wave of healing energy washes out from a point of your choice within range. Choose up to six creatures in a 30-foot-radius sphere centered on that point. Each target regains hit points equal to 3d8 + your spellcasting ability modifier. This spell has no effect on undead or constructs.
\subparagraph*{Overcast} When you cast this spell using more than 9 AET, the healing increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Meld into Stone\label{spell:meld-into-stone}}
{5 AET}
{1 action}
{Touch}
{V, S}
{8 hours}

You step into a stone object or surface large enough to fully contain your body, melding yourself and all the equipment you carry with the stone for the duration. Using your movement, you step into the stone at a point you can touch. Nothing of your presence remains visible or otherwise detectable by non-magical senses.

While merged with the stone, you can't see what occurs outside it, and any Wisdom (Perception) checks you make to hear sounds outside it are made with disadvantage. You remain aware of the passage of time and can cast spells on yourself while merged in the stone. You can use your movement to leave the stone where you entered it, which ends the spell. You otherwise can't move.

Minor physical damage to the stone doesn't harm you, but its partial destruction or a change in its shape (to the extent that you no longer fit within it) expels you and deals 6d6 bludgeoning damage to you. The stone's complete destruction (or transmutation into a different substance) expels you and deals 50 bludgeoning damage to you. If expelled, you fall prone in an unoccupied space closest to where you first entered.


\DndSpellHeader{Message\label{spell:message}}
{cantrip}
{1 action}
{120 feet}
{S, M (a short piece of copper wire)}
{1 round}

You point your finger toward a creature within range and whisper a message. The target (and only the target) hears the message and can reply in a whisper that only you can hear.

You can cast this spell through solid objects if you are familiar with the target and know it is beyond the barrier. Magical silence, 1 foot of stone, 1 inch of common metal, a thin sheet of lead, or 3 feet of wood blocks the spell. The spell doesn't have to follow a straight line and can travel freely around corners or through openings.


\DndSpellHeader{Minor Illusion\label{spell:minor-illusion}}
{cantrip}
{1 action}
{30 feet}
{S, M (a bit of fleece)}
{1 minute}

You create a sound or an image of an object within range that lasts for the duration. The illusion also ends if you dismiss it as an action or cast this spell again.

If you create a sound, its volume can range from a whisper to a scream. It can be your voice, someone else's voice, a lion's roar, a beating of drums, or any other sound you choose. The sound continues unabated throughout the duration, or you can make discrete sounds at different times before the spell ends.

If you create an image of an object---such as a chair, muddy footprints, or a small chest---it must be no larger than a 5-foot cube (ie less than 5ft on any dimension). The image can't create sound, light, smell, or any other sensory effect. Creating a reflective surface (such as a hand mirror) works but only reflects what the viewer expects to see. Physical interaction with the image reveals it to be an illusion, because things can pass through it. 

If a creature uses its action to examine the sound or image, the creature can determine that it is an illusion with a successful Intelligence (Investigation) check against your spell save DC. If a creature discerns the illusion for what it is, the illusion becomes faint to the creature.


\DndSpellHeader{Mirror Image\label{spell:mirror-image}}
{3 AET}
{1 action}
{Self}
{V, S}
{1 minute}

Three illusory duplicates of yourself appear in your space. Until the spell ends, the duplicates move with you and mimic your actions, shifting position so it's impossible to track which image is real. You can use your action to dismiss the illusory duplicates.

Each time a creature targets you with an attack during the spell's duration, you can choose to expend a duplicate as a reaction. If you do so, the duplicate vanishes but the attack deals no damage to you. You can make this decision after you see the attack roll but before damage is assigned.

A creature is unaffected by this spell if it can't see, if it relies on senses other than sight, such as blindsight, or if it can perceive illusions as false, as with truesight.


\DndSpellHeader{Mislead\label{spell:mislead}}
{12 AET}
{1 action}
{Self}
{S}
{Concentration, up to 1 hour}

You become invisible at the same time that an illusory double of you appears where you are standing. The double lasts for the duration, but the invisibility ends if you attack or cast a spell.

You can use your action to move your illusory double up to twice your speed and make it gesture, speak, and behave in whatever way you choose.

You can see through its eyes and hear through its ears as if you were located where it is. On each of your turns as a bonus action, you can switch from using its senses to using your own, or back again. While you are using its senses, you are blinded and deafened in regard to your own surroundings.


\DndSpellHeader{Misty Step\label{spell:misty-step}}
{3 AET}
{1 bonus action}
{Self}
{V, S}
{Instantaneous}

Briefly surrounded by silvery mist, you teleport up to 30 feet to an unoccupied space that you can see. There must be a clear path between you and the target location (ie you can't teleport through transparent total cover such as a window).


\DndSpellHeader{Moonbeam\label{spell:moonbeam}}
{4 AET}
{1 action}
{120 feet}
{V, S, M (several seeds of any moonseed plant and a piece of opalescent feldspar)}
{Concentration, up to 1 minute}

A silvery beam of pale light shines down in a 5-foot-radius, 40-foot-high cylinder centered on a point within range. Until the spell ends, dim light fills the cylinder.

When a creature enters the spell's area for the first time on a turn or starts its turn there, it is engulfed in ghostly flames that cause searing pain, and it must make a Constitution saving throw. It takes 2d10 radiant damage on a failed save, or half as much damage on a successful one.

A creature whose physical shape is not their native one (such as a shapeshifted dragon or disguised succubus) makes its saving throw with disadvantage. If it fails, it also instantly reverts to its original form and can't assume a different form until it leaves the spell's light.

On each of your turns after you cast this spell, you can use an action to move the beam 60 feet in any direction.
\subparagraph*{Overcast} When you cast this spell using more than 4 AET, the damage increases by 1d10 for every 2 additional AET spent.

\DndSpellHeader{Move Earth\label{spell:move-earth}}
{12 AET}
{1 action}
{120 feet}
{V, S, M (an iron blade and a small bag containing a mixture of soils—clay, loam, and sand)}
{Concentration, up to 2 hours}

Choose an area of terrain no larger than 40 feet on a side within range. You can reshape dirt, sand, or clay in the area in any manner you choose for the duration. You can raise or lower the area's elevation, create or fill in a trench, erect or flatten a wall, or form a pillar. The extent of any such changes can't exceed half the area's largest dimension. So, if you affect a 40-foot square, you can create a pillar up to 20 feet high, raise or lower the square's elevation by up to 20 feet, dig a trench up to 20 feet deep, and so on. It takes 10 minutes for these changes to complete.

At the end of every 10 minutes you spend concentrating on the effect, you can choose a new area of terrain to affect.

Because the terrain's transformation occurs slowly, creatures in the area can't usually be trapped or injured by the ground's movement.

This effect can't manipulate natural stone or stone construction. Rocks and structures shift to accommodate the new terrain. If the way you shape the terrain would make a structure unstable, it might collapse.

Similarly, this effect doesn't directly affect plant growth. The moved earth carries any plants along with it.

\subsection{Spells (N)}

\DndSpellHeader{Noxious Stench \label{spell:noxious-stench}}
{3 AET}
{1 action}
{Self (10 ft radius)}
{S, M (a drop of skunk musk or something similar)}
{1 minute}

A burst of noxious fumes billows from you. All creatures other than yourself that have a sense of smell in the area must make a Constitution saving throw. On a failed saving throw, they take 4d8 poison damage and are \nameref{condition:staggered} until the end of their next turn. On a success, they take half as much damage and are not staggered. Creatures with an enhanced sense of smell make the saving throw at disadvantage and are stunned instead of staggered on a failure.

The scent lingers for 10 minutes or until dispersed by a brisk wind, but does not cause saving throws after the initial instantaneous effect. Creatures that have passed through the cloud retain some of the stench for 8 hours or until they thoroughly clean themselves and their gear. This makes most other creatures dislike them, but also foils any attempt to track them via non-magical scent for that period.

\subsection{Spells (O)}

\subsection{Spells (P)}


\DndSpellHeader{Pass without Trace\label{spell:pass-without-trace}}
{4 AET}
{1 action}
{Self}
{V, S, M (ashes from a burned leaf of mistletoe and a sprig of spruce)}
{Concentration, up to 1 hour}

A veil of shadows and silence radiates from you, masking you and your companions from detection. For the duration, each creature you choose within 30 feet of you (including you) treats a d20 result of less than 10 as a 10 on Dexterity (Stealth) checks and any attempt to track them via sight or sound is at disadvantage. Scent is unaffected.


\DndSpellHeader{Passwall\label{spell:passwall}}
{12 AET}
{1 action}
{30 feet}
{V, S, M (a pinch of sesame seeds)}
{1 hour}

A passage appears at a point of your choice that you can see on a wooden, plaster, or stone surface (such as a wall, a ceiling, or a floor) within range, and lasts for the duration. You choose the opening's dimensions: up to 5 feet wide, 8 feet tall, and 20 feet deep. The passage creates no instability in a structure surrounding it.

When the opening disappears, any creatures or objects still in the passage created by the spell are safely ejected to an unoccupied space nearest to the surface on which you cast the spell.


\DndSpellHeader{Phantasmal Killer\label{spell:phantasmal-killer}}
{8 AET}
{1 action}
{120 feet}
{V, S}
{Concentration, up to 1 minute}

You tap into the nightmares of a creature you can see within range and create an illusory manifestation of its deepest fears, visible only to that creature. The target must make an Intelligence saving throw. On a failed save, the target takes 4d10 psychic damage and becomes shaken for the duration. At the end of each of the target's turns before the spell ends, the target must succeed on an Intelligence saving throw, taking 4d10 psychic damage and becoming frightened as well if they were shaken before. On a successful save, the spell ends.

\subparagraph*{Overcast} When you cast this spell using more than 8 AET, the damage increases by 1d10 for every 2 additional AET spent.


\DndSpellHeader{Plant Growth\label{spell:plant-growth}}
{5 AET}
{1 action}
{150 feet}
{V, S}
{Instantaneous}

This spell channels vitality into plants within a specific area, causing them to grow rapidly

If you cast this spell using 1 action, choose a point within range. All normal plants in a 100-foot radius centered on that point become thick and overgrown. A creature moving through the area must spend 4 feet of movement for every 1 foot it moves. This does not stack with difficult terrain.

You can exclude one or more areas of any size within the spell's area from being affected.

\DndSpellHeader{Poison Spray\label{spell:poison-spray}}
{0 AET}
{1 action}
{Self (15 ft cone)}
{V, S}
{Instantaneous}

You spray poison in a cone. All creatures in that cone must make a Constitution saving throw, taking 1d6 poison damage on a failure. The damage from this spell increases to 2d6 at 5th level, 3d6 at 11th level, and 4d6 at 17th level.

\DndSpellHeader{Polymorph\label{spell:polymorph}}
{8 AET}
{1 action}
{60 feet}
{V, S, M (a caterpillar cocoon)}
{Concentration, up to 1 hour}

This spell transforms a creature that you can see within range into a new form. An unwilling creature must make a Charisma saving throw to avoid the effect. The spell has no effect on a shapechanger or a creature with 0 hit points.

The transformation lasts for the duration, or until the target drops to 0 hit points or dies. The new form can be any beast whose offensive rating is equal to or less than half the aether spent on the spell. The target's game statistics, including mental ability scores, are replaced by the statistics of the chosen beast. It retains its alignment and personality and can recognize friends and foes, but does not understand the nature of the change (including its ending conditions).

The target assumes the hit points of its new form. When it reverts to its normal form, the creature returns to the number of hit points it had before it transformed. If it reverts as a result of dropping to 0 hit points, any excess damage carries over to its normal form. As long as the excess damage doesn't reduce the creature's normal form to 0 hit points, it isn't knocked unconscious.

The creature is limited in the actions it can perform by the nature of its new form, and it can't speak, cast spells, or take any other action that requires hands or speech.

The target's gear melds into the new form. The creature can't activate, use, wield, or otherwise benefit from any of its equipment.


\DndSpellHeader{Prayer of Healing\label{spell:prayer-of-healing}}
{3 AET}
{10 minutes}
{30 feet}
{V}
{Instantaneous}

Up to six creatures of your choice that you can see within range each regain hit points equal to 2d8 + your spellcasting ability modifier. This spell has no effect on undead or constructs.
\subparagraph*{Overcast} When you cast this spell and spend more than 3 AET, the healing increases by 1d8 for every AET above 3, up to a maximum of 13d8 + your spellcasting ability modifier.


\DndSpellHeader{Prestidigitation\label{spell:prestidigitation}}
{cantrip}
{1 action}
{10 feet}
{V, S}
{Up to 1 hour}

This spell is a minor magical trick that novice spell-casters use for practice. You create one of the following magical effects within range:

\begin{itemize}
\item You create an instantaneous, harmless sensory effect, such as a shower of sparks, a puff of wind, faint musical notes, or an odd odor.
\item You instantaneously light or snuff out a candle, a torch, or a small campfire.
\item You instantaneously clean or soil an object no larger than 1 cubic foot.
\item You chill, warm, or flavor up to 1 cubic foot of nonliving material for 1 hour.
\item You make a color, a small mark, or a symbol appear on an object or a surface for 1 hour.
\item You create a non-magical trinket or an illusory image that can fit in your hand and that lasts until the end of your next turn.
\end{itemize}

If you cast this spell multiple times, you can have up to three of its non-instantaneous effects active at a time, and you can dismiss such an effect as an action. None of its effects can impose any conditions or deal any damage to a creature or object (although lighting a fire can subsequently damage objects because it is real fire).


\DndSpellHeader{Produce Flame\label{spell:produce-flame}}
{cantrip}
{1 action}
{Self}
{V, S}
{10 minutes}

A flickering flame appears in your hand. The flame remains there for the duration and harms neither you nor your equipment. The flame sheds bright light in a 10-foot radius and dim light for an additional 10 feet. The spell ends if you dismiss it as an action or if you cast it again.

You can also attack with the flame, although doing so ends the spell. When you cast this spell, or as an action on a later turn, you can hurl the flame at a creature within 90 feet of you. Make a ranged spell attack. On a hit, the target takes 1d8 fire damage.

This spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Protection from Energy\label{spell:protection-from-energy}}
{5 AET}
{1 action}
{Touch}
{V, S}
{Concentration, up to 1 hour}

For the duration, the willing creature you touch has resistance to one damage type of your choice: acid, cold, fire, lightning, or thunder.


\DndSpellHeader{Protection from Otherworldly Influence\label{spell:protection-from-otherworldly-influence}}
{2 AET}
{1 action}
{Touch}
{V, S, M (holy water or powdered silver and iron, which the spell consumes)}
{Concentration up to 10 minutes}

Until the spell ends, one willing creature you touch is protected against certain types of creatures: aberrations, celestials, elementals, fey, fiends, and undead.

The protection grants several benefits. Creatures of those types have disadvantage on attack rolls against the target. The target also can't be charmed, frightened, or possessed by them. If the target is already charmed, frightened, or possessed by such a creature, the target has advantage on any new saving throw against the relevant effect.


\DndSpellHeader{Protection from Poison\label{spell:protection-from-poison}}
{3 AET}
{1 action}
{Touch}
{V, S}
{1 hour}

You touch a creature. If it is poisoned, you neutralize the poison. If more than one poison afflicts the target, you neutralize one poison that you know is present, or you neutralize one at random.

For the duration, the target has advantage on saving throws against being poisoned, and it has resistance to poison damage.