\section{Weapons, Magic}\label{mi:weapons}

\subsection{Ammunition of Slaying}
\textit{Weapon (arrow, bolt, or sling stone), Hero} 

An arrow of slaying is a magic weapon meant to slay a particular kind of creature. Some are more focused than others; for example, there are both arrows of dragon slaying and arrows of blue dragon slaying. If a creature belonging to the type, race, or group associated with an arrow of slaying is hit by the arrow, the attack is a critical hit and the creature takes an additional 6d10 piercing damage (not doubled due to the critical hit). In addition, the creature must make a DC 17 Constitution saving throw. On a failed save, the creature is killed outright. If the creature was reduced to zero hit points by the attack's damage (including the extra damage), it automatically fails the saving throw. Once an arrow of slaying deals its extra damage to a creature, it becomes a nonmagical arrow. Other types of magic ammunition of this kind exist, such as bolts of slaying meant for a crossbow, though arrows are most common.

\subsection{Berserker Axe}
\textit{Weapon (any masterwork axe), Hero (requires attunement)}

This is a masterwork, adamantine axe. In addition, while you are attuned to this weapon, your hit point maximum increases by 1 for each level you have attained.

\subsection{Blade of Sharpness}
\textit{Weapon (any weapon that deals slashing damage), Hero (requires attunement)}

When you attack an object with this sword and hit, maximize your weapon damage dice against the target. When you attack a creature with this weapon and roll a 20 on the attack roll, that target takes an extra 4d6 slashing damage. Then roll another d20. If you roll a 20, you lop off one of the target's limbs, with the effect of such loss determined by the GM. If the creature has no limb to sever, you lop off a portion of its body instead.

In addition, you can speak the sword's command word to cause the blade to shed bright light in a 10-foot radius and dim light for an additional 10 feet. Speaking the command word again or sheathing the sword puts out the light.

\subsection{Dagger of Venom}
\textit{Weapon (masterwork dagger), Hero}

This is a masterwork dagger.

You can use an action to cause thick, black poison to coat the blade. The poison remains for 1 minute or until an attack using this weapon hits a creature. That creature must succeed on a DC 15 Constitution saving throw or take 2d10 poison damage and become poisoned for 1 minute. The dagger can't be used this way again until the next dawn.

\subsection{Dancing Sword}
\textit{Weapon (any sword), Hero (requires attunement)}

You can use a bonus action to toss this magic sword into the air and speak the command word. When you do so, the sword begins to hover, flies up to 30 feet, and attacks one creature of your choice within 5 feet of it. The sword uses your attack roll and ability score modifier to damage rolls.

While the sword hovers, you can use a bonus action to cause it to fly up to 30 feet to another spot within 30 feet of you. As part of the same bonus action, you can cause the sword to attack one creature within 5 feet of it.

After the hovering sword attacks for the fourth time, it flies up to 30 feet and tries to return to your hand. If you have no hand free, it falls to the ground at your feet. If the sword has no unobstructed path to you, it moves as close to you as it can and then falls to the ground. It also ceases to hover if you grasp it or move more than 30 feet away from it.

\subsection{Defender}
\textit{Weapon (any masterwork, silvered sword), legendary (requires attunement)}

This is a masterwork, silvered sword. While attuned to it, you can take the Deflect action without expending stamina and the effects last until the start of your next turn instead of for one attack only.

\subsection{Dragon Slayer}
\textit{Weapon (any masterwork sword, axe, spear, lance, halberd or glaive), Hero} 

This is a masterwork weapon. When you hit a dragon with this weapon, the dragon takes an extra 3d6 damage of the weapon's type. For the purpose of this weapon, “dragon” refers to any creature with the dragon type, including dragon turtles and wyverns.

\subsection{Dwarven Thrower}
\textit{Weapon (warhammer), Hero (requires attunement)}

This is a masterwork, adamantine warhammer. It has the thrown property with a normal range of 20 feet and a long range of 60 feet. When you hit with a ranged attack using this weapon, it deals an extra 1d8 damage or, if the target is a giant, 2d8 damage. Immediately after the attack, the weapon flies back to your hand.

\subsection{Energy Bow}
\textit{Weapon (any bow or crossbow), Hero (requires attunement)}

This masterwork weapon has no natural string. Instead, a thread of magical force is generated between the tips of the bowstave. Similarly, it requires no ammunition, because it generates its own that count as silver and adamantine. Energy crossbows ignore the Loading property.

\subsection{Flame Tongue}
\textit{Weapon (any sword), Hero (requires attunement)} 

You can use a bonus action to speak this sword's command word, causing flames to erupt from the blade. These flames shed bright light in a 40-foot radius and dim light for an additional 40 feet. While the sword is ablaze, it deals an extra 2d6 fire damage to any target it hits. The flames last until you use a bonus action to speak the command word again or until you drop or sheathe the sword.

\subsection{Frost Brand}
\textit{Weapon (any sword), Hero (requires attunement)} 

When you hit with an attack using this sword, the target takes an extra 1d6 cold damage. In addition, while you hold the sword, you have resistance to fire damage.

In freezing temperatures, the blade sheds bright light in a 10-foot radius and dim light for an additional 10 feet. When you draw this weapon, you can extinguish all nonmagical flames within 30 feet of you. This property can be used no more than once per hour.

\subsection{Giant Slayer}
\textit{Weapon (any axe or sword), Hero} 

This is a masterwork weapon. When you hit a giant with it, the giant takes an extra 2d6 damage of the weapon's type and must succeed on a DC 15 Strength saving throw or fall prone. For the purpose of this weapon, “giant” refers to any creature with the giant type, including ettins and trolls.

\subsection{Hammer of Thunderbolts}
\textit{Weapon (maul), legendary}

This is a masterwork weapon that counts as adamantine and silvered.

\subparagraph*{Giant's Bane (Requires Attunement)} You must be wearing a belt of giant strength (any variety) and gauntlets of ogre power to attune to this weapon. The attunement ends if you take off either of those items. While you are attuned to this weapon and holding it, your Strength score increases by 2 and can exceed +5, but not +10. When you roll a 20 on an attack roll made with this weapon against a giant, the giant must succeed on a DC 17 Constitution saving throw or die.

The hammer also has 5 charges. While attuned to it, you can expend 1 charge and make a ranged weapon attack with the hammer, hurling it as if it had the thrown property with a normal range of 20 feet and a long range of 60 feet. If the attack hits, the hammer unleashes a thunderclap audible out to 300 feet. The target and every creature within 30 feet of it must succeed on a DC 17 Constitution saving throw or be stunned until the end of your next turn. The hammer regains 1d4 + 1 expended charges daily at dawn.

\subsection{Holy Avenger}
\textit{Weapon (any sword), legendary (requires attunement by an oathbound)}

This is a masterwork weapon that counts as silvered. When you hit a fiend or an undead with it, that creature takes an extra 2d10 radiant damage.

While you hold the drawn sword, it creates an aura in a 10-foot radius around you. You and all creatures friendly to you in the aura have advantage on saving throws against spells and other magical effects. If you have 17 or more levels in the Oathbound class, the radius of the aura increases to 30 feet.

\subsection{Javelin of Lightning}
\textit{Weapon (javelin), Adventurer}

When you hurl this javelin and speak its command word, it transforms into a bolt of lightning, forming a line 5 feet wide that extends out from you to a target within 120 feet. Each creature in the line excluding you and the target must make a DC 13 Dexterity saving throw, taking 4d6 lightning damage on a failed save, and half as much damage on a successful one. The lightning bolt turns back into a javelin when it reaches the target. Make a ranged weapon attack against the target. On a hit, the target takes damage from the javelin plus 4d6 lightning damage.

The javelin's property can't be used again until the next dawn. In the meantime, the javelin can still be used as a magic weapon.

\subsection{Luck Blade}
\textit{Weapon (any sword), Adventurer (requires attunement)}

This is a masterwork weapon. This sword has 3 charges. If the sword is on your person, you can call on its luck and expend a charge (no action required) to reroll one attack roll, ability check, or saving throw you dislike. You must use the second roll. The sword regains 1d3 charges at dawn.

\subsection{Mace of Disruption}
\textit{Weapon (any bludgeoning), Hero (requires attunement)} 

When you hit a fiend or an undead with this magic weapon, that creature takes an extra 2d6 radiant damage. If the target has 25 hit points or fewer after taking this damage, it must succeed on a DC 15 Wisdom saving throw or be destroyed. On a successful save, the creature becomes frightened of you until the end of your next turn.  While you hold this weapon, it sheds bright light in a 20-foot radius and dim light for an additional 20 feet.

\subsection{Mace of Smiting}
\textit{Weapon (any bludgeoning), Hero}

This is a masterwork weapon that counts as adamantine. When used against a construct, you have advantage on the attack.  When you roll a 20 on an attack roll made with this weapon, the target takes an extra 2d6 bludgeoning damage, or 4d6 bludgeoning damage if it's a construct. If a construct has 25 hit points or fewer after taking this damage, it is destroyed.

\subsection{Mace of Terror}
\textit{Weapon (any bludgeoning), Hero (requires attunement)} 

This magic weapon has 3 charges. While holding it, you can use an action and expend 1 charge to release a wave of terror. Each creature of your choice in a 30-foot radius extending from you must succeed on a DC 15 Wisdom saving throw or become frightened of you for 1 minute. While it is frightened in this way, a creature must spend its turns trying to move as far away from you as it can, and it can't willingly move to a space within 30 feet of you. It also can't take reactions. For its action, it can use only the Dash action or try to escape from an effect that prevents it from moving. If it has nowhere it can move, the creature can use the Guard action. At the end of each of its turns, a creature can repeat the saving throw, ending the effect on itself on a success. The mace regains 1d3 expended charges daily at dawn.

\subsection{Moons-glow Weapon}
\textit{Weapon (any sword, axe, or polearm), Journeyman}

This magic weapon was forged and blessed by kami. It counts as being silvered, and when unsheathed in the dark, it glows, shedding bright light for 10 feet and dim light for 10 feet beyond that. The color of the glow changes as the moons wax and wane, ranging from white when Quella is new to pure red when Quella is full.

\subparagraph*{Formula} To craft a moons-glow weapon, you must provide one lb of refined silver metal. At least one day during the crafting period must be spent in a place sacred to a kami, and you must perform rituals appropriate to that kami while crafting. The required proficiency is Smith's tools (+2, +3 if masterwork). Crafting a moons-glow weapon costs 75 gold (150 if you wish to make a masterwork weapon) above the base price of the weapon.

\subsection{Nine Lives Stealer}
\textit{Weapon (any melee), Hero (requires attunement)}

This is a masterwork weapon. The sword has 1d8 + 1 charges. If you score a critical hit against a creature that has fewer than 100 hit points, it must succeed on a DC 15 Constitution saving throw or be slain instantly as the sword tears its life force from its body (a construct or an undead is immune). The sword loses 1 charge if the creature is slain. When the sword has no charges remaining, it loses this property.

\subsection{Oathbow}
\textit{Weapon (longbow), Hero (requires attunement)}

When you nock an arrow on this bow, it whispers in Metsae, “Swift defeat to my enemies.” When you use this weapon to make a ranged attack, you can, as a command phrase, say, “Swift death to you who have wronged me.” The target of your attack becomes your sworn enemy until it dies or until dawn seven days later. You can have only one such sworn enemy at a time. When your sworn enemy dies, you can choose a new one after the next dawn.

When you make a ranged attack roll with this weapon against your sworn enemy, you have advantage on the roll. In addition, your target gains no benefit from cover, other than total cover, and you suffer no disadvantage due to long range. If the attack hits, your sworn enemy takes an extra 3d6 piercing damage. While your sworn enemy lives, you have disadvantage on attack rolls with all other weapons.

\subsection{Scimitar of Speed}
\textit{Weapon (scimitar), Hero (requires attunement)}

This is a masterwork weapon. In addition, you can make one attack with it as a bonus action on each of your turns.

\subsection{Staff of Charming}
\textit{Staff, Hero (requires attunement)} 

While holding this staff, you can use an action to expend 1 of its 10 charges to cast charm person, command, or comprehend languages from it using your spell save DC or 8 + your Charisma modifier + your proficiency bonus if you do not have a spellcasting DC. The staff can also be used as a magic quarterstaff. If you are holding the staff and fail a saving throw against an enchantment spell that targets only you, you can turn your failed save into a successful one. You can't use this property of the staff again until the next dawn. If you succeed on a save against an enchantment spell that targets only you, with or without the staff's intervention, you can use your reaction to expend 1 charge from the staff and turn the spell back on its caster as if you had cast the spell.  The staff regains 1d8 + 2 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, the staff becomes a non-magical quarterstaff.

\subsection{Staff of Fire}
\textit{Staff, Hero (requires attunement)}

You have resistance to fire damage while you hold this staff.

The staff has 10 charges. While holding it, you can use an action to expend 1 or more of its charges to cast one of the following spells from it, using your spell save DC or 8 + your Intelligence modifier + your proficiency bonus if you do not have a spellcasting DC: burning hands (1 charge), fireball (3 charges), or wall of fire (4 charges).

The staff regains 1d6 + 4 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, the staff blackens, crumbles into cinders, and is destroyed.

\subsection{Staff of Frost}
\textit{Staff, Hero (requires attunement)}

You have resistance to cold damage while you hold this staff.

The staff has 10 charges. While holding it, you can use an action to expend 1 or more of its charges to cast one of the following spells from it, using your spell save DC or 8 + your Intelligence modifier + your proficiency bonus if you do not have a spellcasting DC: cone of cold (5 charges), fog cloud (1 charge), ice storm (4 charges), or wall of ice (4 charges).

The staff regains 1d6 + 4 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, the staff turns to water and is destroyed.

\subsection{Staff of Healing}
\textit{Staff, Hero (requires attunement)}

This staff has 10 charges. While holding it, you can use an action to expend 1 or more of its charges to cast one of the following spells from it, using your spell save DC and spellcasting ability modifier or 8 + your Wisdom modifier + your proficiency bonus if you do not have a spellcasting DC (using Wisdom as your spellcasting ability modifier in that case): cure wounds (1 charge per two aether, up to 8 AET), lesser restoration (2 charges), or mass cure wounds (5 charges).

The staff regains 1d6 + 4 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, the staff vanishes in a flash of light, lost forever.

\subsection{Staff of Power}
\textit{Staff, Hero (requires attunement)}

This staff can be wielded as a magic quarterstaff that grants a +4 bonus to damage rolls made with it.

The staff has 20 charges for the following properties. The staff regains 2d8 + 4 expended charges daily at dawn as long as it has at least one charge remaining. If you expend the last charge, roll a d20. On a 20, the staff regains 1d8 + 2 charges.

\subparagraph*{Power Strike} When you hit with a melee attack using the staff, you can expend 1 charge to deal an extra 1d6 bludgeoning damage to the target.

\subparagraph*{Spells}. While holding this staff, you can use an action to expend 1 or more of its charges to cast one of the following spells from it, using your spell save DC and spell attack bonus: cone of cold (5 charges), fireball (cast with 12 AET, 5 charges), globe of invulnerability (6 charges), hold monster (5 charges), levitate (2 charges), lightning bolt (cast with 12 AET, 5 charges), magic missile (1 charge), ray of enfeeblement (1 charge), or wall of force (6 charges).

\subparagraph*{Retributive Strike} You can use an action to break the staff over your knee or against a solid surface, performing a retributive strike. The staff is destroyed and releases its remaining magic in an explosion that expands to fill a 30-foot-radius sphere centered on it.  You have a 50 percent chance to instantly travel to a random plane of existence, avoiding the explosion. If you fail to avoid the effect, you take bludgeoning damage equal to 16 \texttimes the number of charges in the staff. Every other creature in the area must make a DC 17 Dexterity saving throw. On a failed save, a creature takes an amount of damage based on how far away it is from the point of origin, as shown in the following table. On a successful save, a creature takes half as much damage.
\begin{DndTable}{lX}
Distance from Origin &  Damage \\                                   
10 ft. away or closer & 8 \texttimes the number of charges in the staff   \\
11 to 20 ft. away     & 6 \texttimes the  number of charges in the staff  \\
21 to 30 ft. away     & 4 \texttimes the number of charges in the staff   \\
\end{DndTable}

\subsection{Staff of Striking}
\textit{Staff, Hero (requires attunement)}

This staff can be wielded as a masterwork quarterstaff.  The staff has 10 charges. When you hit with a melee attack using it, you can expend up to 3 of its charges. For each charge you expend, the target takes an extra 1d6 bludgeoning damage. The staff regains 1d6 + 4 expended charges daily at dawn as long as it has at least one charge remaining.

\subsection{Staff of Swarming Insects}
\textit{Staff, Hero (requires attunement)}

This staff has 10 charges and regains 1d6 + 4 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, a swarm of insects consumes and destroys the staff, then disperses.

\subparagraph*{Spells} While holding the staff, you can use an action to expend some of its charges to cast one of the following spells from it, using your spell save DC or 8 + your Wisdom modifier + your proficiency bonus if you do not have a spell save DC: giant insect (4 charges) or insect plague (5 charges).

\subparagraph*{Insect Cloud} While holding the staff, you can use an action and expend 1 charge to cause a swarm of harmless flying insects to spread out in a 30-foot radius from you. The insects remain for 10 minutes, making the area heavily obscured for creatures other than you. The swarm moves with you, remaining centered on you. A wind of at least 10 miles per hour disperses the swarm and ends the effect.

\subsection{Staff of the Python}
\textit{Staff, Adventurer (requires attunement)}

You can use an action to speak this staff's command word and throw the staff on the ground within 10 feet of you. The staff becomes a giant constrictor snake under your control and acts on its own initiative count. By using a bonus action to speak the command word again, you return the staff to its normal form in a space formerly occupied by the snake.  On your turn, you can mentally command the snake if it is within 60 feet of you and you aren't incapacitated. You decide what action the snake takes and where it moves during its next turn, or you can issue it a general command, such as to attack your enemies or guard a location.  If the snake is reduced to 0 hit points, it dies and reverts to its staff form. The staff then shatters and is destroyed. If the snake reverts to staff form before losing all its hit points, it regains all of them.

\subsection{Staff of the Woodlands}
\textit{Staff, Hero (requires attunement)}

This staff can be wielded as a masterwork quarterstaff.

The staff has 10 charges for the following properties. It regains 1d6 + 4 expended charges daily at dawn. If you expend the last charge, roll a d20. On a 1, the staff loses its properties and becomes a nonmagical quarterstaff.

\subparagraph*{Spells} You can use an action to expend 1 or more of the staff's charges to cast one of the following spells from it, using your spell save DC: animal friendship (1 charge), awaken(5 charges), barkskin (2 charges), or wall of thorns(6 charges).

In addition, you can use the Voice the Voiceless incantation (Animals and Plants only).

\subparagraph*{Tree Form} You can use an action to plant one end of the staff in fertile earth and expend 1 charge to transform the staff into a healthy tree. The tree is 60 feet tall and has a 5-foot-diameter trunk, and its branches at the top spread out in a 20-foot radius. The tree appears ordinary but radiates a faint aura of transmutation magic if targeted by detect magic. While touching the tree and using another action to speak its command word, you return the staff to its normal form. Any creature in the tree falls when it reverts to a staff.

\subsection{Staff of Thunder and Lightning}
\textit{Staff, Hero (requires attunement)}

This staff can be wielded as a masterwork quarterstaff. It also has the following additional properties. When one of these properties is used, it can't be used again until the next dawn.

\subparagraph*{Lightning} When you hit with a melee attack using the staff, you can cause the target to take an extra 2d6 lightning damage.

\subparagraph*{Thunder} When you hit with a melee attack using the staff, you can cause the staff to emit a crack of thunder, audible out to 300 feet. The target you hit must succeed on a DC 17 Constitution saving throw or become stunned until the end of your next turn.

\subparagraph*{Lightning Strike} You can use an action to cause a bolt of lightning to leap from the staff's tip in a line that is 5 feet wide and 120 feet long. Each creature in that line must make a DC 17 Dexterity saving throw, taking 9d6 lightning damage on a failed save, or half as much damage on a successful one.

\subparagraph*{Thunderclap} You can use an action to cause the staff to issue a deafening thunderclap, audible out to 600 feet. Each creature within 60 feet of you (not including you) must make a DC 17 Constitution saving throw. On a failed save, a creature takes 2d6 thunder damage and becomes deafened for 1 minute. On a successful save, a creature takes half damage and isn't deafened.

\subparagraph*{Thunder and Lightning} You can use an action to use the Lightning Strike and Thunderclap properties at the same time. Doing so doesn't expend the daily use of those properties, only the use of this one.

\subsection{Staff of Withering}
\textit{Staff, Hero (requires attunement)} 

This staff has 3 charges and regains 1d3 expended charges daily at dawn. The staff can be wielded as a quarterstaff. On a hit, it deals damage as a normal quarterstaff, and you can expend 1 charge to deal an extra 2d10 necrotic damage to the target. In addition, the target must succeed on a DC 15 Constitution saving throw or have disadvantage for 1 hour on any ability check or saving throw that uses Strength or Constitution.

\subsection{Sun Blade}
\textit{Weapon (longsword), Hero (requires attunement)}

This item appears to be a longsword hilt. While grasping the hilt, you can use a bonus action to cause a blade of pure radiance to spring into existence, or make the blade disappear. While the blade exists, this magic longsword has the finesse property. If you are proficient with shortswords or longswords, you are proficient with the sun blade.

This is a masterwork weapon which counts as silvered and which deals radiant damage instead of slashing damage. When you hit an undead with it, that target takes an extra 1d8 radiant damage.

The sword's luminous blade emits bright light in a 15-foot radius and dim light for an additional 15 feet. The light is sunlight. While the blade persists, you can use an action to expand or reduce its radius of bright and dim light by 5 feet each, to a maximum of 30 feet each or a minimum of 10 feet each.

\subsection{Sword of Life Stealing}
\textit{Weapon (any sword), Hero (requires attunement)}

When you attack a creature with this magic weapon and roll a 20 on the attack roll, that target takes an extra 3d6 necrotic damage, provided that the target isn't a construct or an undead. You gain temporary hit points equal to the extra damage dealt.

\subsection{Trident of Fish Command}
\textit{Weapon (trident), Adventurer (requires attunement)}

This trident is a magic weapon. It has 3 charges. While you carry it, you can use an action and expend 1 charge to cast
dominate beast(save DC 15) from it on a beast that has an innate swimming speed. The trident regains 1d3 expended charges daily at dawn.

\subsection{Vicious Weapon}
\textit{Weapon (any), Hero}

When you roll a 20 on your attack roll with this magic weapon, your critical hit deals an extra 2d6 damage of the weapon's type.

\subsection{Vorpal Sword}
\textit{Weapon (any sword that deals slashing damage), legendary (requires attunement)}

This is a masterwork weapon that counts as silvered and adamantine. In addition, the weapon ignores resistance to slashing damage.  When you attack a creature that has at least one head with this weapon and roll a 20 on the attack roll, you cut off one of the creature's heads. The creature dies if it can't survive without the lost head. A creature is immune to this effect if it is immune to slashing damage, doesn't have or need a head, has legendary actions, or the GM decides that the creature is too big for its head to be cut off with this weapon. Such a creature instead takes an extra 6d8 slashing damage from the hit.

\subsection{Weapon of Wounding}
\textit{Weapon (any), Hero (requires attunement)}

Hit points lost to this weapon's damage can be regained only through a short or long rest, rather than by regeneration, magic, or any other means. Once per turn, when you hit a creature with an attack using this magic weapon, you can wound the target. At the start of each of the wounded creature's turns, it takes 1d4 necrotic damage for each time you've wounded it, and it can then make a DC 15 Constitution saving throw, ending the effect of all such wounds on itself on a success. Alternatively, the wounded creature, or a creature within 5 feet of it, can use an action to make a DC 15 Wisdom (Medicine) check, ending the effect of such wounds on it on a success.