\FloatBarrier
\section{Celestials}
Creatures of the astral who are encountered in their own forms (either on the Astral plane or having traveled to another plane under their own power or that of an Ascendant) are called "celestials", after the most prominent group, the angels. Additionally, Mortal-plane creatures touched by strong amounts of Ascendant power may become MORE than they were. These two are counted as celestials--among them the heavenly flying horse, the pegasus, and the horned protector and friend to the innocent, the unicorn.
\subsection{Angels}
"Angel" most properly denotes the members of the Angelic Hosts, those sworn to the endless battle to keep the universe safe from threats from Beyond. In common parlance, however, it means all sapient Lucians who act on the orders of the Great Mechanism or various Ascendants. They inhabit either their own, often vaguely humanoid bodies of aether or wondrous shells of aether. Because their essence is of the Astral directly, they don't have the vulnerability to silver that fiends do.

Don't mistake their luminous, kindly appearances for beneficence or goodness or kindness. Most often, angels are sent to the Mortal and other planes on specific missions, often of vengeance, justice, or punishment. Some are good and kindly, others are not. But angels are almost always \textit{dutiful}, even if that means wreaking havoc on mortals.

\begin{DndMonster}{Astral Eye}
	\DndMonsterType{Small celestial}
	\DndMonsterBasics[armor-class={14}, hit-points={45}, speed={0 ft., fly 50 ft. (hover)}]
	\MonsterStats{-1}{+4}{+1}{+1}{+5}{+4}
	\DndMonsterDetails[skills={Perception +9}, damage-resistances={fire, radiant}, senses={truesight 60 ft., blindsight 60 ft., passive Perception 19}, languages={all, telepathy 60 ft.}, challenge={3:1}]
	
	\DndMonsterSection{Actions}
	
	\DndMonsterAction{Multiattack} The eye uses Blazing Beam twice. It can substitute one attack for either Divine Inspiration or Righteous Jaunt.
	
	\DndMonsterAttack[
		name=Blazing Beam,
		distance=ranged,
		type=spell,
		mod=+6,
		range=120 ft.,
		dmg=\DndDice{1d6+4},
		dmg-type=radiant
	]
	\DndMonsterAction{Divine Inspiration (1/rest)} All allies other than the eye within 60 ft. of the eye that can see it gain advantage on their next attack roll, saving throw, or ability check made within 1 minute.
	
	\DndMonsterAction{Righteous Jaunt (3/rest)} The eye teleports to an unoccupied location it can see within 120 ft. Any creature within 5 ft. of its starting location must make a DC 14 Dexterity saving throw, taking 2d6 radiant damage on a failure or half as much on a success. 
\end{DndMonster}

\begin{DndMonster}{Blade Archon, Lesser}
	\DndMonsterType{Medium celestial}
	\DndMonsterBasics[armor-class={15 (scale mail)}, hit-points={60}, speed={30 ft., fly 50 ft. (hover)}]
	\MonsterStats{4}{+2}{+2}{+0}{+3}{+2}
	\DndMonsterDetails[skills={Athletics +6}, damage-resistances={radiant}, damage-immunities={poison}, senses={darkvision 60 ft., passive Perception 19}, languages={celestial}, challenge={4:3}]

	\DndMonsterAction{Astral Edge} The blade archon's weapon attacks deal an additional 1d6 radiant damage (included in the attacks). This is an inherent property of the archon, not a property of the weapon.
	
	\DndMonsterAction{Dancing Blades} The blade archon is accompanied by two floating longswords that it uses to make attacks either in melee or at range, giving it much larger reach than normal.
	
	\DndMonsterSection{Actions}
	\DndMonsterAction{Multiattack} The Archon attacks twice with its longswords. It can use Confeitor in place of either of those attacks if available.
	\DndMonsterAttack[
		name=Floating Longsword,
		distance=both,
		type=weapon,
		mod=+6,
		reach=20 ft.,
		dmg=\DndDice{1d10+4},
		dmg-type=slashing,
		plus-dmg=\DndDice{1d6},
		plus-dmg-type=radiant
	]
	\DndMonsterAction{Confeitor (Recharge 5-6)} The floating blades spin around him, creating after-images. These images shoot outward, piercing all creatures within 15 feet. Targets must make a DC 13 Dexterity saving throw, taking 4d6 psychic damage on a failure or half as much on a success. Creatures that fail are shaken by memories of their sins and are \nameref{condition:frightened} of the archon until the end of their next turn.
\end{DndMonster}


\begin{DndMonster}{Seraph}
    \DndMonsterType{Medium Celestial (Angel)}
    \DndMonsterBasics[
        armor-class={16 (chain)},
        hit-points={104 (16d8 + 32)},
        speed={30 ft., 90 ft. fly}
    ]
    \MonsterStats{0}{+4}{+2}{+1}{+4}{+3}
    \DndMonsterDetails[
        skills={--},
        senses={darkvision 120 ft, passive perception 14},
        languages={celestial, lucian},
        challenge={9:5},
        damage-immunities={--},
        damage-resistances={radiant},
        damage-vulnerabilities={--}
    ]

    \DndMonsterAction{Distant Aid} The angel can cast any spell that only targets allies out to a range of 60 ft, even if the range is smaller. This also extends the radius of aura-type spells.
    \DndMonsterAction{Spellcasting} The seraph casts as a priest. Its spell attack modifier is +7 and its spell save DC is 15. It can cast the following spells:
    \textit{Concentration:} \nameref{spell:hold-person} (2x), \nameref{spell:bless} (1x)

    \textit{Healing:} \nameref{spell:mass-cure-wounds} (1x), \nameref{spell:cure-wounds} (at will)

    \textit{Other:} \nameref{spell:death-ward} (1x, has it active on self)

    \DndMonsterAction{Death Ward} The first time per day the seraph falls to zero hit points or is hit by an effect that would instantly kill it, it is reduced to 1 hp instead.

    \DndMonsterSection{Actions}
    \DndMonsterAttack[
        name={Sacred Lance},
        distance={both},
        type={spell},
        mod={+7},
        reach={5 ft or 90 ft},
        dmg-type={radiant},
        extra={.}
    ]
    

    \DndMonsterAction{Multiattack} The seraph makes 3 attacks. It can exchange two attacks to cast a spell.



\end{DndMonster}
\begin{DndMonster}{Valorite}
    \DndMonsterType{Medium Celestial (Angel)}
    \DndMonsterBasics[
        armor-class={18 (chain + shield)},
        hit-points={91 (14d8 + 28)},
        speed={30 ft., 60 ft. fly}
    ]
    \MonsterStats{+4}{0}{+2}{-1}{+2}{+2}
    \DndMonsterDetails[
        skills={--},
        senses={darkvision 120 ft, passive perception 12},
        languages={celestial, lucian},
        challenge={5:6},
        damage-immunities={--},
        damage-resistances={radiant},
        damage-vulnerabilities={--}
    ]

    \DndMonsterAction{Shield Wall} As long as another valorite is within 5ft, the angel makes all DEX saves at advantage and all ranged attacks against it are at disadvantage.
    \DndMonsterAction{Angelic Weapons} The angel's weapon attacks deal an extra 2d8 radiant damage (included in the attack). This is an innate property.

    \DndMonsterSection{Actions}
    \DndMonsterAttack[
        name={Gladius},
        distance={melee},
        type={weapon},
        mod={+7},
        reach={5 ft},
        dmg-type={piercing},
        extra={ plus 2d8 radiant.}
    ]
    
    \DndMonsterAttack[
        name={Javelin (3x)},
        distance={ranged},
        type={weapon},
        mod={+7},
        reach={30/120 ft},
        dmg-type={piercing},
        extra={ plus 2d8 radiant.}
    ]
    
    \DndMonsterAction{Multiattack} The angel makes 2 gladius attacks.

    \DndMonsterSection{Variants}
        \DndMonsterAction{Sergeant} Upgrade to a longsword (1d8 + 4) slashing. Gain trait Commanding Presence: All angels within 60 ft have advantage on their first attack on each of their turns. No change to ratings.
\end{DndMonster}
\FloatBarrier

\subsection{Astral-Touched Mortal Creatures}
\begin{DndMonster}{Pegasus}
\DndMonsterType{Large celestial}
\DndMonsterBasics[armor-class={12}, hit-points={59 (7d10 + 21)}, speed={60 ft., fly 90 ft.}]
\MonsterStats{+4}{+2 (+4)}{+3}{+0}{+2 (+4)}{+1 (+3)}
\DndMonsterDetails[saving-throws={}, skills={Perception +6}, damage-immunities={}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={passive Perception 16}, languages={understands Celestial, Common, Elvish, and Sylvan but can't speak}, challenge={1:2}]

\DndMonsterSection{Actions}
\DndMonsterAttack[
	name=Hooves,
	distance=melee,
	type=weapon,
	mod=+6,
	reach=5,
	dmg=\DndDice{2d6 + 4},
	dmg-type=bludgeoning
]
\end{DndMonster}

\begin{DndMonster}[float*=htb,width=\textwidth + 8pt]{Couatl}
\begin{multicols}{2}
\DndMonsterType{Medium celestial}
\DndMonsterBasics[armor-class={16 (natural armor)}, hit-points={97 (13d8 + 39)}, speed={30 ft., fly 90 ft.}]
\MonsterStats{+3}{+5}{+3 (+5)}{+4}{+5 (+7)}{+4 (+6)}
\DndMonsterDetails[saving-throws={Con +5, Wis +7, Cha +6}, skills={}, damage-immunities={psychic}, damage-resistances={radiant}, damage-vulnerabilities={}, condition-immunities={}, senses={truesight 120 ft., passive Perception 15}, languages={all, telepathy 120 ft.}, challenge={3:6}]

\DndMonsterAction{Innate Spellcasting} The couatl's spellcasting ability is Charisma (spell save DC 14). It can innately cast the following spells, requiring only verbal components:

\begin{itemize}
\item[] At will: \nameref{spell:detect-thoughts}
\item[] 3/day each: \nameref{spell:bless},  \nameref{spell:cure-wounds}, \nameref{spell:protection-from-poison}, \nameref{spell:sanctuary}, \nameref{spell:shield}
\item[] 1/day each: \nameref{spell:flame-strike}
\end{itemize} 

\DndMonsterAction{Shielded Mind} The couatl is immune to scrying and to any effect that would sense its emotions, read its thoughts, or detect its location.

\DndMonsterSection{Actions}
\DndMonsterAttack[
	name=Bite,
	distance=melee,
	type=weapon,
	mod=+6,
	reach=5,
	dmg=\DndDice{1d6 + 3},
	dmg-type=piercing,
	extra={, and the target must succeed on a DC 13 Constitution saving throw or be poisoned for 24 hours. Until this poison ends, the target is unconscious. Another creature can use an action to shake the target awake.}
]
\DndMonsterAttack[
	name=Constrict,
	distance=melee,
	type=weapon,
	mod=+6,
	reach=10,
	dmg=\DndDice{2d6 + 3},
	dmg-type=bludgeoning,
	extra={, and the target is grappled (escape DC 15). Until this grapple ends, the target is restrained, and the couatl can't constrict another target.}
]

\DndMonsterAction{Change Shape}
The couatl magically polymorphs into a humanoid or beast that has a challenge rating equal to or less than its own, or back into its true form. It reverts to its true form if it dies. Any equipment it is wearing or carrying is absorbed or borne by the new form (the couatl's choice).

In a new form, the couatl retains its game statistics and ability to speak, but its AC, movement modes, Strength, Dexterity, and other actions are replaced by those of the new form, and it gains any statistics and capabilities (except class features, legendary actions, and lair actions) that the new form has but that it lacks. If the new form has a bite attack, the couatl can use its bite in that form.
\end{multicols}
\end{DndMonster}

\begin{DndMonster}[float*=htb,width=\textwidth + 8pt]{Unicorn}
\begin{multicols}{2}
\DndMonsterType{Large celestial}
\DndMonsterBasics[armor-class={12}, hit-points={67 (9d10 + 18)}, speed={50 ft.}]
\MonsterStats{+4}{+2}{+2}{+0}{+3}{+3}
\DndMonsterDetails[saving-throws={}, skills={}, damage-immunities={poison}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={charmed, paralyzed, poisoned}, senses={darkvision 60 ft., passive Perception 13}, languages={Celestial, Elvish, Sylvan, telepathy 60 ft.}, challenge={6:4}]
\DndMonsterAction{Charge} If the unicorn moves at least 20 feet straight toward a target and then hits it with a horn attack on the same turn, the target takes an extra 9 (2d8) piercing damage. If the target is a creature, it must succeed on a DC 15 Strength saving throw or be knocked prone.

\DndMonsterAction{Innate Spellcasting} The unicorn's innate spellcasting ability is Charisma (spell save DC 14). The unicorn can innately cast the following spells, requiring no components:
\begin{itemize}
	\item[] At will: \nameref{inc:sense-aura}, \nameref{spell:pass-without-trace}
	\item[] 1/day each: \nameref{spell:calm-emotions}, \nameref{spell:dispel-otherworldly-influence}, \nameref{spell:entangle}
\end{itemize}

\DndMonsterAction{Magic Resistance} The unicorn has advantage on saving throws against spells and other magical effects.

\DndMonsterAction{Magic Weapons} The unicorn's weapon attacks are magical.

\DndMonsterSection{Actions}
\DndMonsterAction{Multiattack} The unicorn makes two attacks: one with its hooves and one with its horn.
\DndMonsterAttack[
	name=Hooves,
	distance=melee,
	type=weapon,
	mod=+7,
	reach=5,
	dmg=\DndDice{2d6 + 4},
	dmg-type=bludgeoning
]
\DndMonsterAttack[
	name=Horn,
	distance=melee,
	type=weapon,
	mod=+7,
	reach=5,
	dmg=\DndDice{1d8 + 4},
	dmg-type=piercing
]
\DndMonsterAction{Healing Horn}
The unicorn touches another creature with its horn. The target magically regains 11 (2d8 + 2) hit points. In addition, the touch removes all diseases and neutralizes all poisons afflicting the target.

\DndMonsterAction{Teleport}
The unicorn magically teleports itself and up to three willing creatures it can see within 5 feet of it, along with any equipment they are wearing or carrying, to a location the unicorn is familiar with, up to 1 mile away.


\DndMonsterSection{Legendary Actions}
The Unicorn can take 3 legendary actions, choosing from the options below. Only one legendary action option can be used at a time and only at the end of another creature's turn. The Unicorn regains spent legendary actions at the start of its turn.

\begin{DndMonsterLegendaryActions}
\DndMonsterLegendaryAction{Hooves}{The unicorn makes one attack with its hooves.}
\DndMonsterLegendaryAction{Shimmering Shield (Costs 2 Actions)}{The unicorn creates a shimmering, magical field around itself or another creature it can see within 60 feet of it. The target gains a +2 bonus to AC until the end of the unicorn's next turn.}
\DndMonsterLegendaryAction{Heal Self (Costs 3 Actions)}{The unicorn magically regains 11 (2d8 + 2) hit points.}
\end{DndMonsterLegendaryActions}
\end{multicols}
\end{DndMonster}