\subsection{}
\begin{DndMonster}[float=*b,width\textwidth + 8pt]{}
\begin{multicols}{2}
\DndMonsterType{I'm tired of having a crap-ton of dragon stat blocks}
\DndMonsterBasics[armor-class={}, hit-points={}, speed={}]
\MonsterStats{0}{0}{0}{0}{0}{0}
\DndMonsterDetails[saving-throws={}, skills={}, damage-immunities={}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={}, languages={}, challenge={0}]
\DndMonsterSection{Actions}
\end{multicols}
\end{DndMonster}
\subsection{Dragon, Wyrmling}
\begin{DndMonster}[float=*b,width\textwidth + 8pt]{Dragon, Wyrmling}
\begin{multicols}{2}
\DndMonsterType{Medium dragon}
\DndMonsterBasics[armor-class={17 (natural armor)}, hit-points={33 (6d8 + 6)}, speed={30 ft., fly 60 ft.}]
\MonsterStats{+2}{+0}{+1}{+0}{+0}{+1}
\DndMonsterDetails[saving-throws={Dex +2, Con +3, Wis +2, Cha +3}, skills={Perception +4, Stealth +4}, damage-immunities={Variable based on color}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={blindsight 10 ft., darkvision 60 ft., passive Perception 14}, languages={Draconic}, challenge={2 (450 XP)}]
\DndMonsterAction{Special Traits} Pick one of the following based on color:\\n- Amphibious: The dragon can breath both air and water.\\n- Ice Walk. The dragon can move across and climb icy surfaces without needing to make an ability check. Additionally, difficult terrain composed of ice or snow doesn’t cost it extra moment.

\DndMonsterAction{Damage Type} The dragon's bite attacks and breath weapon deal damage based on its color. The saving throw also depends on the color\\n- Black: Necrotic (CON)\\n- Blue: Lightning (DEX)\\n- Brass: Thunder (CON)\\n- Bronze or Red: Fire (DEX)\\n- Copper Acid (CON)\\n- Green: Acid (DEX)\\n- Gold: Radiant (DEX)\\n- Silver or White: Cold (CON)

\DndMonsterAction{Blast Shape} The dragon's breath weapon targets an area chosen from the list below:\\n- Ball: 10 ft radius sphere centered on a point within 60 ft\\n- Cone: 15 ft cone\\n- Line: 15 ft line that is 5 ft wide

\DndMonsterSection{Actions}
**_Bite_**. _Melee Weapon Attack:_ +4 to hit, reach 5 ft., one target. _Hit:_ 7 (1d10 + 2) piercing damage plus 2 (1d4) damage of the dragon's damage type.
The dragon exhales energy in a burst (see Blast Shape). Each creature in that area must make a DC 12 saving throw (DEX or CON based on color), taking 22 (6d6) damage of the dragon's Damage Type on a failed save, or half as much damage on a successful one.
\subsection{Scaling}
To make a CR 3 version, add 2 hit dice (15 HP), increase the ability scores to 18/12/16/13/11/15, which increases the bite's modifier to +6 and the damage to 1d10 + 4 + 2d4. Don't change the breath weapon.
To make a CR 4 version, add 6 hit dice (45 HP), increase the ability scores to 19/12/17/13/11/15, which increases the bite's modifier to +6 and the damage to 1d10 + 4 + 2d4. The breath weapon goes up to DC 13 and gains 1d6 damage.
\end{multicols}
\end{DndMonster}
\subsection{Dragon, Young}
\begin{DndMonster}[float=*b,width\textwidth + 8pt]{Dragon, Young}
\begin{multicols}{2}
\DndMonsterType{Large dragon}
\DndMonsterBasics[armor-class={18 (natural armor)}, hit-points={142 (15d10 + 45)}, speed={40 ft., fly 80 ft.}]
\MonsterStats{+5}{+1}{+3}{+2}{+1}{+3}
\DndMonsterDetails[saving-throws={Dex +4, Con +7, Wis +4, Cha +6}, skills={Perception +6, +one}, damage-immunities={Variable by color}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={blindsight 30 ft., darkvision 120 ft., passive Perception 16}, languages={Common, Draconic}, challenge={8}]
\DndMonsterAction{Special Traits} Pick one of the following based on color:\\n- Amphibious: The dragon can breath both air and water.\\n- Ice Walk. The dragon can move across and climb icy surfaces without needing to make an ability check. Additionally, difficult terrain composed of ice or snow doesn’t cost it extra moment.\\n- Shapechange. The dragon can use its action to transform into a humanoid shape of Medium or smaller size. Its features other than its size do not change.\\n- Spellcasting. The dragon can cast spells. These generally should be defensive, control, or utility spells, as the dragon has enough raw firepower. Its spellcasting DC is 15.\\n- Pick 3 spells that cost 2 AET. It can cast these 3x/day each.\\n- Pick 2 spells that cost 3-4 AET. It can cast these 2x/day each.\\n- Pick 1 spell that costs 5 AET. It can cast this once/day.

\DndMonsterAction{Damage Type} The dragon's bite attacks and breath weapon deal damage based on its color. The saving throw also depends on the color\\n- Black: Necrotic (CON)\\n- Blue: Lightning (DEX)\\n- Brass: Thunder (CON)\\n- Bronze or Red: Fire (DEX)\\n- Copper Acid (CON)\\n- Green: Acid (DEX)\\n- Gold: Radiant (DEX)\\n- Silver or White: Cold (CON)

\DndMonsterAction{Blast Shape} The dragon's breath weapon targets an area chosen from the list below:\\n- Ball: 15 ft radius sphere centered on a point within 60 ft\\n- Cone: 30 ft cone\\n- Line: 30 ft line that is 5 ft wide

\DndMonsterSection{Actions}
\DndMonsterAction{Multiattack} The dragon makes three attacks: one with its bite and two with its claws.
**_Bite_**. _Melee Weapon Attack:_ +8 to hit, reach 10 ft., one target. _Hit:_ 16 (2d10 + 5) piercing damage plus 4 (1d8) damage of the dragon's Damage Type.
**_Claw_**. _Melee Weapon Attack:_ +8 to hit, reach 5 ft., one target. _Hit:_ 12 (2d6 + 5) slashing damage.
The dragon exhales energy in a burst (see Blast Shape). Each creature in that area must make a DC 15 saving throw (DEX or CON based on color), taking 45 (10d8) damage of the dragon's Damage Type on a failed save, or half as much damage on a successful one.
\subsection{Scaling}
To increase the dragon's CR to 9, increase its hit points by 15 (no extra hit dice but +2 CON), increase its STR, CON, and CHA by 2 each (increasing its attack modifier by 2 and physical damage modifiers by +1), its breath weapon by 9 (+2d8), and its save DC (spellcasting and breath weapon) by 1 (to 16). Its proficiency goes up by 1 as well.
To increase the dragon's CR to 10, increase its hit points by 21 (adding 2 hit dice), increase its STR, CON, and CHA by +3 and its DEX by 1 (increasing its attack modifier by 2 and physical damage modifiers by +1). Its breath weapon increases by 9 (+2d8) and its save DC increases by 2 (to 17). Its proficiency goes up by 1 as well.
\end{multicols}
\end{DndMonster}
\subsection{Dragon, Adult}
\begin{DndMonster}[float=*b,width\textwidth + 8pt]{Dragon, Adult}
\begin{multicols}{2}
\DndMonsterType{Huge dragon}
\DndMonsterBasics[armor-class={19 (natural armor)}, hit-points={195 (17d12 + 85)}, speed={40 ft., fly 80 ft., swim 40 ft.}]
\MonsterStats{+6}{+1}{+5}{+3}{+2}{+3}
\DndMonsterDetails[saving-throws={Dex +6, Con +10, Wis +6, Cha +8}, skills={Perception +11, Stealth +7}, damage-immunities={Variable based on color}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={blindsight 60 ft., darkvision 120 ft., passive Perception 21}, languages={Common, Draconic}, challenge={14}]
\DndMonsterAction{Special Traits} Pick one of the following based on color:\\n- Amphibious: The dragon can breath both air and water.\\n- Ice Walk. The dragon can move across and climb icy surfaces without needing to make an ability check. Additionally, difficult terrain composed of ice or snow doesn’t cost it extra moment.\\n- Shapechange. The dragon can use its action to transform into a humanoid shape of Medium or smaller size. Its features other than its size do not change.\\n- Spellcasting. The dragon can cast spells. These generally should be defensive, control, or utility spells, as the dragon has enough raw firepower. Its spellcasting DC is 16.\\n- Pick 3 spells that cost 4 AET or less. It can cast these 3x/day each. \\n- Pick 2 spells that cost 7 AET or less. It can cast these 2x/day each.\\n- Pick 1 spell that costs 8 AET. It can cast this once/day.

\DndMonsterAction{Damage Type} The dragon's bite attacks and breath weapon deal damage based on its color. The saving throw also depends on the color\\n- Black: Necrotic (CON)\\n- Blue: Lightning (DEX)\\n- Brass: Thunder (CON)\\n- Bronze or Red: Fire (DEX)\\n- Copper Acid (CON)\\n- Green: Acid (DEX)\\n- Gold: Radiant (DEX)\\n- Silver or White: Cold (CON)

\DndMonsterAction{Blast Shape} The dragon's breath weapon targets an area chosen from the list below:\\n- Ball: 20 ft radius sphere centered on a point within 90 ft\\n- Cone: 60 ft cone\\n- Line: 60 ft line that is 10 ft wide

\DndMonsterAction{Legendary Resistance (3/Day)} If the dragon fails a saving throw, it can choose to succeed instead.

\DndMonsterSection{Actions}
\DndMonsterAction{Multiattack} The dragon can use its Frightful Presence. It then makes three attacks: one with its bite and two with its claws.
**_Bite_**. _Melee Weapon Attack:_ +11 to hit, reach 10 ft., one target. _Hit:_ 17 (2d10 + 6) piercing damage plus 4 (1d8) damage of the dragon's Damage Type.
**_Claw_**. _Melee Weapon Attack:_ +11 to hit, reach 5 ft., one target. _Hit:_ 13 (2d6 + 6) slashing damage.
**_Tail_**. _Melee Weapon Attack:_ +11 to hit, reach 15 ft., one target. _Hit:_ 15 (2d8 + 6) bludgeoning damage.
Each creature of the dragon’s choice that is within 120 feet of the dragon and aware of it must succeed on a DC 16 Wisdom saving throw or become frightened for 1 minute. A creature can repeat the saving throw at the end of each of its turns, ending the effect on itself on a success. If a creature’s saving throw is successful or the effect ends for it, the creature is immune to the dragon’s Frightful Presence for the next 24 hours.
The dragon exhales energy in a burst (see Blast Shape). Each creature in that area must make a DC 18 Dexterity saving throw, taking 54 (12d8) acid damage on a failed save, or half as much damage on a successful one.
\DndMonsterSection{{Legendary Actions}}
The Dragon, Adult can take 3 legendary actions, choosing from the options below. Only one legendary action option can be used at a time and only at the end of another creature's turn. The Dragon, Adult regains spent legendary actions at the start of its turn.
\begin{{DndMonsterLegendaryActions}}
\DndMonsterLegendaryAction{Detect}{The dragon makes a Wisdom (Perception) check.}
\DndMonsterLegendaryAction{Tail Attack}{The dragon makes a tail attack.}
\DndMonsterLegendaryAction{Wing Attack (Costs 2 Actions)}{The dragon beats its wings. Each creature within 10 feet of the dragon must succeed on a DC 19 Dexterity saving throw or take 13 (2d6 + 6) bludgeoning damage and be knocked prone. The dragon can then fly up to half its flying speed.}
\subsection{Scaling}
To increase the dragon's CR to 15, increase its hit points to 212, its ability scores to 25/12/23/17/15/17 (increasing its attack bonus to +12 and its weapon damage output by 1), its breath weapon to 63 (14d8), and increase each of its saving throw DCs by 1.
To increase the dragon's CR to 16, increase its hit points to 225, its ability scores to 25/12/23/16/15/19, its breath weapon to 67 (15d8), its frightful presence DC to 18, its breath weapon DC to 19, and its wing attack DC to 20.
\end{multicols}
\end{DndMonster}
\subsection{Dragon, Ancient}
\begin{DndMonster}[float=*b,width\textwidth + 8pt]{Dragon, Ancient}
\begin{multicols}{2}
\DndMonsterType{Gargantuan dragon}
\DndMonsterBasics[armor-class={22 (natural armor)}, hit-points={350 (20d20 + 140)}, speed={40 ft., fly 80 ft.}]
\MonsterStats{+8}{+1}{+7}{+4}{+2}{+4}
\DndMonsterDetails[saving-throws={Dex +8, Con +14, Wis +9, Cha +11}, skills={Perception +16, Stealth +9}, damage-immunities={Variable based on color}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={blindsight 60 ft., darkvision 120 ft., passive Perception 26}, languages={Common, Draconic}, challenge={21}]
\DndMonsterAction{Special Traits} Pick one of the following based on color:\\n- Amphibious: The dragon can breath both air and water.\\n- Ice Walk. The dragon can move across and climb icy surfaces without needing to make an ability check. Additionally, difficult terrain composed of ice or snow doesn’t cost it extra moment.\\n- Shapechange. The dragon can use its action to transform into a humanoid shape of Medium or smaller size. Its features other than its size do not change.\\n- Spellcasting. The dragon can cast spells. These generally should be defensive, control, or utility spells, as the dragon has enough raw firepower. Its spellcasting DC is 19.\\n- Pick 3 spells that cost 5 AET or less. It can cast these 3x/day each. \\n- Pick 2 spells that cost 8 AET or less. It can cast these 2x/day each.\\n- Pick 1 spell that costs 13 AET. It can cast this once/day.

\DndMonsterAction{Damage Type} The dragon's bite attacks and breath weapon deal damage based on its color. The saving throw also depends on the color\\n- Black: Necrotic (CON)\\n- Blue: Lightning (DEX)\\n- Brass: Thunder (CON)\\n- Bronze or Red: Fire (DEX)\\n- Copper Acid (CON)\\n- Green: Acid (DEX)\\n- Gold: Radiant (DEX)\\n- Silver or White: Cold (CON)

\DndMonsterAction{Blast Shape} The dragon's breath weapon targets an area chosen from the list below:\\n- Ball: 40 ft radius sphere centered on a point within 120 ft\\n- Cone: 90 ft cone\\n- Line: 90 ft line that is 10 ft wide

\DndMonsterAction{Legendary Resistance (3/Day)} If the dragon fails a saving throw, it can choose to succeed instead.

\DndMonsterSection{Actions}
\DndMonsterAction{Multiattack} The dragon can use its Frightful Presence. It then makes three attacks: one with its bite and two with its claws.
**_Bite_**. _Melee Weapon Attack:_ +15 to hit, reach 15 ft., one target. _Hit:_ 19 (2d10 + 8) piercing damage plus 9 (2d8) damage of the dragon's Damage Type.
**_Claw_**. _Melee Weapon Attack:_ +15 to hit, reach 10 ft., one target. _Hit:_ 15 (2d6 + 8) slashing damage.
**_Tail_**. _Melee Weapon Attack:_ +15 to hit, reach 20 ft., one target. _Hit:_ 17 (2d8 + 8) bludgeoning damage.
Each creature of the dragon’s choice that is within 120 feet of the dragon and aware of it must succeed on a DC 19 Wisdom saving throw or become frightened for 1 minute. A creature can repeat the saving throw at the end of each of its turns, ending the effect on itself on a success. If a creature’s saving throw is successful or the effect ends for it, the creature is immune to the dragon’s Frightful Presence for the next 24 hours.
The dragon exhales energy in an area (see Blast Shape). Each creature in that area must make a DC 22 saving throw of the type indicated for the dragon's color, taking 71 (13d10) damage of the dragon's Damage Type on a failed save, or half as much damage on a successful one.
\DndMonsterSection{{Legendary Actions}}
The Dragon, Ancient can take 3 legendary actions, choosing from the options below. Only one legendary action option can be used at a time and only at the end of another creature's turn. The Dragon, Ancient regains spent legendary actions at the start of its turn.
\begin{{DndMonsterLegendaryActions}}
\DndMonsterLegendaryAction{Detect}{The dragon makes a Wisdom (Perception) check.}
\DndMonsterLegendaryAction{Tail Attack}{The dragon makes a tail attack.}
\DndMonsterLegendaryAction{Wing Attack (Costs 2 Actions)}{The dragon beats its wings. Each creature within 15 feet of the dragon must succeed on a DC 23 Dexterity saving throw or take 15 (2d6 + 8) bludgeoning damage and be knocked prone. The dragon can then fly up to half its flying speed.}
\subsection{Scaling}
To increase the dragon's CR to 22, increase its HP to 385 (22d20+154), its AC to 22, its breath weapon damage to 77 (14d10), and its saving throw DCs to 20 (Frightful Presence and spellcasting), 23 (Breath weapon), and 24 (Wing Attack).
To increase the dragon's CR to 23, increase its HP to 487 (25d20+225), its AC to 23, its ability scores to 30/12/29/18/15/23 (increasing its attack bonus to +17 and the damage modifier of each physical attack to +10), its breath weapon damage to 88 (16d10), and its saving throw DCs to 21 (Frightful Presence and spellcasting), 24 (Breath weapon), and 25 (Wing Attack)/
\end{multicols}
\end{DndMonster}
