\clearpage
\section{Fey}
The fey are non-mortal, sometimes-embodied collections of the spirits that inhabit the natural world and Shadow (aka ihminen or kami). These beings, when exposed to mortal influences and strong emotions, often conglomerate and take form, mimicing the acts and emotions that brought them forth. This leads to what appears to be obsessive behavior--a fey born of violence will rampage destructively or tempt others into such acts; a fey born of intense romantic feelings may become a serial seducer. The longer the fey exists without being destroyed or losing interest, however, the more this obsession develops into a full-on mimicry of mortal life in many aspects. They retain an interest in that one particular obsession, but it generally broadens and becomes less of a caricature. Make no mistake--fey are dangerous. They do not, except the strongest and oldest, understand \textit{death}. The more humanoid and "civilized" a fey, the more powerful.

Some fey, in contrast to this strongly mortal-centric view, are born from elemental influences or those of strong, highly-magical areas (especially trees, for some reason). These are place-bound fey, bound to a particular natural feature. They often form deep, almost symbiotic bonds with mortal creatures, trading vastly extended life and power to the mortal in exchange for a shared consciousness and being bound close to one location for the rest of their existence. These fey tend to be much less fickle, if bound to a mortal. The mortal provides an understanding of death, of change, and of the mortal world. Instead, these fey tend to be very fiercely protective of the natural feature they're bound to.

Fey cross over into Shadow naturally, existing on both sides of the veil almost simultaneously. Many of the most powerful make their homes there, rarely manifesting on the Mortal. Others have no home and flit about wherever their fancy (and obsession) take them. 

\begin{DndMonster}{Blink Dog}
\DndMonsterType{Medium fey}
\DndMonsterBasics[armor-class={13}, hit-points={22 (4d8 + 4)}, speed={40 ft.}]
\MonsterStats{+1}{+3}{+1}{+0}{+1}{+0}
\DndMonsterDetails[saving-throws={}, skills={Perception +3, Stealth +5}, damage-immunities={}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={passive Perception 13}, languages={Blink Dog, understands Sylvan but can't speak it}, challenge={1/4:1/2}]
\DndMonsterAction{Keen Hearing and Smell} The dog has advantage on Wisdom (Perception) checks that rely on hearing or smell.

\DndMonsterSection{Actions}
\DndMonsterAttack[
	name=Bite,
	distance=melee,
	type=weapon,
	mod=+3,
	reach=5,
	dmg=\DndDice{1d6 + 1},
	dmg-type=piercing
]
\DndMonsterAction{Blink}
The dog magically teleports, along with any equipment it is wearing or carrying, up to 40 feet to an unoccupied space it can see. Before or after teleporting, the dog can make one bite attack.
\end{DndMonster}


\begin{DndMonster}{Kipuliin, Lesser}
    \DndMonsterType{Small Fey}
    \DndMonsterBasics[
        armor-class={14 (natural armor)},
        hit-points={51 (6d8 + 24)},
        speed={30 ft.}
    ]
    \MonsterStats{+3}{+2}{+4 (+6)}{-1}{-1}{+2}
    \DndMonsterDetails[
        skills={--},
        senses={darkvision 60 ft, passive perception 9},
        languages={Common, Sylvan},
        challenge={2:3},
        damage-immunities={--},
        damage-resistances={all damage when bloodied},
        damage-vulnerabilities={--}
    ]

    \DndMonsterAction{Masochism} While bloodied, the kipuliin has resistance to all damage and regains 2hp the first time it hits with an attack on a turn. If the target hit is bloodied, it regains 10 hp instead.
    \DndMonsterAction{Sadism} When the kipuliin hits a bloodied creature with a weapon attack, its weapons deal an additional 1d4 damage.

    \DndMonsterSection{Actions}
		\DndMonsterAction{Multiattack} The kipuliin makes two barbed whip attacks.
    \DndMonsterAttack[
        name={Barbed Whip},
        distance={melee},
        type={weapon},
        mod={+5},
        reach={10 ft},
        dmg-type={slashing},
        extra={ plus 1d4 extra against bloodied targets.}
    ]
\end{DndMonster}

\begin{DndMonster}{Satyr}
	\DndMonsterType{Medium fey}
	\DndMonsterBasics[armor-class={13}, hit-points={31 (7d8)}, speed={40 ft.}]
	\MonsterStats{+1}{+3}{+0}{+1}{+0}{+2}
	\DndMonsterDetails[saving-throws={}, skills={Perception +2, Performance +6, Stealth +5}, damage-immunities={}, damage-resistances={}, damage-vulnerabilities={}, condition-immunities={}, senses={passive Perception 12}, languages={Common, Elvish, Sylvan}, challenge={1/2:1/2}]
	\DndMonsterAction{Magic Resistance} The satyr has advantage on saving throws against spells and other magical effects.
	
	\DndMonsterSection{Actions}
	\DndMonsterAttack[
		name=Ram,
		distance=melee,
		type=weapon,
		mod=+3,
		reach=5,
		dmg=\DndDice{2d4 + 1},
		dmg-type=bludgeoning
	]
	\DndMonsterAttack[
		name=Shortsword,
		distance=melee,
		type=weapon,
		mod=+5,
		reach=5,
		dmg=\DndDice{1d6 + 3},
		dmg-type=piercing
	]
	\DndMonsterAttack[
		name=Shortbow,
		distance=ranged,
		type=weapon,
		mod=+5,
		range=80/320,
		dmg=\DndDice{1d6 + 3},
		dmg-type=piercing
	]

	\DndMonsterAction{Aura of Drunkenness} The satyr's nature overflows. All creatures within 30 feet must make a DC 12 Constitution saving throw. On a failed save they are poisoned for one minute. Creatures that end their turn more than 30 feet from the satyr can reattempt the saving throw at the end of each of their turns, ending the effect on a success.
\end{DndMonster}