\section{Armor, Magic}\label{mi:armor}

This section contains the various magic items that are worn as armor or clothing, as well as shields.

\subparagraph*{Boots, belts, cloaks, gloves, and helmets}
Even though many types of armor (especially heavy armor) have integrated sabatons (armored boots), belts, gauntlets, and helmets, you can wear magic items listed as boots, belts, cloaks, gloves (or gauntlets), or helmets (or hats) without reducing the protective qualities of the armor. Some magic items (such as the Demon Armor) may impose more specific restrictions. You must wear both of a pair of magic gloves or magic boots to gain any benefit from them. You can only gain the benefits of one pair of boots, one cloak, one pair of gloves or gauntlets, one helmet or hat, and one belt at the same time, even if they do not require attunement.

\subparagraph*{Robes}
Robes cannot be worn with any armor, magical or not, except a mithral breastplate or chain shirt (or armor like Aelven Chain, which acts as a mithral chain shirt).

\subsection{Adamantine Armor}

\textit{Armor (medium or heavy, but not hide), Adventurer}

This suit of armor is reinforced with adamantine, one of the hardest substances in existence. While you're wearing it, any critical hit against you becomes a normal hit.

\subparagraph*{Formula} To craft Adamantine Armor, you must procure one-quarter of the weight of the armor in refined adamantine metal (three-quarters of the weight in raw adamantine ore). The cost of the crafting is twice the base price of the armor, four times if you wish to have masterwork adamantine armor. Crafting adamantine armor requires a +3 proficiency bonus (+4 if you are crafting masterwork) and proficiency in smith's tools.

\subsection{Aelven Chain}
Armor (masterwork chain shirt), Hero

This also counts as mithril. You are considered proficient with this armor even if you lack proficiency with medium armor.

\subsection{Animated Shield}
\textit{Armor (shield), Hero (requires attunement)} 

While holding this shield, you can speak its command word as a bonus action to cause it to animate. The shield leaps into the air and hovers in your space to protect you as if you were wielding it, leaving your hands free. The shield remains animated for 1 minute, until you use a bonus action to end this effect, or until you are incapacitated or die, at which point the shield falls to the ground or into your hand if you have one free.

\subsection{Armor of Invulnerability}
\textit{Armor (plate), legendary (requires attunement)}

You have resistance to all damage while you wear this armor. Additionally, you can use an action to make yourself immune to damage from non-legendary effects for 10 minutes or until you are no longer wearing the armor. Once this special action is used, it can't be used again until the next dawn.

\subsection{Armor of Energy Resistance}
\textit{Armor (light, medium, or heavy), Hero (requires attunement)}

You have resistance to one type of damage while you wear this armor. The GM chooses the type or determines it randomly from the options below.
\begin{DndTable}{llXll}
    \textbf{d8} & \textbf{Damage Type} & & \textbf{d8} &  \textbf{Damage Type} \\ 
    1  &  Acid      &&    5  &  Necrotic \\     
    2  &  Cold      &&    6  & Psychic        \\
    3  &  Fire      &&    7  & Radiant       \\
    4  &  Lightning &&    8  & Thunder         \\
\end{DndTable}
     
\subsection*{Armor of Physical Resistance}
\textit{Armor (light, medium, or heavy), Hero (requires attunement)} 

While wearing this armor, you have resistance to one of the following damage types: bludgeoning, piercing, or slashing. The GM chooses the type or determines it randomly.

\subsection{Arrow-Catching Shield}
\textit{Armor (shield), Hero (requires attunement)} 

You gain half cover against ranged attacks while you wield this shield. This bonus is in addition to the shield's normal bonus to AC. In addition, whenever an attacker makes a ranged attack against a target within 5 feet of you, you can use your reaction to become the target of the attack instead.

\subsection{Belt of Giant Strength}
Belt, varying tier (requires attunement)

While wearing this belt, your Strength score changes to a score granted by the belt. If your Strength is already equal to or greater than the belt's score, the item has no effect on you.

Six varieties of this belt exist, corresponding with and having rarity according to the six kinds of true giants. The belt of stone giant strength and the belt of frost giant strength look different, but they have the same effect.

\begin{DndTable}{Xll}
\textbf{Type}    & \textbf{Strength} & \textbf{Tier} \\    
Hill giant        & +5  &      Hero       \\
Stone/frost giant & +6  &      Hero       \\
Fire giant        & +7  &      Legendary  \\
Cloud giant       & +8  &      Legendary  \\
Storm giant       & +9  &      Legendary  \\
\end{DndTable}

\subsection{Boots of Levitation}
Boots, Hero (requires attunement) 

While you wear these boots, you can use an action to cast the levitate spell on yourself at will.

\subsection{Boots of Speed}
Boots, Hero (requires attunement) 

While you wear these boots, you can use a bonus action and click the boots' heels together. If you do, the boots double your walking speed, and any creature that makes an opportunity attack against you has disadvantage on the attack roll. If you click your heels together again, you end the effect; otherwise, the effect ends after 1 minute. The boots have three charges; each use expends a charge and all expended charges are restored at dawn.

\subsection{Boots of Striding and Springing}
Boots, Adventurer (requires attunement) 

While you wear these boots, your walking speed becomes 30 feet, unless your walking speed is higher, and you have advantage on any saving throw or ability check made to avoid being grappled or restrained. In addition, you can jump three times the normal distance, though you can't jump farther than your remaining movement would allow.

\subsection{Boots of the Winterlands}
Boots, Adventurer (requires attunement)

These furred boots are snug and feel quite warm. While you wear them, you gain the following benefits:
\begin{itemize}
 \item You have resistance to cold damage.
 \item You ignore difficult terrain created by ice or snow.
 \item You can tolerate temperatures as low as \textminus50degrees Fahrenheit without any additional protection. If you wear heavy clothes, you can tolerate temperatures as low as \textminus100 degreesFahrenheit.
\end{itemize}

\subsection{Cape of the Mountebank}
Cloak, Hero

This cape smells faintly of brimstone. While wearing it, you can use it to cast the \nameref{spell:dimension-door} spell as an action. This property of the cape can't be used again until the next dawn. When you disappear, you leave behind a cloud of smoke, and you appear in a similar cloud of smoke at your destination. The smoke lightly obscures the space you left and the space you appear in, and it dissipates at the end of your next turn. A light or stronger wind disperses the smoke.

\subsection{Circlet of Blasting}
Hat, Adventurer

While wearing this circlet, you can use an action to cast the \nameref{spell:scorching-ray} spell with it. When you make the spell's attacks, you do so with an attack bonus of +5. The circlet can't be used this way again until the next dawn.

\subsection{Cloak of Arachnida}
Cloak, Hero (requires attunement) 

This fine garment is made of black spider-silk interwoven with faint silvery threads. While wearing it, you gain the following benefits:
\begin{itemize}
 \item You have resistance to poison damage.
 \item You have a climbing speed equal to your walking speed.
 \item You can move up, down, and across vertical surfaces and upside down along ceilings, while leaving your hands free.
 \item You can't be caught in webs of any sort and can move through webs as if they were difficult terrain.
 \item You can use an action to cast the web spell (save DC 13). The web created by the spell fills twice its normal area. Once used, this property of the cloak can't be used again until the next dawn.
\end{itemize}

\subsection{Cloak of Displacement}
Cloak, Hero (requires attunement) 

While you wear this cloak, it projects an illusion that makes you appear to be standing in a place near your actual location, causing any creature to have disadvantage on attack rolls against you. If you take damage, the property ceases to function until the start of your next turn. This property is suppressed while you are incapacitated, restrained, or otherwise unable to move.

\subsection{Cloak of Protection}
Cloak, Adventurer (requires attunement)

When you take the \nameref{action:deflect} action or the \nameref{action:exert} action (the latter on a saving throw), you gain a +1 bonus to the effect.

\subsection{Cloak of the Bat}
Cloak, Hero (requires attunement)

While wearing this cloak, you have advantage on Dexterity (Stealth) checks. In an area of dim light or darkness, you can grip the edges of the cloak with both hands and use it to fly at a speed of 40 feet. If you ever fail to grip the cloak's edges while flying in this way, or if you are no longer in dim light or darkness, you lose this flying speed.

While wearing the cloak in an area of dim light or darkness, you can use your action to cast \nameref{spell:polymorph} on yourself, transforming into a bat. While you are in the form of the bat, you retain your Intelligence, Wisdom, and Charisma scores. The cloak can't be used this way again until the next dawn.

\subsection{Cloak of the Manta Ray}
Cloak, Adventurer

While wearing this cloak with its hood up, you can breathe underwater, and you have a swimming speed of 60 feet. Pulling the hood up or down requires an action.

\subsection{Demon Armor}
\textit{Armor (masterwork plate), Adventurer (requires attunement)}

While wearing this armor, you can understand and speak Abyssal. In addition, the armor's clawed gauntlets are light melee weapons that deal slashing damage with a +2 bonus to damage rolls and a damage die of 1d8. You have proficiency in these weapons. You cannot wear any other gloves while attuned to this armor.

\subsection{Dragon Scale Mail}
Armor (masterwork scale mail), Hero (requires attunement) 

Dragon scale mail is made of the scales of one kind of dragon. Sometimes dragons collect their cast-off scales and gift them to humanoids. Other times, hunters carefully skin and preserve the hide of a dead dragon. In either case, dragon scale mail is highly valued.  While wearing this armor, you have advantage on saving throws against the Frightful Presence and breath weapons of dragons, and you have resistance to one damage type that is determined by the kind of dragon that provided the scales (see the table).  Additionally, you can focus your senses as an action to magically discern the distance and direction to the closest dragon within 30 miles of you that is of the same type as the armor. This special action can't be used again until the next dawn.

\begin{DndTable}{llXll}
Dragon  & Resistance && Dragon & Resistance \\ 
Black &  Acid      &&  Gold  &  Fire        \\
Blue  &  Lightning &&  Green &  Poison \\      
Brass &  Fire      &&  Red   & Fire     \\   
Bronze & Lightning &&  Silver & Cold      \\  
Copper & Acid      &&  White &  Cold \\
\end{DndTable}

\subsection{Dwarven Plate}
Armor (masterwork plate), Hero 

This armor counts as adamantine. In addition, if an effect moves you against your will along the ground, you can use your reaction to reduce the distance you are moved by up to 10 feet. You also have advantage on any check or saving throw made to resist or escape a grapple.

\subsection{Gauntlets of Ogre Power}
Gloves, Adventurer (requires attunement)

Your Strength score is +4 while you wear these gauntlets. They have no effect on you if your Strength is already +4 or higher.

\subsection{Glamoured reinforced Leather}
Armor (masterwork reinforced leather), Adventurer 

While wearing this armor, you can use a bonus action to cast \nameref{spell:disguise-self} at will. Your spell save DC for this is 8 + your Intelligence modifier + your proficiency bonus.

\subsection{Gloves of Missile Snaring}
Gloves, Adventurer (requires attunement) 

These gloves seem to almost meld into your hands when you don them. When a ranged weapon attack hits you while you're wearing them, you can use your reaction to reduce the damage by 1d10 + your Dexterity modifier, provided that you have a free hand. If you reduce the damage to 0, you can catch the missile if it is small enough for you to hold in that hand. The damage reduction provided by these gloves stacks with the Brawler ability Deflect Missiles.

\subsection{Gloves of Swimming and Climbing}
Gloves, Adventurer (requires attunement) 

While wearing these gloves, climbing and swimming don't cost you extra movement, and you gain a +5 bonus to Strength (Athletics) checks made to climb or swim.

\subsection{Hat of Disguise}
Hat, Adventurer (requires attunement) 

While wearing this hat, you can use an action to cast the \nameref{spell:disguise-self} spell from it at will (DC 10). The spell ends if the hat is removed.

\subsection{Headband of Intellect}
Hat, Adventurer (requires attunement) 

Your Intelligence score is +4 while you wear this headband. It has no effect on you if your Intelligence is already +4 or higher.

\subsection{Helm of Brilliance}
Hat, Hero (requires attunement) 

This dazzling helm is set with 5 diamonds, 10 rubies, 15 fire opals, and 20 opals. Any gem pried from the helm crumbles to dust. When all the gems are removed or destroyed, the helm loses its magic. You gain the following benefits while wearing it:
\begin{itemize}
    \item You can use an action to cast one of the following spells (save DC 18), using one of the helm's gems of the specified type as a component: \nameref{spell:daylight} (opal), \nameref{spell:fireball} (fire opal), \nameref{spell:true-seeing} (diamond), or \nameref{spell:wall-of-fire} (ruby). The gem is destroyed when the spell is cast and disappears from the helm.
    \item As long as it has at least one diamond, the helm emits dim light in a 30-foot radius when at least one undead is within that area. Any undead that starts its turn in that area takes 1d6 radiant damage.
    \item As long as the helm has at least one ruby, you have resistance to fire damage.
    \item As long as the helm has at least one fire opal, you can use an action and speak a command word to cause one weapon you are holding to burst into flames. The flames emit bright light in a 10-foot radius and dim light for an additional 10 feet. Theflames are harmless to you and the weapon. When you hit with an attack using the blazing weapon,the target takes an extra 1d6 fire damage. The flames last until you use a bonus action to speak the command word again or until you drop or stow the weapon.
\end{itemize}

\subsection{Helm of Comprehending Languages}
Hat, Adventurer

While wearing this helm, you understand all spoken languages and can be understood by anyone who speaks at least one language.

\subsection{Helm of Telepathy}
Hat, Adventurer (requires attunement)

While wearing this helm, you can use an action to cast the \nameref{spell:detect-thoughts} spell (save DC 13) from it. As long as you maintain concentration on the spell, you can use a bonus action to send a telepathic message to a creature you are focused on. It can reply-using
a bonus action to do so-while your focus on it continues. While focusing on a creature with detect thoughts, you can use an action to cast the
\nameref{spell:suggestion} spell (save DC 13) from the helm on that creature. Once used, the suggestion property can't be used again until the next dawn.

\subsection{Helm of Teleportation}
Hat, Hero (requires attunement)

This helm has 3 charges. While wearing it, you can use an action and expend 1 charge to cast the \nameref{inc:teleport} incantation from it. The helm regains 1 expended charge at dawn the first day of each week.

\subsection{Ihmisi Boots}
Boots, Adventurer 

While you wear these boots, your steps make no sound, regardless of the surface you are moving across. You also have advantage on Dexterity (Stealth) checks that rely on moving silently.

\subsection{Mantle of Spell Resistance}
Cloak, Hero (requires attunement)

You have advantage on saving throws against spells while you wear this cloak.

\subsection{Mithral Armor}
Armor (medium or heavy, but not hide), Adventurer 

Mithral is a light, flexible metal. A mithral chain shirt or breastplate can be worn under normal clothes. If the armor normally imposes disadvantage on Dexterity (Stealth) checks or has a Strength requirement, the mithral version of the armor doesn't.

\subparagraph*{Formula} To craft mithral armor, you must procure one-quarter of the weight of the armor in refined mithral metal (three-quarters of the weight in raw mithral ore). The cost of the crafting is twice the base price of the armor or four times if you wish to have masterwork mithral armor. Crafting mithral armor requires a +3 proficiency bonus (+4 if you are crafting masterwork) and proficiency in smith's tools.

\subsection{Plate Armor of Etherealness}
Armor (plate), legendary (requires attunement) 

While you're wearing this armor, you can speak its command word as an action to gain the effect of the \nameref{spell:etherealness} legendary effect, which last for 10 minutes or until you remove the armor or use an action to speak the command word again. This property of the armor can't be used again until the next dawn.

\subsection{Robe of Eyes}
Robe, Hero (requires attunement)

This robe is adorned with eye like patterns. While you wear the robe, you gain the following benefits:
\begin{itemize}
 \item The robe lets you see in all directions, and you have advantage on Wisdom (Perception) checks that rely on sight.
 \item You have darkvision out to a range of 120 feet.
 \item You can see invisible creatures and objects, as well as see into the Ethereal Plane, out to a range of120 feet.The eyes on the robe can't be closed or averted.Although you can close or avert your own eyes, you are never considered to be doing so while wearing this robe.
\end{itemize}

A light spell cast on the robe or a daylight spell cast within 5 feet of the robe causes you to be blinded for 1 minute. At the end of each of your turns, you can make a Constitution saving throw (DC 11 for light or DC 15 for daylight), ending the blindness on a success.

\subsection{Robe of Scintillating Colors}
Robe, Hero (requires attunement)

This robe has 3 charges, and it regains 1d3 expended charges daily at dawn. While you wear it, you can use an action and expend 1 charge to cause the garment to display a shifting pattern of dazzling hues until the end of your next turn. During this time, the robe sheds bright light in a 30-foot radius and dim light for an additional 30 feet. Creatures that can see you have disadvantage on attack rolls against you. In addition, any creature in the bright light that can see you when the robe's power is activated must succeed on a DC 15 Wisdom saving throw or become stunned until the effect ends.

\subsection{Robe of Stars}
Wondrous item, Hero (requires attunement) 

This black or dark blue robe is embroidered with small white or silver stars. Six stars, located on the robe's upper front portion, are particularly large. While wearing this robe, you can use an action to pull off one of the stars and use it to cast \nameref{spell:magic-missile}, creating 8 darts. Daily at dusk, 1d6 removed stars reappear on the robe.  While you wear the robe, you can use an action to enter the Border Shadow along with everything you are wearing and carrying. You remain there until you use an action to return to the plane you were on. You reappear in the last space you occupied, or if that space is occupied, the nearest unoccupied space.

\subsection{Robe of the Archmagi}
Wondrous item, legendary (requires attunement by an arcanist) 

This elegant garment is made from exquisite cloth of white, gray, or black and adorned with silvery runes. You gain these benefits while wearing the robe:
\begin{itemize}
    \item Your Armor Class is 15 + your Dexterity modifier.
    \item You have advantage on saving throws against spells and other magical effects.
    \item Your maximum aether increases by your Intelligence modifier.
\end{itemize}

\subsection{Robe of Useful Items}
Wondrous item, Adventurer 

This robe has cloth patches of various shapes and colors covering it. While wearing the robe, you can use an action to detach one of the patches, causing it to become the object or creature it represents. Once the last patch is removed, the robe becomes an ordinary garment. The robe has two of each of the following patches:
\begin{itemize}
	\item Dagger
	\item Bullseye lantern (filled and lit)
	\item Steel mirror
	\item 10-foot pole
	\item Hempen rope (50 feet, coiled)
	\item Sack
\end{itemize}

In addition, the robe has 4d4 other patches. The GM chooses the patches or determines them randomly.
\begin{DndTable}{lX}
    \textbf{d100} & \textbf{Patch} \\
    01-08 & Bag of 100gp \\
    09-15 & Silver coffer (1 foot long, 6 inches wide and  deep) worth 500 gp \\
    16-22 & Iron door (up to 10 feet wide  and 10 feet high, barred on one side of your choice), which you can place in an opening you can reach; it conforms to fit the opening, attaching and hinging itself \\
    23-30 & 10 gems worth 100 gp each \\
    31-44 & Wooden ladder (24 feet long) \\
    45-51 & A riding horse with saddle bags (friendly toward you) \\
    52-59 & Pit (a cube 10 feet on a side), which you can place on the ground within 10 feet of you \\
    60-68 & 4 potions of healing \\
    69-75 & Rowboat (12 feet long) \\
    76-83 & Spell scroll containing one spell costing 5 or less AET. \\
    84-90 & 2 mastiffs (friendly toward you) \\
    91-96 & Window (2 feet by 4 feet, up to 2 feet deep), which you can place on a vertical surface you can reach \\
    97-00 & Portable ram \\
\end{DndTable}

\subsection{Rune-inscribed Belt}
Belt, Hero (requires attunement)

While wearing this belt, your Constitution score becomes 19 if it was lower. You also have advantage on Charisma (Persuasion) checks made to interact with dwarves, giants, and jazuu if this belt is worn.

\subsection{Shield of Deflection}
Armor (shield), Hero (requires attunement) 

While holding this shield, you have resistance to damage from ranged attacks and magical abilities that affect an area and cause half damage on a successful Dexterity saving throw.

\subsection{Slippers of Spider Climbing}
Boots, Adventurer (requires attunement) 

While you wear these light shoes, you can move up, down, and across vertical surfaces and upside down along ceilings, while leaving your hands free. You have a climbing speed equal to your walking speed. However, the slippers don't allow you to move this way on a slippery surface, such as one covered by ice or oil.

\subsection{Spellguard Shield}
Armor (shield), Hero (requires attunement)

While holding this shield, you have advantage on saving throws against spells and other magical effects, and spell attacks have disadvantage against you.

\subsection{Winged Boots}
Boots, Adventurer (requires attunement) 

These boots have 3 charges. While you wear these boots, you can expend a charge to gain a flying speed equal to your walking speed for 1 hour. If you are still flying when a charge expires, you descend at a rate of 50 ft per round, taking no fall damage. All expended charges are restored at dawn.

\subsection{Wings of Flying}
Cloak, Hero (requires attunement) 

While wearing this cloak, you can use an action to speak its command word. This turns the cloak into a pair of bat wings or bird wings on your back for 1 hour or until you repeat the command word as an action. The wings give you a flying speed of 60 feet. When they disappear, you can't use them again for 1d12 hours.

\subsection{Woodland Cloak}
Cloak, Adventurer (requires attunement) 

While you wear this cloak with its hood up, Wisdom (Perception) checks made to see you have disadvantage, and you have advantage on Dexterity (Stealth) checks made to hide, as the cloak's color shifts to camouflage you. Pulling the hood up or down requires an action.
