\section{Miscellaneous Magic Items}\label{mi:misc}

\subsection{Bag of Holding}
Wondrous item, Adventurer 

This bag has an interior space considerably larger than its outside dimensions, roughly 2 feet in diameter at the mouth and 4 feet deep. The bag can hold up to 500 pounds, not exceeding a volume of 64 cubic feet. The bag weighs 15 pounds, regardless of its contents. Retrieving an item from the bag requires an action.

If the bag is overloaded, pierced, or torn, it ruptures and is destroyed, and its contents fall out. If the bag is turned inside out, its contents spill forth, unharmed, but the bag must be put right before it can be used again. Breathing creatures inside the bag can survive up to a number of minutes equal to 10 divided by the number of creatures (minimum 1 minute), after which time they begin to suffocate.

If you place a bag of holding inside another item that manipulates space or weight, such as a handy haversack or a portable hole, it counts towards the space and weight capacity of that other item as if it was a non-magical bag of its total volume and present contents.

\subsection{Bag of Tricks}
Wondrous item, Adventurer

This ordinary bag, made from gray, rust, or tan cloth, appears empty. Reaching inside the bag, however, reveals the presence of a small, fuzzy object. The bag weighs 1/2 pound.

You can use an action to pull the fuzzy object from the bag and throw it up to 20 feet. When the object
lands, it transforms into a creature you determine by rolling a d8 and consulting the table that corresponds to the bag's color. The creature vanishes at the next dawn or when it is reduced to 0 hit points.

The creature is friendly to you and your companions, and it acts on your turn. You can use a bonus action to command how the creature moves and what action it takes on its next turn, or to give it general orders, such as to attack your enemies. In the absence of such orders, the creature acts in a fashion appropriate to its nature.

Once three fuzzy objects have been pulled from the bag, the bag can't be used again until the next dawn.

\begin{DndTable}[header=Gray Bag of Tricks]{lX}
d8 & Creature \\     
1 &  Weasel    \\    
2 &  Giant rat \\    
3 &  Badger    \\    
4 &  Boar      \\    
5 &  Panther   \\    
6 &  Giant badger \\
7 & Dire wolf \\ 
8 & Giant elk \\
\end{DndTable}  

\begin{DndTable}[header=Rust Bag of Tricks]{lX}
d8 & Creature   \\ 
1  & Rat        \\ 
2  & Owl        \\ 
3  & Mastiff    \\ 
4  & Goat       \\ 
5  & Giant goat \\ 
6  & Giant boar \\ 
7  & Lion       \\ 
8  & Brown bear \\ 
\end{DndTable}

\begin{DndTable}[header=Tan Bag of Tricks]{lX}
d8 & Creature \\      
1  & Jackal   \\     
2  & Ape      \\     
3  & Baboon   \\     
4  & Axe beak \\     
5  & Black bear \\   
6  & Giant weasel \\ 
7  & Giant hyena  \\
8  & Tiger
\end{DndTable}      

\subsection{Bowl of Commanding Water Elementals}
Wondrous item, Hero

While this bowl is filled with water, you can use an action to speak the bowl's command word and summon a water elemental, as if you had cast the
\nameref{spell:conjure-elemental} spell. The bowl can't be used this way again until the next dawn.

The bowl is about 1 foot in diameter and half as deep. It weighs 3 pounds and holds about 3 gallons.

\subsection{Brazier of Commanding Fire Elementals}
Wondrous item, Hero

While a fire burns in this brass brazier, you can use an action to speak the brazier's command word and summon a fire elemental, as if you had cast the \nameref{spell:conjure-elemental} spell. The brazier can't be used this way again until the next dawn.

The brazier weighs 5 pounds.

\subsection{Broom of Flying}
Wondrous item, Hero 

This wooden broom, which weighs 3 pounds, functions like a mundane broom until you stand astride it and speak its command word. It then hovers beneath you and can be ridden in the air. It has a flying speed of 50 feet. It can carry up to 400 pounds, but its flying speed becomes 30 feet while carrying over 200 pounds. The broom stops hovering when you land. You can send the broom to travel alone to a destination within 1 mile of you if you speak the command word, name the location, and are familiar with that place. The broom comes back to you when you speak another command word, provided that the broom is still within 1 mile of you.

\subsection{Carpet of Flying}
Wondrous item, Hero

You can speak the carpet's command word as an action to make the carpet hover and fly. It moves according to your spoken directions, provided that you are within 30 feet of it.

Four sizes of carpet of flying exist. The GM chooses the size of a given carpet or determines it randomly.
\begin{DndTable}{lXll}
d100     &   Size  &         Capacity & Flying Speed  \\  
01-20    &  3 ft. \texttimes 5 ft. & 200 lb. &  80  feet \\      
21-55    &  4 ft. \texttimes 6 ft. & 400 lb. &  60 feet  \\    
56-80    &  5 ft. \texttimes 7 ft. & 600 lb. &  40 feet  \\  
81 - 100 &  6 ft. \texttimes 9 ft. &  800 lb. &  30 feet  \\
\end{DndTable}

A carpet can carry up to twice the weight shown on the table, but it flies at half speed if it carries more than its normal capacity.

\subsection{Censer of Controlling Air Elementals}
Wondrous item, Hero

While incense is burning in this censer, you can use an action to speak the censer's command word and summon an air elemental, as if you had cast the \nameref{spell:conjure-elemental} spell. The censer can't be used this way again until the next dawn.

This 6-inch-wide, 1-foot-high vessel resembles a chalice with a decorated lid. It weighs 1 pound.

\subsection{Crystal Ball}
Wondrous item, Hero or legendary (requires attunement)

The typical crystal ball, a Hero item, is about 6 inches in diameter. While touching it, you can cast the scrying incantation (save DC 17) with it.
The following crystal ball variants are legendary items and have additional properties.

\subparagraph*{Crystal Ball of Mind Reading} You can use an action to cast the
detect thoughts spell (save DC 17) while you are scrying with the crystal ball, targeting creatures you can see within 30 feet of the spell's sensor. You don't need to concentrate on this detect thoughts to maintain it during its duration, but it ends if scrying ends.

\subparagraph*{Crystal Ball of Telepathy} While scrying with the crystal ball, you can communicate telepathically with creatures you can see within 30 feet of the spell's sensor. You can also use an action to cast the suggestion spell (save DC 17) through the sensor on one of those creatures. You don't need to concentrate on this suggestion to maintain it during its duration, but it ends if scrying ends. Once used, the suggestion power of the crystal ball can't be used again until the next dawn.

\subparagraph*{Crystal Ball of True Seeing} While scrying with the crystal ball, you have truesight with a radius of 120 feet centered on the spell's sensor.

\subsection{Cube of Force}
Wondrous item, Hero (requires attunement) 

This cube is about an inch across. Each face has a distinct marking on it that can be pressed. The cube starts with 36 charges, and it regains 1d20 expended charges daily at dawn.  You can use an action to press one of the cube's faces, expending a number of charges based on the chosen face, as shown in the Cube of Force Faces table. Each face has a different effect. If the cube has insufficient charges remaining, nothing happens. Otherwise, a barrier of invisible force springs into existence, forming a cube 15 feet on a side. The barrier is centered on you, moves with you, and lasts for 1 minute, until you use an action to press the cube's sixth face, or the cube runs out of charges. You can change the barrier's effect by pressing a different face of the cube and expending the requisite number of charges, resetting the duration.  If your movement causes the barrier to come into contact with a solid object that can't pass through the cube, you can't move any closer to that object as long as the barrier remains.

\begin{DndTable}[header=Faces]{llX}
Face & Charges & Effect \\                                                                                                               
1  &   1  &      Gases, wind, and  fog can't pass through  the barrier. \\                                                               
2  &   2  &      Nonliving matter can't pass through the  barrier. Walls, floors, and ceilings can pass  through at your discretion. \\ 
3  &   3  &      Living matter can't pass through the  barrier. \\                                                                       
4  &   4  &      Spell effects can't pass through the barrier.  \\                                                                      
5  &   5  &      Nothing can pass through the barrier.  Walls, floors, and ceilings can pass through  at your discretion. \\            
6  &   0  &      The barrier deactivates.                                                                                             
\end{DndTable}

\subsection{Cubic Gate}
Wondrous item, legendary

This cube is 3 inches across and radiates palpable magical energy. The six sides of the cube are each keyed to a different plane of existence, one of which is the Material Plane. The other sides are linked to planes determined by the GM.

You can use an action to press one side of the cube to cast the gate incantation with it, opening a portal to the plane keyed to that side. 

Alternatively, if you use an action to press one side twice, you can cast the plane shift spell (save DC 17) with the cube and transport the targets to the plane keyed to that side.

The cube has 3 charges. Each use of the cube expends 1 charge. The cube regains 1d3 expended charges daily at dawn.

\subsection{Decanter of Endless Water}
Wondrous item, Adventurer 

This stoppered flask sloshes when shaken, as if it contains water. The decanter weighs 2 pounds.  You can use an action to remove the stopper and speak one of three command words, whereupon an amount of fresh water or salt water (your choice) pours out of the flask. The water stops pouring out at the start of your next turn. Choose from the following options:
\begin{itemize}
\item “Stream” produces 1 gallon of water.
\item “Fountain” produces 5 gallons of water.
\item “Geyser” produces 30 gallons of water that gushes forth in a geyser 30 feet long and 1 foot wide. As a bonus action while holding the decanter, you can aim the geyser at a creature you can see within 30feet of you. The target must succeed on a DC 13 Strength saving throw or take 1d4 bludgeoning damage and fall prone. Instead of a creature, you can target an object that isn't being worn or carried and that weighs no more than 200 pounds.The object is either knocked over or pushed up to 15 feet away from you.
\end{itemize}

\subsection{Efficient Quiver}
Wondrous item, Adventurer 

Each of the quiver's three compartments connects to an extradimensional space that allows the quiver to hold numerous items while never weighing more than 2 pounds. The shortest compartment can hold up to sixty arrows, bolts, or similar objects. The midsize compartment holds up to eighteen javelins or similar objects. The longest compartment holds up to six long objects, such as bows, quarterstaffs, or spears.  You can draw any item the quiver contains as if doing so from a regular quiver or scabbard.

\subsection{Eversmoking Bottle}
Wondrous item, Adventurer

Smoke leaks from the lead-stoppered mouth of this brass bottle, which weighs 1 pound. When you use an action to remove the stopper, a cloud of thick smoke pours out in a 60-foot radius from the bottle. The cloud's area is heavily obscured. Each minute the bottle remains open and within the cloud, the radius increases by 10 feet until it reaches its maximum radius of 120 feet.

The cloud persists as long as the bottle is open. Closing the bottle requires you to speak its command word as an action. Once the bottle is closed, the cloud disperses after 10 minutes. A moderate wind (11 to 20 miles per hour) can also disperse the smoke after 1 minute, and a strong wind (21 or more miles per hour) can do so after 1 round.

\subsection{Figurine of Wondrous Power}
Wondrous item, tier varies

A figurine of wondrous power is a statuette of a beast small enough to fit in a pocket. If you use an action to speak the command word and throw the figurine to a point on the ground within 60 feet of you, the figurine becomes a living creature. If the space where the creature would appear is occupied by other creatures or objects, or if there isn't enough space for the creature, the figurine doesn't become a creature.

The creature is friendly to you and your companions. It understands your languages and obeys your spoken commands. If you issue no commands, the creature defends itself but takes no other actions.  The creature exists for a duration specific to each figurine. At the end of the duration, the creature reverts to its figurine form. It reverts to a figurine early if it drops to 0 hit points or if you use an action to speak the command word again while touching it. When the creature becomes a figurine again, its property can't be used again until a certain amount of time has passed, as specified in the figurine's description.

Bronze Griffon (Hero). This bronze statuette is of a griffon rampant. It can become a griffon for up to 6 hours. Once it has been used, it can't be used again until 5 days have passed.

Ebony Fly (Hero). This ebony statuette is carved in the likeness of a horsefly. It can become a giant fly for up to 12 hours and can be ridden as a mount. Once it has been used, it can't be used again until 2 days have passed. The giant fly has the statistics of a riding horse except with a fly speed of 60 feet.

Golden Lions (Hero). These gold statuettes of lions are always created in pairs. You can use one figurine or both simultaneously. Each can become a lion for up to 1 hour. Once a lion has been used, it can't be used again until 7 days have passed.

Ivory Goats (Hero). These ivory statuettes of goats are always created in sets of three. Each goat looks unique and functions differently from the others. Their properties are as follows:
\begin{itemize}
\item The goat of traveling can become a Large goat with the same statistics as a riding horse. It has 24 charges, and each hour or portion thereof it spends in beast form costs 1 charge. While it has charges, you can use it as often as you wish. When it runs out of charges, it reverts to a figurine and can't be used again until 7 days have passed, when it regains all its charges.
\item The goat of travail becomes a giant goat for up to 3hours. Once it has been used, it can't be used again until 30 days have passed.
\item The goat of terror becomes a giant goat for up to 3hours. The goat can't attack, but you can remove its horns and use them as weapons. One horn becomes a +1 lance, and the other becomes a +2 longsword. Removing a horn requires an action, and the weapons disappear and the horns return when the goat reverts to figurine form. In addition,the goat radiates a 30-foot-radius aura of terror while you are riding it. Any creature hostile to you that starts its turn in the aura must succeed on a DC 15 Wisdom saving throw or be frightened of the goat for 1 minute, or until the goat reverts to figurine form. The frightened creature can repeat the saving throw at the end of each of its turns,ending the effect on itself on a success. Once it successfully saves against the effect, a creature is immune to the goat's aura for the next 24 hours. Once the figurine has been used, it can't be used again until 15 days have passed.
\end{itemize}

Marble Elephant (Hero).This marble statuette is about 4 inches high and long. It can become an elephant for up to 24 hours. Once it has been used, it can't be used again until 7 days have passed.

Obsidian Steed (Hero).This polished obsidian horse can become a nightmare for up to 24 hours. The nightmare fights only to defend itself. Once it has been used, it can't be used again until 5 days have passed.

Onyx Dog (Hero).This onyx statuette of a dog can become a mastiff for up to 6 hours. The mastiff has an Intelligence of 8 and can speak Common. It also has darkvision out to a range of 60 feet and can see invisible creatures and objects within that range. Once it has been used, it can't be used again until 7 days have passed.

Serpentine Owl (Hero).This serpentine statuette of an owl can become a giant owl for up to 8 hours. Once it has been used, it can't be used again until 2 days have passed. The owl can telepathically communicate with you at any range if you and it are on the same plane of existence.

Silver Raven (Adventurer).This silver statuette of a raven can become a raven for up to 12 hours. Once
it has been used, it can't be used again until 2 days have passed. While in raven form, the figurine allows you to cast the animal messenger spell on it at will.

\subsection{Folding Boat}
Wondrous item, Hero 

This object appears as a wooden box that measures 12 inches long, 6 inches wide, and 6 inches deep. It weighs 4 pounds and floats. It can be opened to store items inside. This item also has three command words, each requiring you to use an action to speak it. One command word causes the box to unfold into a boat 10 feet long, 4 feet wide, and 2 feet deep. The boat has one pair of oars, an anchor, a mast, and a lateen sail. The boat can hold up to four Medium creatures comfortably.  The second command word causes the box to unfold into a ship 24 feet long, 8 feet wide, and 6 feet deep. The ship has a deck, rowing seats, five sets of oars, a steering oar, an anchor, a deck cabin, and a mast with a square sail. The ship can hold fifteen Medium creatures comfortably.  When the box becomes a vessel, its weight becomes that of a normal vessel its size, and anything that was stored in the box remains in the boat.  The third command word causes the folding boat to fold back into a box, provided that no creatures are aboard. Any objects in the vessel that can't fit inside the box remain outside the box as it folds. Any objects in the vessel that can fit inside the box do so.

\subsection{Gem of Seeing}
Wondrous item, Hero (requires attunement)

This gem has 3 charges. As an action, you can speak the gem's command word and expend 1 charge. For the next 10 minutes, you have truesight out to 120 feet when you peer through the gem.

The gem regains 1d3 expended charges daily at dawn.

\subsection{Handy Haversack}
Wondrous item, Hero

This backpack has a central pouch and two side pouches, each of which is an extradimensional space. Each side pouch can hold up to 20 pounds of material, not exceeding a volume of 2 cubic feet. The large central pouch can hold up to 8 cubic feet or 80 pounds of material. The backpack always weighs 5 pounds, regardless of its contents.

Placing an object in the haversack follows the normal rules for interacting with objects. Retrieving an item from the haversack requires you to use an action. When you reach into the haversack for a specific item, the item is always magically on top.

The haversack has a few limitations. If it is overloaded, or if a sharp object pierces it or tears it, the haversack ruptures and is destroyed. If the haversack is destroyed, its contents spill forth, unharmed. If the haversack is turned inside out, its contents spill forth, unharmed, and the haversack must be put right before it can be used again. If a breathing creature is placed within the haversack, the creature can survive for up to 10 minutes, after which time it begins to suffocate.

If you place a handy haversack inside another item that manipulates space or weight, such as a bag of holding or a portable hole, it counts towards the space and weight capacity of that other item as if it was a non-magical bag of its total volume and present contents.

\subsection{Horn of Blasting}
Wondrous item, Hero 

You can use an action to speak the horn's command word and then blow the horn, which emits a thunderous blast in a 30-foot cone that is audible 600 feet away. Each creature in the cone must make a DC 15 Constitution saving throw. On a failed save, a creature takes 5d6 thunder damage and is deafened for 1 minute. On a successful save, a creature takes half as much damage and isn't deafened. Creatures and objects made of glass or crystal have disadvantage on the saving throw and take 10d6 thunder damage instead of 5d6.  Each use of the horn's magic has a 20 percent chance of causing the horn to explode. The explosion deals 10d6 fire damage to the blower and destroys the horn.

\subsection{Horn of Valhalla}
Wondrous item, Hero (silver or brass), Hero (bronze), or legendary (iron) You can use an action to blow this horn. In response, warrior spirits from the Valhalla appear within 60 feet of you. They use the statistics of a berserker.They return to Valhalla after 1 hour or when they drop to 0 hit points. Once you use the horn, it can't be used again until 7 days have passed.  Four types of horn of Valhalla are known to exist, each made of a different metal. The horn's type determines how many berserkers answer its summons, as well as the requirement for its use. The GM chooses the horn's type or determines it randomly.

\begin{DndTable}{lllX}
d100  & Horn Type & Berserkers Summoned & Requirement \\                           
01-40 & Silver &     2d4 + 2 &              None \\                                   
41-75 & Brass &      3d4 + 3 &              Proficiency with all simple weapons \\   
76-90 & Bronze &    4d4 + 4   &           Proficiency with all medium armor    \\ 
91-00 & Iron   &    5d4 + 5   &            Proficiency with all martial weapons  
\end{DndTable}

If you blow the horn without meeting its requirement, the summoned berserkers attack you. If you meet the requirement, they are friendly to you and your companions and follow your verbal commands.

\subsection{Horseshoes of a Zephyr}
Wondrous item, Hero

These iron horseshoes come in a set of four. While all four shoes are affixed to the hooves of a horse or similar creature, they allow the creature to move normally while floating 4 inches above the ground. This effect means the creature can cross or stand above nonsolid or unstable surfaces, such as water or lava. The creature leaves no tracks and ignores difficult terrain. In addition, the creature can move at normal speed for up to 12 hours a day without suffering exhaustion from a forced march.

\subsection{Horseshoes of Speed}
Wondrous item, Hero

These iron horseshoes come in a set of four. While all four shoes are affixed to the hooves of a horse or similar creature, they increase the creature's walking speed by 30 feet.

\subsection{Immovable Rod}
Rod, Adventurer

This flat iron rod has a button on one end. You can use an action to press the button, which causes the rod to become magically fixed in place relative to the surface of the planet or a vehicle of Huge size or larger that you are riding. Until you or another creature uses an action to push the button again, the rod doesn't move, even if it is defying gravity. The rod can hold up to 8,000 pounds of weight. More weight causes the rod to deactivate and fall. A creature can use an action to make a DC 30 Strength check, moving the fixed rod up to 10 feet on a success.

\subsection{Instant Fortress}
Wondrous item, Hero

You can use an action to place this 1-inch metal cube on the ground and speak its command word. The cube rapidly grows into a fortress that remains until you use an action to speak the command word that dismisses it, which works only if the fortress is empty.

The fortress is a square tower, 20 feet on a side and 30 feet high, with arrow slits on all sides and a battlement atop it. Its interior is divided into two floors, with a ladder running along one wall to connect them. The ladder ends at a trapdoor leading to the roof. When activated, the tower has a small door on the side facing you. The door opens only at your command, which you can speak as a bonus action. It is immune to the
knockspell and similar magic, such as that of a chime of opening.

Each creature in the area where the fortress appears must make a DC 15 Dexterity saving throw, taking 10d10 bludgeoning damage on a failed save, or half as much damage on a successful one. In either case, the creature is pushed to an unoccupied space outside but next to the fortress. Objects in the area that aren't being worn or carried take this damage and are pushed automatically.  The tower is made of adamantine, and its magic prevents it from being tipped over. The roof, the door, and the walls each have 100 hit points, immunity to damage from nonmagical weapons excluding siege weapons, and resistance to all other damage. Damage done to the tower cannot be repaired, and if any piece is destroyed, the fortress immediately returns to its cube form and no longer functions.

\subsection{Iron Bands of Binding}
Wondrous item, Hero 

This rusty iron sphere measures 3 inches in diameter and weighs 1 pound. You can use an action to speak the command word and throw the sphere at a Huge or smaller creature you can see within 60 feet of you. As the sphere moves through the air, it opens into a tangle of metal bands.  Make a ranged attack roll with an attack bonus equal to your Dexterity modifier plus your proficiency bonus. On a hit, the target is restrained until you take a bonus action to speak the command word again to release it. Doing so, or missing with the attack, causes the bands to contract and become a sphere once more.  A creature, including the one restrained, can use an action to make a DC 20 Strength check to break the iron bands. On a success, the item is destroyed, and the restrained creature is freed. If the check fails, any further attempts made by that creature automatically fail until 24 hours have elapsed.  Once the bands are used, they can't be used again until the next dawn.

\subsection{Iron Flask}
Wondrous item, legendary 

This iron bottle has a brass stopper. You can use an action to speak the flask's command word, targeting a creature that you can see within 60 feet of you. If the target is native to a plane of existence other than the one you're on, the target must succeed on a DC 17 Wisdom saving throw or be trapped in the flask. If the target has been trapped by the flask before, it has advantage on the saving throw. Once trapped, a creature remains in the flask until released. The flask can hold only one creature at a time. A creature trapped in the flask doesn't need to breathe, eat, or drink and doesn't age.  You can use an action to remove the flask's stopper and release the creature the flask contains. The creature is friendly to you and your companions for 1 hour and obeys your commands for that duration. If you give no commands or give it a command that is likely to result in its death, it defends itself but otherwise takes no actions. At the end of the duration, the creature acts in accordance with its normal disposition. An identify spell reveals that a creature is inside the flask, but the only way to determine the type of creature is to open the flask. A newly discovered bottle might already contain a creature chosen by the GM.

\subsection{Lantern of Revealing}
Wondrous item, Adventurer

While lit, this hooded lantern burns for 6 hours on 1 pint of oil, shedding bright light in a 30-foot radius and dim light for an additional 30 feet. Invisible creatures and objects are visible as long as they are in the lantern's bright light. You can use an action to lower the hood, reducing the light to dim light in a 5-foot radius.

\subsection{Manual of Golems}
Wondrous item, Hero

This tome contains information and incantations necessary to make a particular type of golem. The GM chooses the type or determines it randomly. To decipher and use the manual, you must be at least 12th level and have an Intelligence modifier of +3 or higher. A creature that can't use a
manual of golems and attempts to read it takes 6d6 psychic damage.

\begin{DndTable}{lllX}
d20    &  Golem & Time     &  Cost \\        
1 - 5  &  Clay  & 30 days  &  65,000 gp \\   
6 - 17 &  Flesh & 60 days  &  50,000 gp \\  
18     &  Iron  & 120 days &  100,000 gp \\ 
19 - 20 & Stone & 90 days  &  80,000 gp \\  
\end{DndTable}

To create a golem, you must spend the time shown on the table, working without interruption with the manual at hand and resting no more than 8 hours per day. You must also pay the specified cost to purchase supplies.  Once you finish creating the golem, the book is consumed in eldritch flames. The golem becomes animate when the ashes of the manual are sprinkled on it. It is under your control, and it understands and obeys your spoken commands.

\subsection{Mirror of Life Trapping}
Wondrous item, Hero

When this 4-foot-tall mirror is viewed indirectly, its surface shows faint images of creatures. The mirror weighs 50 pounds, and it has AC 11, 10 hit points, and vulnerability to bludgeoning damage. It shatters and is destroyed when reduced to 0 hit points. If the mirror is hanging on a vertical surface and you are within 5 feet of it, you can use an action to speak its command word and activate it. It remains activated until you use an action to speak the command word again.

Any creature other than you that sees its reflection in the activated mirror while within 30 feet of it must succeed on a DC 15 Charisma saving throw or be trapped, along with anything it is wearing or carrying, in one of the mirror's twelve extradimensional cells. This saving throw is made with advantage if the creature knows the mirror's nature, and constructs succeed on the saving throw automatically.

An extradimensional cell is an infinite expanse filled with thick fog that reduces visibility to 10 feet. Creatures trapped in the mirror's cells don't age, and they don't need to eat, drink, or sleep. A creature trapped within a cell can escape using magic that permits planar travel. Otherwise, the creature is confined to the cell until freed.

If the mirror traps a creature but its twelve extradimensional cells are already occupied, the mirror frees one trapped creature at random to accommodate the new prisoner. A freed creature appears in an unoccupied space within sight of the mirror but facing away from it. If the mirror is shattered, all creatures it contains are freed and appear in unoccupied spaces near it.

While within 5 feet of the mirror, you can use an action to speak the name of one creature trapped in it or call out a particular cell by number. The creature named or contained in the named cell appears as an image on the mirror's surface. You and the creature can then communicate normally.
In a similar way, you can use an action to speak a second command word and free one creature trapped in the mirror. The freed creature appears, along with its possessions, in the unoccupied space nearest to the mirror and facing away from it.

\subsection{Pipes of Haunting}
Wondrous item, Adventurer

You must be proficient with wind instruments to use these pipes. They have 3 charges. You can use an action to play them and expend 1 charge to create an eerie, spellbinding tune. Each creature within 30 feet of you that hears you play must succeed on a DC 15 Wisdom saving throw or become frightened of you for 1 minute. If you wish, all creatures in the area that aren't hostile toward you automatically succeed on the saving throw. A creature that fails the saving throw can repeat it at the end of each of its turns, ending the effect on itself on a success. A creature that succeeds on its saving throw is immune to the effect of these pipes for 24 hours. The pipes regain 1d3 expended charges daily at dawn.

\subsection{Pipes of the Sewers}
Wondrous item, Adventurer (requires attunement)

You must be proficient with wind instruments to use these pipes. While you are attuned to the pipes, ordinary rats and giant rats are indifferent toward you and will not attack you unless you threaten or harm them.

The pipes have 3 charges. If you play the pipes as an action, you can use a bonus action to expend 1 to 3 charges, calling forth one swarm of rats with each expended charge, provided that enough rats are within half a mile of you to be called in this fashion (as determined by the GM). If there aren't enough rats to form a swarm, the charge is wasted. Called swarms move toward the music by the shortest available route but aren't under your control otherwise. The pipes regain 1d3 expended charges daily at dawn.  Whenever a swarm of rats that isn't under another creature's control comes within 30 feet of you while you are playing the pipes, you can make a Charisma check contested by the swarm's Wisdom check. If you lose the contest, the swarm behaves as it normally would and can't be swayed by the pipes' music for the next 24 hours. If you win the contest, the swarm is swayed by the pipes' music and becomes friendly to you and your companions for as long as you continue to play the pipes each round as an action. A friendly swarm obeys your commands. If you issue no commands to a friendly swarm, it defends itself but otherwise takes no actions. If a friendly swarm starts its turn and can't hear the pipes' music, your control over that swarm ends, and the swarm behaves as it normally would and can't be swayed by the pipes' music for the next 24 hours.

\subsection{Portable Hole}
Wondrous item, Hero 

This fine black cloth, soft as silk, is folded up to the dimensions of a handkerchief. It unfolds into a circular sheet 6 feet in diameter.  You can use an action to unfold a portable hole and place it on or against a solid surface, whereupon the portable hole creates an extradimensional hole 10 feet deep. The cylindrical space within the hole exists on a different plane, so it can't be used to create open passages. Any creature inside an open portable hole can exit the hole by climbing out of it.  You can use an action to close a portable hole by taking hold of the edges of the cloth and folding it up. Folding the cloth closes the hole, and any creatures or objects within remain in the extradimensional space. No matter what's in it, the hole weighs next to nothing.  If the hole is folded up, a creature within the hole's extradimensional space can use an action to make a DC 10 Strength check. On a successful check, the creature forces its way out and appears within 5 feet of the
portable hole or the creature carrying it. A breathing creature within a closed portable hole can survive for up to 10 minutes, after which time it begins to suffocate.

If you place a portable hole inside another item that manipulates space or weight, such as a handy haversack or a bag of holding, it counts towards the space and weight capacity of that other item as if it was a non-magical bag of its total volume and present contents.

\subsection{Rope of Climbing}
Wondrous item, Adventurer 

This 60-foot length of silk rope weighs 3 pounds and can hold up to 3,000 pounds. If you hold one end of the rope and use an action to speak the command word, the rope animates. As a bonus action, you can command the other end to move toward a destination you choose. That end moves 10 feet on your turn when you first command it and 10 feet on each of your turns until reaching its destination, up to its maximum length away, or until you tell it to stop. You can also tell the rope to fasten itself securely to an object or to unfasten itself, to knot or unknot itself, or to coil itself for carrying.  If you tell the rope to knot, large knots appear at 1-foot intervals along the rope. While knotted, the rope shortens to a 50-foot length and grants advantage on checks made to climb it.

The rope has AC 20 and 20 hit points. It regains 1 hit point every 5 minutes as long as it has at least 1 hit point. If the rope drops to 0 hit points, it is destroyed.

\subsection{Rope of Entanglement}
Wondrous item, Hero

This rope is 30 feet long and weighs 3 pounds. If you hold one end of the rope and use an action to speak its command word, the other end darts forward to entangle a creature you can see within 20 feet of you. The target must succeed on a DC 15 Dexterity saving
throw or become restrained.

You can release the creature by using a bonus action to speak a second command word. A target restrained by the rope can use an action to make a DC 15 Strength or Dexterity check (target's choice). On a success, the creature is no longer restrained by the rope.

The rope has AC 20 and 20 hit points. It regains 1 hit point every 5 minutes as long as it has at least 1 hit point. If the rope drops to 0 hit points, it is destroyed. 

\subsection{Stone of Controlling Earth Elementals}
Wondrous item, Hero 

If the stone is touching the ground, you can use an action to speak its command word and summon an earth elemental, as if you had cast the conjure elemental spell. The stone can't be used this way again until the next dawn. The stone weighs 5 pounds.

\subsection{Well of Many Worlds}
Wondrous item, legendary

This fine black cloth, soft as silk, is folded up to the dimensions of a handkerchief. It unfolds into a circular sheet 6 feet in diameter.
You can use an action to unfold and place the well of many worlds on a solid surface, whereupon it creates a two-way portal to another world or plane of existence. Each time the item opens a portal, the GM decides where it leads. You can use an action to close an open portal by taking hold of the edges of the cloth and folding it up. Once well of many worlds has opened a portal, it can't do so again for 1d8 hours.

\subsection{Wind Fan}
Wondrous item, Adventurer

While holding this fan, you can use an action to cast the gust of windspell (save DC 13) from it. Once used, the fan shouldn't be used again until the next dawn. Each time it is used again before then, it has a cumulative 20 percent chance of not working and tearing into useless, nonmagical tatters.