\section{Rings and Amulets}\label{mi:rings}

This section contains items worn as jewelry. Rings, amulets, broaches, earrings, bracers, etc. Almost all of these require attunement, which serves as the limiting factor on how many you can gain benefit from at a time.

\subsection{Amulet of Health}
\textit{Wondrous item, Hero (requires attunement)}

Your Constitution score is +4 while you wear this amulet. It has no effect on you if your Constitution is already +4 or higher.

\subsection{Amulet of Nondetection}
\textit{Wondrous item, Adventurer (requires attunement)}

While wearing this amulet, you are hidden from divination magic as if under the effects of the \nameref{spell:nondetection}. You can't be targeted by such magic or perceived through magical scrying sensors.

\subsection{Amulet of the Planes}
\textit{Wondrous item, Hero (requires attunement)}

While wearing this amulet, you can use an action to name a location that you are familiar with on another plane of existence. Then make a DC 15 Intelligence check. On a successful check, you cast the plane shift incantation. On a failure, you and each creature and object within 15 feet of you travel to a random destination. Roll a d100. On a 1-60, you travel to a random location on the plane you named. On a 61-100, you travel to a randomly determined plane of existence.

\subsection{Bracers of Archery}
\textit{Wondrous item, Adventurer (requires attunement)}

While wearing these bracers, you have proficiency with the longbow and shortbow, and you gain a +2 bonus to damage rolls on ranged attacks made with such weapons.

\subsection{Bracers of Defense}
\textit{Wondrous item, Hero (requires attunement)}

While attuned to these bracers, you gain the benefits of the \nameref{spell:mage-armor} spell.

\subsection{Brooch of Shielding}
\textit{Wondrous item, Adventurer (requires attunement)}

While wearing this brooch, you have resistance to bludgeoning damage, and you have immunity to damage from the \nameref{spell:magic-missile} spell.

\subsection{Dimensional Shackles}
\textit{Wondrous item, Hero}

You can use an action to place these shackles on an incapacitated creature. The shackles adjust to fit a creature of Small to Large size. In addition to serving as mundane manacles, the shackles prevent a creature bound by them from using any method of extradimensional movement, including teleportation or travel to a different plane of existence. They don't prevent the creature from passing through an interdimensional portal. You and any creature you designate when you use the shackles can use an action to remove them. Once every 30 days, the bound creature can make a DC 30 Strength (Athletics) check. On a success, the creature breaks free and destroys the shackles.

\subsection{Eyes of Charming}
\textit{Wondrous item, Adventurer (requires attunement)}

These crystal lenses fit over the eyes. They have 3 charges. While wearing them, you can expend 1 charge as an action to cast the charm person spell (save DC 13) on a humanoid within 30 feet of you, provided that you and the target can see each other. The lenses regain all expended charges daily at dawn.

\subsection{Eyes of Minute Seeing}
\textit{Wondrous item, Adventurer}

These crystal lenses fit over the eyes. While wearing them, you can see much better than normal out to a range of 1 foot. You have advantage on Intelligence (Investigation) checks that rely on sight while searching an area or studying an object within that range.

\subsection{Eyes of the Eagle}
\textit{Wondrous item, Adventurer (requires attunement)}

These crystal lenses fit over the eyes. While wearing them, you have advantage on Wisdom (Perception) checks that rely on sight. In conditions of clear visibility, you can make out details of even extremely distant creatures and objects as small as 2 feet across.

\subsection{Goggles of Night}
\textit{Wondrous item, Adventurer} 

While wearing these dark lenses, you have darkvision out to a range of 60 feet. If you already have darkvision, wearing the goggles increases its range by 60 feet.

\subsection{Medallion of Thoughts}
\textit{Wondrous item, Adventurer (requires attunement)}

The medallion has 3 charges. While wearing it, you can use an action and expend 1 charge to cast the \nameref{spell:detect-thoughts} spell (save DC 13) from it. The medallion regains 1d3 expended charges daily at dawn.

\subsection{Necklace of Adaptation}
\textit{Wondrous item, Adventurer (requires attunement)}

While wearing this necklace, you can breathe normally in any environment, and you have advantage on saving throws made against harmful gases and vapors (such as cloudkill and stinking cloud effects, inhaled poisons, and the breath weapons of some dragons).

\subsection{Pearl of Power}
\textit{Wondrous item, Adventurer (requires attunement by a spellcaster)}

While this pearl is on your person, you can use an action to speak its command word, instantly regaining aether (AET) equal to your proficiency bonus up to your maximum. Once you use the pearl, it can`t be used again until the next dawn.

\subsection{Periapt of Health}
\textit{Wondrous item, Adventurer}

You are immune to contracting any disease while you wear this pendant. If you are already infected with a disease, the effects of the disease are suppressed and you count as succeeding on any requisite saving throws you while you wear the pendant.

\subsection{Periapt of Proof against Poison}
\textit{Wondrous item, Hero}

This delicate silver chain has a brilliant-cut black gem pendant. While you wear it, poisons have no effect on you. You are immune to the poisoned condition and have immunity to poison damage.

\subsection{Periapt of Wound Closure}
\textit{Wondrous item, Adventurer (requires attunement)}

While you wear this pendant, you stabilize whenever you are dying at the start of your turn. In addition, whenever you roll a Hit Die to regain hit points, double the number of hit points it restores.

\subsection{Ring of Animal Influence}
\textit{Ring, Hero} 

This ring has 3 charges, and it regains 1d3 expended charges daily at dawn. While wearing the ring, you can use an action to expend 1 of its charges to cast one of the following spells:
\begin{itemize}
 \item Animal friendship (save DC 13)
 \item Fear (save DC 13), targeting only beasts that have an Intelligence of 3 or lower
 \item Speak with animals
\end{itemize}

\subsection{Ring of Djinni Summoning}
\textit{Ring, legendary (requires attunement)}

While wearing this ring, you can speak its command word as an action to summon a particular djinni from one of the Elemental Planes of Air. The djinni appears in an unoccupied space you choose within 120 feet of you. It remains as long as you concentrate (as if concentrating on a spell), to a maximum of 1 hour, or until it drops to 0 hit points. It then returns to its home plane.

While summoned, the djinni is friendly to you and your companions. It obeys any commands you give it, no matter what language you use. If you fail to command it, the djinni defends itself against attackers but takes no other actions.

After the djinni departs, it can't be summoned again for 24 hours, and the ring becomes nonmagical if the djinni dies.

\subsection{Ring of Elemental Command}
\textit{Ring, legendary (requires attunement)}

This ring is linked to one of the four Prime Elemental Planes. The GM chooses or randomly determines the linked plane.

While wearing this ring, you have advantage on attack rolls against elementals from the linked plane, and they have disadvantage on attack rolls against you. In addition, you have access to properties based on the linked plane.

The ring has 5 charges. It regains 1d4 +1 expended charges daily at dawn. Spells cast from the ring have a save DC of 17.

\subparagraph*{Ring of Air Elemental Command} You can expend 2 of the ring's charges to cast dominate monster on an air elemental. In addition, when you fall, you descend 60 feet per round and take no damage from falling. You can also speak and understand Auran. If you help slay an air elemental while attuned to the ring, you gain access to the following additional properties:
\begin{itemize}
\item You have resistance to lightning damage.
\item You have a flying speed equal to your walking speed and can hover.
\item You can cast the following spells from the ring,expending the necessary number of charges: chain lightning (3 charges), gust of wind (2 charges), or wind wall (1 charge).
\end{itemize}

\subparagraph*{Ring of Earth Elemental Command} You can expend 2 of the ring's charges to cast dominate monster on an earth elemental. In addition, you can move in difficult terrain that is composed of rubble, rocks, or dirt as if it were normal terrain. You can also speak and understand Terran.  If you help slay an earth elemental while attuned to the ring, you gain access to the following additional properties:
\begin{itemize}
\item You have resistance to acid damage.
\item You can move through solid earth or rock as if those areas were difficult terrain. If you end your turn there, you are shunted out to the nearest unoccupied space you last occupied.
\item You can cast the following spells from the ring,expending the necessary number of charges: stone shape (2 charges), stoneskin (3 charges), or wall of stone (3 charges).
\end{itemize}

\subparagraph*{Ring of Fire Elemental Command}
You can expend 2 of the ring's charges to cast dominate monster on a fire elemental. In addition, you have resistance to fire damage. You can also speak and understand Ignan. If you help slay a fire elemental while attuned to the ring, you gain access to the following additional properties:
\begin{itemize}
\item You are immune to fire damage.
\item You can cast the following spells from the ring,expending the necessary number of charges:burning hands (1 charge), fireball (2 charges), and wall of fire (3 charges).
\end{itemize}

\subparagraph*{Ring of Water Elemental Command}
You can expend 2 of the ring's charges to cast dominate monster on a water elemental. In addition, you can stand on and walk across liquid surfaces as if they were solid ground. You can also speak and understand Aquan.  If you help slay a water elemental while attuned to the ring, you gain access to the following additional properties:
\begin{itemize}
\item You can breathe underwater and have a swimming speed equal to your walking speed.
\item You can cast the following spells from the ring, expending the necessary number of charges: create or destroy water (1 charge), control water (3 charges), ice storm (2 charges), or wall of ice (3 charges).
\end{itemize}

\subsection{Ring of Evasion}
\textit{Ring, Hero (requires attunement)} 

This ring has 3 charges, and it regains 1d3 expended charges daily at dawn. When you fail a Dexterity saving throw while wearing it, you can use your reaction to expend 1 of its charges to succeed on that saving throw instead.

\subsection{Ring of Feather Falling}
\textit{Ring, Hero (requires attunement)}

When you fall while wearing this ring, you descend 60 feet per round and take no damage from falling.

\subsection{Ring of Free Action}
\textit{Ring, Hero (requires attunement)}

While you wear this ring, difficult terrain doesn't cost you extra movement. In addition, magic can neither reduce your speed nor cause you to be paralyzed or restrained.

\subsection{Ring of Invisibility}
\textit{Ring, legendary (requires attunement)}

While wearing this ring, you can turn invisible as an action. Anything you are wearing or carrying is invisible with you. You remain invisible until the ring is removed, until you attack or cast a spell, or until you use a bonus action to become visible again.

\subsection{Ring of Jumping}
\textit{Ring, Adventurer (requires attunement)}

While wearing this ring, you can cast the jump spell from it as a bonus action at will, but can target only yourself when you do so.

\subsection{Ring of Mind Shielding}
\textit{Ring, Adventurer (requires attunement)}

While wearing this ring, you are immune to magic that allows other creatures to read your thoughts, determine whether you are lying, know your alignment, or know your creature type. Creatures can telepathically communicate with you only if you allow it.

You can use an action to cause the ring to become invisible until you use another action to make it visible, until you remove the ring, or until you die.

\subsection{Ring of Protection}
\textit{Ring, Hero (requires attunement)}

This ring has 3 charges. You can expend a charge to take the Deflect, Exert, or Focus actions without expending stamina or aether. The ring regains all expended charges at dawn.

\subsection{Ring of Regeneration}
\textit{Ring, Hero (requires attunement)}

While wearing this ring, you regain 1d6 hit points every 10 minutes, provided that you have at least 1 hit point. If you lose a body part, the ring causes the missing part to regrow and return to full functionality after 1d6 + 1 days if you have at least 1 hit point the whole time.

\subsection{Ring of Energy Resistance}
\textit{Ring, Hero (requires attunement)} 

You have resistance to one damage type while wearing this ring. The gem in the ring indicates the type, which the GM chooses or determines randomly.
\begin{DndTable}{lXX}
d10 & Damage Type & Gem \\         
1  &  Acid  &       Pearl \\      
2  &  Cold  &       Tourmaline \\  
3  &  Fire   &      Garnet      \\
4  &  Lightning &   Citrine    \\ 
5  &  Necrotic  &   Jet         \\   
7  &  Psychic   &   Jade        \\
8  &  Radiant   &   Topaz       \\
9  & Thunder    &  Spinel      \\
10 & Reroll & --- \\
\end{DndTable}

\subsection{Ring of Shooting Stars}
\textit{Ring, Hero (requires attunement outdoors at night)}

While wearing this ring in dim light or darkness, you can cast dancing lights and light from the ring at will. Casting either spell from the ring requires an action. The ring has 6 charges for the following other properties. The ring regains 1d6 expended charges daily at dawn.

\subparagraph*{Faerie Fire} You can expend 1 charge as an action to cast faerie fire from the ring.

\subparagraph*{Ball Lightning} You can expend 2 charges as an action to create one to four 3-foot-diameter spheres of lightning. The more spheres you create, the less powerful each sphere is individually.  Each sphere appears in an unoccupied space you can see within 120 feet of you. The spheres last as long as you concentrate (as if concentrating on a spell), up to 1 minute. Each sphere sheds dim light in a 30-foot radius.

As a bonus action, you can move each sphere up to 30 feet, but no farther than 120 feet away from you. When a creature other than you comes within 5 feet of a sphere, the sphere discharges lightning at that creature and disappears. That creature must make a DC 15 Dexterity saving throw. On a failed save, the creature takes lightning damage based on the number of spheres you created.
\begin{DndTable}{ll}
Spheres & Lightning Damage \\  
4   &     2d4 \\              
3   &     2d6 \\              
2   &     5d4 \\              
1   &     4d12 \\
\end{DndTable}             

\subparagraph*{Shooting Stars} You can expend 1 to 3 charges as an action. For every charge you expend, you launch a glowing mote of light from the ring at a point you can see within 60 feet of you. Each creature within a 15-foot cube originating from that point is showered in sparks and must make a DC 15 Dexterity saving throw, taking 5d4 fire damage on a failed save, or half as much damage on a successful one.

\subsection{Ring of Spell Storing}
\textit{Ring, Hero (requires attunement)}

This ring stores spells cast into it, holding them until the attuned wearer uses them. The ring can store up to 12 aether worth of spells at a time. When found, it contains 6 aether of stored spells chosen by the GM.

Any creature can cast a spell costing 12 or less aether into the ring by touching the ring as the spell is cast. The spell has no effect, other than to be stored in the ring. If the ring can't hold the spell, the spell is expended without effect. The base aether cost of the spell determines how much space it uses.

While wearing this ring, you can cast any spell stored in it. The spell uses the aether cost, spell save DC, spell attack bonus, and spellcasting ability of the original caster, but is otherwise treated as if you cast the spell. The spell cast from the ring is no longer stored in it, freeing up space.

\subsection{Ring of Spell Turning}
\textit{Ring, legendary (requires attunement)}

While wearing this ring, you have advantage on saving throws against any spell that targets only you (not in an area of effect). In addition, if you roll a 20 for the save and the spell is 7th level or lower, the spell has no effect on you and instead targets the caster, using the slot level, spell save DC, attack bonus, and spellcasting ability of the caster.

\subsection{Ring of Swimming}
\textit{Ring, Adventurer} 

You have a swimming speed of 40 feet while wearing this ring.

\subsection{Ring of Telekinesis}
\textit{Ring, Hero (requires attunement)}

While wearing this ring, you can cast the telekinesis spell at will, but you can target only objects that aren't being worn or carried.

\subsection{Ring of the Ram}
\textit{Ring, Hero (requires attunement)}

This ring has 3 charges, and it regains 1d3 expended charges daily at dawn. While wearing the ring, you can use an action to expend 1 to 3 of its charges to attack one creature you can see within 60 feet of you. The ring produces a spectral ram's head and makes its attack roll with a +7 bonus. On a hit, for each charge you spend, the target takes 2d10 bludgeoning damage and is pushed 5 feet away from you.  Alternatively, you can expend 1 to 3 of the ring's charges as an action to try to break an object you can see within 60 feet of you that isn't being worn or carried. The ring makes a Strength check with a +5 bonus for each charge you spend.

\subsection{Ring of Three Wishes}
\textit{Ring, legendary}

While wearing this ring, you can use an action to expend 1 of its 3 charges to make a wish. The ring becomes nonmagical when you use the last charge.

Wishes must be clearly stated in a few simple sentences. Attempts to cleverly word wishes to evade the restrictions described here result in the GM warning you or refusing that wish (without expending a wish). No wish can reverse a past event (although it might be able to undo its effects, such as by raising a person who was killed), can make someone fall in love with another person, or grant more wishes. Where possible, the GM will answer the wish by replicating the effect of a spell or legendary effect.

\subsection{Ring of Warmth}
\textit{Ring, Adventurer (requires attunement)}

While wearing this ring, you have resistance to cold damage. In addition, you and everything you wear and carry are unharmed by temperatures as low as \textminus50 degrees Fahrenheit.

\subsection{Ring of Water Walking}
\textit{Ring, Adventurer}

While wearing this ring, you can stand on and move across any liquid surface as if it were solid ground.

\subsection{Ring of X-ray Vision}
\textit{Ring, Hero (requires attunement)}

While wearing this ring, you can use an action to speak its command word. When you do so, you can see into and through solid matter for 1 minute. This vision has a radius of 30 feet. To you, solid objects within that radius appear transparent and don't prevent light from passing through them. The vision can penetrate 1 foot of stone, 1 inch of common metal, or up to 3 feet of wood or dirt. Thicker substances block the vision, as does a thin sheet of lead.

Whenever you use the ring again before taking a long rest, you must succeed on a DC 15 Constitution saving throw or gain one level of exhaustion.

\subsection{Scarab of Protection}
\textit{Wondrous item, legendary (requires attunement)}

If you hold this beetle-shaped medallion in your hand for 1 round, an inscription appears on its surface revealing its magical nature. It provides two benefits while it is on your person:
\begin{itemize}
\item You have advantage on saving throws against spells.
\item The scarab has 12 charges. If you fail a saving throw against a spell that deals necrotic damage or a harmful effect originating from an undead creature, you can use your reaction to expend 1 charge and turn the failed save into a successful one. The scarab crumbles into powder and is destroyed when its last charge is expended.
\end{itemize}

\subsection{Stone of Good Luck (Luckstone)}
\textit{Wondrous item, Adventurer (requires attunement)} 

While this polished agate is on your person, you gain a +1 bonus to ability checks and saving throws.