

Apparatus of the Crab
Wondrous item, legendary This item first appears to be a Large sealed iron barrel weighing 500 pounds. The barrel has a hidden catch, which can be found with a successful DC 20 Intelligence (Investigation) check. Releasing the catch unlocks a hatch at one end of the barrel, allowing two Medium or smaller creatures to crawl inside. Ten levers are set in a row at the far end, each in a neutral position, able to move either up or down. When certain levers are used, the apparatus transforms to resemble a giant lobster.  The apparatus of the Crab is a Large object with the following statistics:
Armor Class: 20
Hit Points: 200
Speed: 30 ft., swim 30 ft. (or 0 ft. for both if the legs and tail aren't extended)
Damage Immunities: poison, psychic To be used as a vehicle, the apparatus requires one pilot. While the apparatus's hatch is closed, the compartment is airtight and watertight. The compartment holds enough air for 10 hours of breathing, divided by the number of breathing creatures inside.  The apparatus floats on water. It can also go underwater to a depth of 900 feet. Below that, the vehicle takes 2d6 bludgeoning damage per minute from pressure.  A creature in the compartment can use an action to move as many as two of the apparatus's levers up or down. After each use, a lever goes back to its neutral position. Each lever, from left to right, functions as shown in the Apparatus of the Crab Levers table.

Apparatus of the Crab Levers
Lever                                      Up                                                                                                                                    Down                                                                                                                                                
1                                          Legs and tail extend,                                                                                                                 Legs and tail retract,                                                                                                                              
allowing the apparatus  to walk and swim.  reducing the apparatus's  speed to 0 and making it  unable to benefit from  bonuses to speed.                                         
2                                          Forward window shutter  opens.                                                                                                        Forward window shutter  closes.                                                                                                                     
3                                          Side window shutters  open (two per side).                                                                                            Side window shutters  close (two per side).                                                                                                         
4                                          Two claws  extend from  the front sides of the  apparatus.                                                                            The claws retract.                                                                                                                                  
5                                          Each extended claw  makes the following  melee weapon attack:  +8 to hit, reach 5 ft., one target. Hit: 7 (2d6 ) bludgeoning damage.  Each extended claw  makes the following  melee weapon attack:   +8  to hit, reach 5 ft., one  target. Hit: The target is  grappled (escape DC 15).  
6                                          The apparatus walks or  swims forward.                                                                                                The apparatus walks or  swims backward.                                                                                                             
7                                          The apparatus turns 90  degrees left.                                                                                                 The apparatus turns 90  degrees right.                                                                                                              
8                                          Eyelike fix tur es emit  bright light in a 30 --foot  radius and dim light for  an additional 30 feet.                                The light turns off.                                                                                                                                
9                                          The apparatus sinks as  much as 20 feet in  liquid.                                                                                   The apparatus rises up  to 20 feet in liquid.                                                                                                       
10                                         The rear hatch unseals  and opens.                                                                                                    The rear hatch  closes  and seals.                                                                                                                  






Bag of Beans
Wondrous item, Hero
Inside this heavy cloth bag are 3d4 dry beans. The bag weighs 1/2 pound plus 1/4 pound for each bean it contains.
If you dump the bag's contents out on the ground, they explode in a 10-foot radius, extending from the beans. Each creature in the area, including you, must make a DC 15 Dexterity saving
throw, taking 5d4 fire damage on a failed save, or half as much damage on a successful one. The fire ignites flammable objects in the area that aren't being worn or carried.
If you remove a bean from the bag, plant it in dirt or sand, and then water it, the bean produces an effect 1 minute later from the ground where it was planted. The GM can choose an effect from the following table, determine it randomly, or create an effect.
d100      Effect                                                                                                                                                                                                                                                                                                                                                                         
01        5d4 toadstools sprout. If a creature eats a  toadstool, roll any die. On an odd roll, the eater  must succeed on a DC 15 Constitution saving  throw or take 5d6 poison damage and become  poisoned for 1 hour. On an even roll, the eater  gains 5d6 temporary hit po ints for 1 hour.                                                                                         
02 - 10   A geyser erupts and spouts water, beer, berry  juice, tea, vinegar, wine, or oil (GM's choice) 30  feet into the air for 1d12 rounds.                                                                                                                                                                                                                                          
11 - 20   A treant sprouts. There's a 50 percent chance  that the treant is chaotic evil and attacks.                                                                                                                                                                                                                                                                                    
2 1 - 30  An animate, immobile stone statue in your  likeness rises. It makes verbal threats against  you. If you leave it and others come near, it  describes you as the most heinous of villains and  directs the newcomers to find and attack you. If  you are on the s ame plane of existence as the  statue, it knows where you are. The statue  becomes inanimate after 24 hours.  
31 - 40   A campfire with blue flames springs forth and  burns for 24 hours (or until it is extinguished).                                                                                                                                                                                                                                                                               
41 - 50   1d6 + 6 shriekers sprout                                                                                                                                                                                                                                                                                                                                                       
51 - 60   1d4 + 8   bright pink toads crawl forth. Whenever  a toad is touched, it transforms into a Large or  smaller monster of the GM's choice. The  monster remains for 1 minute, then disappears  in a puff of bright pink smoke.                                                                                                                                                   
61 - 70   A hungry bulette burrows up and attacks.                                                                                                                                                                                                                                                                                                                                       
71 - 80   A fruit tree grows. It has 1d10 + 20 fruit, 1d8 of  which act as randomly determined magic  potions, while one acts as an ingested poison of  the GM's choice. The tree vanishes after 1 hour.  Picked fruit remains, retaining any magic for 30  days.                                                                                                                        
81 - 90   A nest of 1d4 + 3 eggs springs up. Any creature                                                                                                                                                                                                                                                                                                                                

       that eats an egg must make a DC 20 Constitution  saving throw. On a successful save, a creature  permanently increases its lowest ability score by  1, randomly choosing among equally low scores.  On a failed save, the creature takes 10d6 force  damage from an internal magical explosion.  
91-99  A pyramid with a 60-- foot-- square base bursts  upward. Inside is a sarcophagus containing a  mummy lord. The pyramid is treated as the  mummy lord's lair, and its sarcophagus con ta ins  treasure of the GM's choice.                                                                        
00     A giant beanstalk sprouts, growing to a height of  the GM's choice. The top leads where the GM  chooses, such as to a great view, a cloud giant's  castle, or a different plane of existence.                                                                                                    


Bag of Devouring
Wondrous item, Hero This bag superficially resembles a bag of holding but is a feeding orifice for a gigantic extradimensional creature. Turning the bag inside out closes the orifice.  The extradimensional creature attached to the bag can sense whatever is placed inside the bag. Animal or vegetable matter placed wholly in the bag is devoured and lost forever. When part of a living creature is placed in the bag, as happens when someone reaches inside it, there is a 50 percent chance that the creature is pulled inside the bag. A creature inside the bag can use its action to try to escape with a successful DC 15 Strength check. Another creature can use its action to reach into the bag to pull a creature out, doing so with a successful DC 20 Strength check (provided it isn't pulled inside the bag first). Any creature that starts its turn inside the bag is devoured, its body destroyed.  Inanimate objects can be stored in the bag, which can hold a cubic foot of such material. However, once each day, the bag swallows any objects inside it and spits them out into another plane of existence. The GM determines the time and plane.  If the bag is pierced or torn, it is destroyed, and anything contained within it is transported to a random location on the Astral Plane.



Deck of Illusions
Wondrous item, Adventurer This box contains a set of parchment cards. A full deck has 34 cards. A deck found as treasure is usually missing 1d20 \textminus 1 cards.  The magic of the deck functions only if cards are drawn at random (you can use an altered deck of playing cards to simulate the deck). You can use an action to draw a card at random from the deck and throw it to the ground at a point within 30 feet of you.  An illusion of one or more creatures forms over the thrown card and remains until dispelled. An illusory creature appears real, of the appropriate size, and behaves as if it were a real creature except that it can do no harm. While you are within 120 feet of the illusory creature and can see it, you can use an action to move it magically anywhere within 30 feet of its card. Any physical interaction with the illusory creature reveals it to be an illusion, because objects pass through it. Someone who uses an action to visually inspect the creature identifies it as illusory with a successful DC 15 Intelligence (Investigation) check. The creature then appears translucent.
The illusion lasts until its card is moved or the illusion is dispelled. When the illusion ends, the image on its card disappears, and that card can't be used again.
Playing Card       Illusion                            
Ace of hearts      Red dragon                          
King of hearts     Knight and four guards              
Queen of hearts    Succubus or incubus                 
Jack of hearts     Druid                               
Ten of hearts      Cloud giant                         
Nine of hearts     Ettin                               
Eight of hearts    Bugbear                             
Two of  hearts     Goblin                              
Ace of diamonds    Beholder                            
King of diamonds   Archmage and mage apprentice        
Queen of diamonds  Night hag                           
Jack of diamonds   Assassin                            
Ten of diamonds    Fire giant                          
Nine of diamonds   Ogre mage                           
Eight of diamonds  Gnoll                               
Two of diamonds    Kobold                              
Ace of  spades     Lich                                
King of spades     Priest and two acolytes             
Queen of spades    Medusa                              
Jack of spades     Veteran                             
Ten of spades      Frost giant                         
Nine of spades     Troll                               
Eight of spades    Hobgoblin                           
Two of spades      Goblin                              
Ace of clubs       Iron golem                          
King of clubs      Bandit captain and   three bandits  
Queen of clubs     Erinyes                             
Jack of clubs      Berserker                           
Ten of clubs       Hill giant                          
Nine of clubs      Ogre                                
Eight of clubs     Orc                                 
Two of clubs       Kobold                              
Jokers (2)         You (the deck's owner)












