\section{Wands and Rods}\label{mi:wands}

\subsection{Rod of Absorption}
\textit{Rod, Hero (requires attunement)}
While holding this rod, you can use your reaction to absorb a spell that is targeting only you and not with an area of effect. The absorbed spell's effect is canceled, and the spell's energy-not the spell itself-is stored in the rod. The energy has the same level as the spell when it was cast. The rod can absorb and store up to 50 levels of energy over the course of its existence. Once the rod absorbs 50 levels of energy, it can't absorb more. If you are targeted by a spell that the rod can't store, the rod has no effect on that spell.
When you become attuned to the rod, you know how many levels of energy the rod has absorbed over the course of its existence, and how many levels of spell energy it currently has stored.
If you are a spellcaster holding the rod, you can convert energy stored in it into spell slots to cast spells you have prepared or know. You can create spell slots only of a level equal to or lower than your own spell slots, up to a maximum of 5th level. You use the stored levels in place of your slots, but otherwise cast the spell as normal. For example, you can use 3 levels stored in the rod as a 3rd-level spell slot.
A newly found rod has 1d10 levels of spell energy stored in it already. A rod that can no longer absorb spell energy and has no energy remaining becomes nonmagical.

\subsection{Rod of Alertness}
Rod, Hero (requires attunement) 

This rod has a flanged head and the following properties.
Alertness. While holding the rod, you have advantage on Wisdom (Perception) checks and on rolls for initiative.
Spells. While holding the rod, you can use an action to cast one of the following spells from it: detect evil and good, detect magic, detect poison and disease, or see invisibility.
Protective Aura. As an action, you can plant the haft end of the rod in the ground, whereupon the rod's head sheds bright light in a 60-foot radius and dim light for an additional 60 feet. While in that bright light, you and any creature that is friendly to you gain a +1 bonus to AC and saving throws and can sense the location of any invisible hostile creature that is also in the bright light.  The rod's head stops glowing and the effect ends after 10 minutes, or when a creature uses an action to pull the rod from the ground. This property can't be used again until the next dawn.

Rod of Lordly Might
Rod, legendary (requires attunement) This rod has a flanged head, and it functions as a magic mace that grants a +3 bonus to attack and damage rolls made with it. The rod has properties associated with six different buttons that are set in a row along the haft. It has three other properties as well, detailed below.
Six Buttons. You can press one of the rod's six buttons as a bonus action. A button's effect lasts until you push a different button or until you push the same button again, which causes the rod to revert to its normal form.  If you press
button 1, the rod becomes a flame tongue, as a fiery blade sprouts from the end opposite the rod's flanged head.  If you press
button 2, the rod's flanged head folds down and two crescent-shaped blades spring out, transforming the rod into a magic battleaxe that grants a +3 bonus to attack and damage rolls made with it.  If you press
button 3, the rod's flanged head folds down, a spear point springs from the rod's tip, and the rod's handle lengthens into a 6-foot haft, transforming the rod into a magic spear that grants a +3 bonus to attack and damage rolls made with it.If you press
button 4, the rod transforms into a climbing pole up to 50 feet long, as you specify. In surfaces as hard as granite, a spike at the bottom and three hooks at the top anchor the pole. Horizontal bars 3 inches long fold out from the sides, 1 foot apart, forming a ladder. The pole can bear up to 4,000 pounds. More weight or lack of solid anchoring causes the rod to revert to its normal form.
If you press
button 5, the rod transforms into a handheld battering ram and grants its user a +10 bonus to Strength checks made to break through doors, barricades, and other barriers.
If you press
button 6, the rod assumes or remains in its normal form and indicates magnetic north. (Nothing happens if this function of the rod is used in a location that has no magnetic north.) The rod also gives you knowledge of your approximate depth beneath the ground or your height above it.
Drain Life.When you hit a creature with a melee attack using the
rod, you can force the target to make a DC 17 Constitution saving throw. On a failure, the target takes an extra 4d6 necrotic damage, and you regain a number of hit points equal to half that necrotic damage. This property can't be used again until the next dawn.
Paralyze.When you hit a creature with a melee attack using the rod, you can force the target to make a DC 17 Strength saving throw. On a failure, the target is paralyzed for 1 minute. The target can repeat the saving throw at the end of each of its turns, ending the effect on a success. This property can't be used again until the next dawn.
Terrify.While holding the rod, you can use an action to force each creature you can see within 30 feet of you to make a DC 17 Wisdom saving throw. On a failure, a target is frightened of you for 1 minute. A frightened target can repeat the saving throw at the end of each of its turns, ending the effect on itself on a success. This property can't be used again until the next dawn.

Rod of Rulership
Rod, Hero (requires attunement)
You can use an action to present the rod and command obedience from each creature of your choice that you can see within 120 feet of you. Each target must succeed on a DC 15 Wisdom saving throw or be charmed by you for 8 hours. While charmed in this way, the creature regards you as its
trusted leader. If harmed by you or your companions, or commanded to do something contrary to its nature, a target ceases to be charmed in this way. The rod can't be used again until the next dawn.

Rod of Security
Rod, Hero While holding this rod, you can use an action to activate it. The rod then instantly transports you and up to 199 other willing creatures you can see to a paradise that exists in an extraplanar space. You choose the form that the paradise takes. It could be a tranquil garden, lovely glade, cheery tavern, immense palace, tropical island, fantastic carnival, or whatever else you can imagine. Regardless of its nature, the paradise contains enough water and food to sustain its visitors. Everything else that can be interacted with inside the extraplanar space can exist only there. For example, a flower picked from a garden in the paradise disappears if it is taken outside the extraplanar space.  For each hour spent in the paradise, a visitor regains hit points as if it had spent 1 Hit Die. Also, creatures don't age while in the paradise, although time passes normally. Visitors can remain in the paradise for up to 200 days divided by the number of creatures present (round down).  When the time runs out or you use an action to end it, all visitors reappear in the location they occupied when you activated the rod, or an unoccupied space nearest that location. The rod can't be used again until ten days have passed.

Wand of Binding
Wand, Hero (requires attunement by a spellcaster) This wand has 7 charges for the following properties. It regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.
Spells. While holding the wand, you can use an action to expend some of its charges to cast one of the following spells (save DC 17): hold monster (5 charges) or hold person (2 charges).
Assisted Escape. While holding the wand, you can use your reaction to expend 1 charge and gain advantage on a saving throw you make to avoid being paralyzed or restrained, or you can expend 1 charge and gain advantage on any check you make to escape a grapple.

Wand of Enemy Detection
Wand, Hero (requires attunement)
This wand has 7 charges. While holding it, you can use an action and expend 1 charge to speak its command word. For the next minute, you know the direction of the nearest creature hostile to you within 60 feet,
but not its distance from you. The wand can sense the presence of hostile creatures that are ethereal, invisible, disguised, or hidden, as well as those in plain sight. The effect ends if you stop holding the wand.
The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Fear
Wand, Hero (requires attunement)
This wand has 7 charges for the following properties. It regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.
Command. While holding the wand, you can use an action to expend 1 charge and command another creature to flee or grovel, as with the
commandspell (save DC 15).
Cone of Fear. While holding the wand, you can use an action to expend 2 charges, causing the wand's tip to emit a 60-foot cone of amber light. Each creature in the cone must succeed on a DC 15 Wisdom saving throw or become frightened of you for 1 minute. While it is frightened in this way, a creature must spend its turns trying to move as far away from you as it can, and it can't willingly move to a space within 30 feet of you. It also can't take reactions. For its action, it can use only the Dash action or try to escape from an effect that prevents it from moving. If it has nowhere it can move, the creature can use the Dodge action. At the end of each of its turns, a creature can repeat the saving throw, ending the effect on itself on a success.

Wand of Fireballs
Wand, Hero (requires attunement by a spellcaster)
This wand has 7 charges. While holding it, you can use an action to expend 1 or more of its charges to cast the
fireballspell (save DC 15) from it. For 1 charge, you cast the 3rd-level version of the spell. You can increase the spell slot level by one for each additional charge you expend.
The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a
d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Lightning Bolts
Wand, Hero (requires attunement by a spellcaster) This wand has 7 charges. While holding it, you can use an action to expend 1 or more of its charges to cast the lightning bolt spell (save DC 15) from it. For 1 charge, you cast the 3rd-level version of the spell. You can increase the spell slot level by one for each additional charge you expend.  The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Magic Detection
Wand, Adventurer This wand has 3 charges. While holding it, you can expend 1 charge as an action to cast the detect magicspell from it. The wand regains 1d3 expended charges daily at dawn.

Wand of Magic Missiles
Wand, Adventurer This wand has 7 charges. While holding it, you can use an action to expend 1 or more of its charges to cast the magic missile spell from it. For 1 charge, you cast the 1st-level version of the spell. You can increase the spell slot level by one for each additional charge you expend.  The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Paralysis
Wand, Hero (requires attunement by a spellcaster) This wand has 7 charges. While holding it, you can use an action to expend 1 of its charges to cause a thin blue ray to streak from the tip toward a creature you can see within 60 feet of you. The target must succeed on a DC 15 Constitution saving throw or be paralyzed for 1 minute. At the end of each of the target's turns, it can repeat the saving throw, ending the effect on itself on a success.  The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Polymorph
Wand, Hero (requires attunement by a spellcaster)
This wand has 7 charges. While holding it, you can use an action to expend 1 of its charges to cast the
polymorphspell (save DC 15) from it.
The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Secrets
Wand, Adventurer
The wand has 3 charges. While holding it, you can use an action to expend 1 of its charges, and if a secret door or trap is within 30 feet of you, the wand pulses and points at the one nearest to you. The wand regains 1d3 expended charges daily at dawn.

Wand of Web
Wand, Adventurer (requires attunement by a spellcaster)
This wand has 7 charges. While holding it, you can use an action to expend 1 of its charges to cast the
webspell (save DC 15) from it.
The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into ashes and is destroyed.

Wand of Wonder
Wand, Hero (requires attunement by a spellcaster)
This wand has 7 charges. While holding it, you can use an action to expend 1 of its charges and choose a target within 120 feet of you. The target can be a creature, an object, or a point in space. Roll d100 and consult the following table to discover what happens.
If the effect causes you to cast a spell fromthe wand, the spell's save DC is 15. If the spell normally has a range expressed in feet, its range becomes 120 feet if it isn't already.
If an effect covers an area, you must center the spell on and include the target. If an effect has
multiple possible subjects, the GM randomly determines which ones are affected.  The wand regains 1d6 + 1 expended charges daily at dawn. If you expend the wand's last charge, roll a d20. On a 1, the wand crumbles into dust and is destroyed.
d100                                              Effect                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  
01-05                                             You cast  slow.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                         
06-10                                             You cast faerie fire.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                   
11-15                                             You are stunned until the start of your next turn,  believing something awesome just happened.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                          
16-20                                             You cast gust of wind.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                  
21 - 25                                           You cast detect thoughts on the target you  chose. If you didn't target a creature,  you  instead take 1d6 psychic damage.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                              
26-30                                             You cast stinking cloud.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
31 - 33                                           Heavy rain falls in a 60 -- foot radius centered on  the target. The area becomes lightly obscured.  The rain falls until the start of your next turn.                                                                                                                                                                                                                                                                                                                                                                                                                                                                  
34-36                                             An animal appears in the unoccupied space  nearest the target. The animal isn't under your  control and acts as it normally would. Roll a d100  to determine which animal appears. On a 01 -25,  a rhinoceros appears; on a 26 -50, an elephant  appears; and on a 51 - 1 00, a rat appears.                                                                                                                                                                                                                                                                                                                            
37-46                                             You cast lightning bolt.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                
47-49                                             A cloud of 600 oversized butterflies fills a 30-- foot  radius centered on the target. The area becomes  heavily obscured. The butterflies remain for 10  minutes.                                                                                                                                                                                                                                                                                                                                                                                                                                                      
50 - 53                                           You enlarge the target as if you  had cast  enlarge/reduce. If the target can't be affected by  that spell, or if you didn't target a creature, you  become the target.                                                                                                                                                                                                                                                                                                                                                                                                                                                 
54 - 58                                           You cast darkness.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      
59-62                                             Grass grows on the ground in a 60 --foot radius  centered on the target. If grass is already  there,  it grows to ten times its normal size and remains  overgrown for 1 minute.                                                                                                                                                                                                                                                                                                                                                                                                                                        
63-65                                             An object of the GM's choice disappears into the  Ethereal Plane. The object must be neither worn  nor carried, within 120 feet of the target, and no  larger than 10 feet  in any dimension.                                                                                                                                                                                                                                                                                                                                                                                                                           
66-69                                             You shrink yourself as if you had cast  enlarge/reduce on yourself.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                     
70 - 79                                           You cast fireball.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      
80-84                                             You cast invisibility on yourself.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                                      
85-87                                             Leaves grow from the target. If you chose a point  in space as the target, leaves sprout  from the  creature nearest to that point. Unless they are picked off, the leaves turn brown and fall off  after 24 hours.                                                                                                                                                                                                                                                                                                                                                                                                     
88-90                                             A stream of 1d4 \texttimes 10 gems, each worth 1 gp,  shoots from the wand's tip in a line 30 feet long  and 5 feet wide. Each gem deals 1 b ludgeoning  damage, and the total damage of the gems is                                                                                                                                                                                                                                                                                                                                                                                                                             
divided equally among all creatures in the line.  
91 - 95                                           A burst of colorful shimmering light extends  from you in a 30 -- foot radius. You and each  creature in the area that can see must succeed  on a DC 15   Constitution saving throw or become  blinded for 1 minute. A creature can repeat the  saving throw at the end of each of its turns,  ending the effect on itself on a success.                                                                                                                                                                                                                                                                                
96 - 97                                           The target's skin turns bright blue for 1d10 days.  If you chose a point in s pace, the creature  nearest to that point is affected.                                                                                                                                                                                                                                                                                                                                                                                                                                                                                    
98 - 00                                           If you targeted a creature, it must make a DC 15  Constitution saving throw. If you didn't target a  creature, you become the target and must make  the saving throw. If the saving throw fails by 5 o r  more, the target is instantly petrified. On any  other failed save, the target is restrained and  begins to turn to stone. While restrained in this  way, the target must repeat the saving throw at  the end of its next turn, becoming petrified on a  failure or   ending the effect on a success. The  petrification lasts until the target is freed by the  greater restoration spell or similar magic. 