\section{Armsman}

The armsman is a master of weapons. His power comes from training and discipline and superior techniques that allow him to break the boundaries of what is possible for others.

Design Discussion: This replaces the fighter. Heavy armor is the signature, plus getting to attack more than most. Gets ability to use any weapon with either ability score. Some subclass-level overt magic. SDCT 3/6/5/6. Control is mostly in the form of punishing those who ignore them.

\subsection{Class Features}

As a armsman, you gain the following class features.

\subsubsection{Hit Points}

\textbf{Hit Dice:} 1d10 per armsman level

\textbf{Hit Points at 1st Level:} 10 + your Constitution modifier

\textbf{Hit Points at Higher Levels:} 1d10 (or 6) + your Constitution modifier per armsman level after 1st

\subsubsection{Proficiencies}

\textbf{Armor:} All armor, shields

\textbf{Weapons:} Simple weapons, martial weapons

\textbf{Tools:} None

\textbf{Saving Throws:} Strength, Constitution

\textbf{Skills:} Choose two skills from Acrobatics, Animal, Handling, Athletics, History, Insight, Intimidation, Perception, and Survival

\subsubsection{Equipment}

You start with the following equipment, in addition to the equipment granted by your background:

\begin{itemize}
\item (\textit{a}) chain mail or (\textit{b}) leather armor, longbow, and 20 arrows
\item (\textit{a}) a martial weapon and a shield or (\textit{b}) two martial weapons
\item (\textit{a}) a light crossbow and 20 bolts or (\textit{b}) two handaxes
\item (\textit{a}) a dungeoneer's pack or (\textit{b}) an explorer's pack
\end{itemize}

\textbf{The Armsman (table)}


\begin{DndTable}[header=Armsman\label{tbl:armsman}]{XXXXXX}
 \textbf{Level} & \textbf{Proficiency Bonus} & \textbf{Features} & \textbf{Stamina} & \textbf{Aether} & \textbf{Aether Limit}\\ 
 1st   & +2  & Fighting Style, Second Wind                       & 1 + CON        & 1     & 1 \\
 2nd   & +2  & Action Surge                           			 & 2 + CON        & 1     & 1 \\
 3rd   & +2  & Martial Archetype                                 & 3 + CON         & 2     & 1 \\
 4th   & +2  & Ability Score Improvement                         & 4 + CON         & 2     & 1 \\
 5th   & +3  & Extra Attack                                      & 5 + CON         & 3     & 2 \\
 6th   & +3  & Versatility                         				 & 6 + CON         & 3     & 2 \\
 7th   & +3  & Martial Archetype Feature                         & 7 + CON         & 4     & 2 \\
 8th   & +3  & Ability Score Improvement                         & 8 + CON         & 4     & 2 \\
 9th   & +4  & Indomitable, Everything's a Weapon                & 9 + CON         & 5     & 2 \\
 10th  & +4  & Martial Archetype Feature                         & 10 + CON        & 5     & 3 \\
 11th  & +4  & Extra Attack (2)                                  & 11 + CON        & 6     & 3 \\
 12th  & +4  & Ability Score Improvement                         & 12 + CON        & 6     & 3 \\
 13th  & +5  & Flash Step                            			 & 13 + CON        & 7     & 3 \\
 14th  & +5  & Deathblow                         				 & 14 + CON        & 7     & 3 \\
 15th  & +5  & Martial Archetype Feature                         & 15 + CON        & 8     & 3 \\ 
 16th  & +5  & Ability Score Improvement                         & 16 + CON        & 8     & 3 \\
 17th  & +6  & Extra Attack (3) 								 &17 + CON        & 9     & 4 \\
 18th  & +6  & Martial Archetype Feature                         &18 + CON        & 9     & 4 \\
 19th  & +6  & Ability Score Improvement                         &19 + CON        & 10     & 4 \\
 20th  & +6  & Improved Deathblow                                  &20 + CON        & 10     & 4 \\
\end{DndTable}


\subsection{Fighting Style}

You are better than most at using the additional properties of your weapon. You gain a bonus depending on the additional property. If the weapon has multiple additional properties, you must choose which bonus to apply on any individual attack. If a bonus calls for a saving throw, the DC = 8 + your Strength modifier + your proficiency bonus.

\subparagraph*{Battering} Once per turn when you hit with a battering weapon, you can force the target to make a Strength saving throw. On a failed save, the target is knocked prone.

\subparagraph*{Cleaving} You can attempt to cleave even if you miss. If you do so, roll a new attack with the same modifiers and compare it to the new target's AC.

\subparagraph*{Heavy} You can choose to forgo your proficiency bonus to the attack roll. If you still hit, you can add twice your proficiency bonus to the damage dealt.

\subparagraph*{Light} When you make the additional attack with a light weapon, you add your ability modifier to the damage dealt.

\subparagraph*{Loading} You ignore the normal effect of this property. Instead, when you hit with an attack from a loading weapon and drop the target to 0 HP, you can choose to have the bolt pass through at a creature behind the slain creature. The closest creature on a 5' wide line connecting you to the slain creature and extending 30' behind him acts as the new target. Make an attack at disadvantage against that creature. If it hits, it takes damage as normal from the attack.

\subparagraph*{Parrying} The bonus from this property increases to +4.

\subparagraph*{Precise} You score a critical hit on an 18, 19, or 20 instead of on a 19 or 20.

\subparagraph*{Reach} You can make opportunity attacks when a creature enters your range as well as leaves it.

\subparagraph*{Thrown} You can draw thrown weapons as part of the attack. In addition, the damage die increases by one step when thrown and you do not suffer disadvantage out to the long range of the attack.

\subparagraph*{Versatile} You get the increased damage die even when wielding it in one hand.

\subsection{Second Wind}

You have a limited well of stamina that you can draw on to protect yourself from harm. On your turn, you can spend 1 STA use a bonus action to regain hit points equal to 1d10 + your armsman level. The cost increases by 1 STA each time you use it, resetting to 1 when you finish a long or short rest.

\subsection{Action Surge}

Starting at 2nd level, you can push yourself beyond your normal limits for a moment. On your turn, you can spend 2 STA to take one additional action on top of your regular action and a possible bonus action.

Once you use this feature, you cannot use this feature again until you next roll initiative. Starting at 17th level, you can use it as many times as you have stamina for, but only once on the same turn.

\subsection{Martial Archetype}

At 3rd level, you choose an archetype that you strive to emulate in your combat styles and techniques. Choose TODO, all detailed at the end of the class description. The archetype you choose grants you features at 3rd level and again at 7th, 10th, 15th, and 18th level.

\subsection{Ability Score Improvement}

When you reach 4th level, and again at 6th, 8th, 12th, 16th, and 19th level, you can increase one ability score of your choice by 2, or you can increase two ability scores of your choice by 1. As normal, you can't increase an ability score above 20 using this feature.

\subsection{Extra Attack}

Beginning at 5th level, you can attack twice, instead of once, whenever you take the Attack action on your turn.

The number of attacks increases to three when you reach 11th level in this class and to four when you reach 17th level in this class.

\subsection{Versatility}
Beginning at 6th level, you can use Strength or Dexterity as the modifier for weapon attacks and damage regardless of the type of weapon.

In addition, you can choose to make a Strength check when a Dexterity check would otherwise be called for and vice versa.

\subsection{Everything's a Weapon}
Beginning at 9th level, you've discovered that the same techniques you use with your weapons also applies to other situations. Choose one of the approaches below.

\begin{itemize}
	\item \textbf{Direct.} When you make an ability check involving Intimidation, Athletics, or any ability check involving Constitution, you may add twice your proficiency bonus instead of any proficiency (including none) that may have applied.
	\item \textbf{Gregarious.} When you make an ability check involving Charisma or Insight, you may add twice your proficiency bonus instead of any proficiency (including none) that may have applied.
	\item \textbf{Inquisitive.} When you make an ability check involving Intelligence, Insight, or Animal Handling, you may add twice your proficiency bonus instead of any proficiency (including none) that may have applied.
	\item \textbf{Intuitive.} When you make an ability check involving Wisdom or any saving throw against being charmed or scryed on, you may add twice your proficiency bonus instead of any proficiency (including none) that may have applied.
	\item \textbf{Subtle.} When you make an ability check involving Acrobatics, Stealth, Sleight of Hand, or Deception, you may add twice your proficiency bonus instead of any proficiency (including none) that may have applied.
\end{itemize}

\subsection{Indomitable}

Beginning at 9th level, you can choose to succeed on a saving throw instead of rolling. If you do so, you cannot use this feature again until you finish a long rest.

You can use this feature twice between long rests starting at 13th level and three times between long rests starting at 17th level.

\subsection{Flash Step}
Starting at 13th level, you have learned to move so fast over short distaneces that it appears you can teleport. When you move on your turn, you can choose to instead teleport to the chosen location as long as you have a clear path to the target location and it is no further than your speed would allow. The clear path to the target does not have to be in a straight line, but you cannot pass through areas too small to squeeze through.

\subsection{Deathblow}
Starting at 14th level, you can attempt to strike down a wounded foe. When you hit an enemy with a weapon attack and the enemy has less than 25 HP, you can choose to spend 5 STA to drop it to 0 HP instead of dealing normal damage. You can choose whether this is lethal or nonlethal if you hit with a melee attack.

If the target has hit points above this threshold, you can choose to force the target to make a Constitution saving throw at disadvantage against a DC of 8 + your Strength + your proficiency bonus. On a failed save they are stunned until the end of your next turn. Alternatively, you can choose to refund the STA spent.

\subsection{Improved Deathblow}
Starting at 20th level, your touch is death for most weaker foes. You no longer need to expend STA to use Deathblow if the target's current or maximum HP is below 50, and can spend stamina to use Deathblow as long as the target's current HP is under 100.

\section{Martial Archetypes}

Different armsmen choose different approaches to perfecting their fighting prowess. The martial archetype you choose to emulate reflects your approach.

\subsection{Defender}
The defender archetype focuses on protecting allies from harm while locking down their enemies. While most use a one-handed weapon and a shield, others trust in their heavy armor to protect them.

\subsubsection{Thicket of Blades}
Starting at 3rd level, opponents provoke opportunity attacks from you by moving within your reach, making attacks against anyone but you, or casting a spell. You can spend 1 STA to make an opportunity attack without consuming your reaction. No individual creature can provoke more than one opportunity attack per movement. Example: if a goblin starts 20' away and runs past a Defender without stopping, the defender can only make one opportunity attack but may make it either when they move within the reach or move out of the reach. But if the goblin attacks an ally and then moves within reach, the Defender can make two opportunity attacks by spending 1 STA and his reaction or 2 STA (keeping his reaction). 

\subsubsection{Shielding Bulwark}
Starting at 3rd level, when you are the target of an effect that targets a point in space and allows a Dexterity saving throw to take half damage, you can use your weapon and shield to diffuse and deflect the energies, protecting yourself and those behind you. Expend 2 STA. You and all creatures in a 15' long line that is 10' wide behind you (relative to the target point in space) gain advantage on the saving throw and the shielded creatures take no damage if they succeed on the saving throw and only half damage if they fail.

\subsubsection{Combat Challenge}
Starting at 7th level, your training has taught you how to magically compel a target to focus on you. As a bonus action, spend 2 AET and choose a target that can hear you and that you can see within 60 feet. The target must make a Charisma saving throw against a DC of 8 + your Charisma modifier + your proficiency bonus. On a failed save, the target cannot make attacks against anyone but you or target any of your allies with an ability (magical or otherwise). Additionally, they cannot willingly move further away from you. This effect lasts for one minute or until you use this ability against someone else or you are incapacitated.

\subsubsection{Calming Words}
Starting at 10th level you've become particularly adept at predicting what will cause controversy--words are weapons too. As such, you can add twice your proficiency bonus to any Charisma check you make to defuse tense situations. In addition, when an ally you can hear makes a Charisma (Persuasion) check and you Help them, they can add your Charisma (Persuasion) modifier to theirs when resolving the check.

\subsubsection{Resilience}
Starting at 15th level, you can ignore your wounds and keep fighting even through attacks that would normally incapacitate you. When you are reduced to 0 HP or would be outright killed (such as by a Power Word: Kill effect) and have less than 5 levels of exhaustion, you choose to instead heal to half your maximum HP. If you do so, you gain one level of exhaustion and cannot use this feature again until you finish a long or short rest.

\subsubsection{Total Prediction}
Starting at 18th level, you have trained enough to be able to magically sense your enemies' actions a few steps ahead. As an action you can expend 4 AET, gaining limited precognitive abilities for one hour. For the duration, you cannot be surprised and have advantage on attack rolls. Additionally, other creatures have disadvantage on attack rolls against you for the duration.

\subsection{Sword Saint}
Don't let the name confuse you, there are sword saints devoted to all forms of weapons. The sword saint takes their weapon skills to an entirely new level. On the battlefield they are a flash of lightning, a stroke of thunder, an explosion of strikes. Many of their techniques draw on pure aether, mixing magic and weapon play.

% \begin{DndTable}[header=Sword Saint]{XXXX}
% 	\textbf{Armsman Level} & \textbf{Feature} & \textbf{Aether} & \textbf{Aether Limit} \\ 
% 	3 & Exceptional Aether & 4 & 1 \\
% 	4 & -- & 6 & 2 \\
% 	5 & -- & 7 & 2 \\
% 	6 & -- & 9 & 2 \\
% 	7 & Something & 12 & 3 \\
% 	8 & -- & 13 & 3 \\
% 	9 & -- & 18 & 3 & \\
% 	10 & Something & 20 & 3 \\ 
% 	11 & -- & 23 & 3 \\
% 	12 & -- & 23 & 3 \\
% 	13 & -- & 27 & 5 \\
% 	14 & -- & 27 & 5 \\
% 	15 & Something & 30 & 5 \\ 
% 	16 & -- & 30 & 5 \\
% 	17 & -- & 33 & 5 \\
% 	18 & Something & 33 & 5 \\ 
% 	19 & -- & 36 & 8 \\
% 	20 & -- & 36 & 8 
% \end{DndTable}

\subsubsection{Exceptional Aether}
Starting at 3rd level, sword saints cultivate their aether thoroughly than most armsmen. Use the Sword Saint table instead of the Armsman table to determine your maximum aether and your aether limit as you gain levels.

\subsubsection{Weapon Flexibility}
Beginning at 3rd level, you can use your weapons in unexpected ways. At the beginning of your turn you can exchange any of the following weapon properties for any of the others in the list: Battering, Cleaving, Finesse, Parrying, Precise, or Thrown. If you chose to make 