\section{Brawler}

\subsection{Class Features}

As a brawler, you gain the following class features.

\subsection{Hit Points}

\textbf{Hit Dice:} 1d8 per brawler level

\textbf{Hit Points at 1st Level:} 8 + your Constitution modifier

\textbf{Hit Points at Higher Levels}: 1d8 (or 5) + your Constitution modifier per brawler level after 1st

\subsection{Proficiencies}

\textbf{Armor:} None

\textbf{Weapons:} Simple weapons, shortswords

\textbf{Tools:} Choose one type of artisan's tools or one musical instrument

\textbf{Saving Throws:} Strength, Dexterity

\textbf{Skills:} Choose two from Acrobatics, Athletics, History, Insight, Religion, and Stealth

\subsection{Equipment}

You start with the following equipment, in addition to the equipment granted by your background:
\begin{itemize}
\item (\textit{a}) a shortsword or (\textit{b}) any simple weapon
\item (\textit{a}) a dungeoneer's pack or (\textit{b}) an explorer's pack
\item 10 darts
\end{itemize}

\begin{DndTable}[header=The Brawler\label{tbl:brawler}]{XXXXXXXX}
 Level & Proficiency Bonus & Martial Arts & Unarmored Movement & Features & Stamina & Aether & Aether Limit \\                                        
 1st   & +2 & 1d6  & -       & Unarmored Defense, Martial Arts             & 1 + CON & 1 & 1 \\                 
 2nd   & +2 & 1d6  & +10 ft. & Conditioning, Unarmored Movement      & 2 + CON + WIS & 1 & 1 \\                 
 3rd   & +2 & 1d6  & +10 ft. & Martial Tradition, Deflect Missiles   & 3 + CON + WIS & 2 & 1 \\            
 4th   & +2 & 1d6  & +10 ft. & Ability Score Improvement, Slow Fall  & 4 + CON + WIS & 2 & 1 \\            
 5th   & +3 & 1d8  & +10 ft. & Extra Attack, Stunning Strike         & 5 + CON + WIS & 3 & 2 \\            
 6th   & +3 & 1d8  & +15 ft. & Empowered Strikes, Martial Tradition Feature & 6 + CON + WIS & 3 & 2\\
 7th   & +3 & 1d8  & +15 ft. & Evasion, Stillness of Mind           & 7 + CON + WIS & 4 & 2\\  
 8th   & +3 & 1d8  & +15 ft. & Ability Score Improvement            & 8 + CON + WIS & 4 & 2\\            
 9th   & +4 & 1d8  & +15 ft. & Unarmored Movement improvement       & 9 + CON + WIS & 5 & 2\\            
 10th  & +4 & 1d8  & +20 ft. & Purity of Body                      & 10 + CON + WIS & 5 & 3\\            
 11th  & +4 & 1d10  & +20 ft. & Martial Tradition Feature         & 11 + CON + WIS & 6 & 3\\            
 12th  & +4 & 1d10  & +20 ft. & Ability Score Improvement          & 12 + CON + WIS & 6 & 3\\            
 13th  & +5 & 1d10  & +20 ft. &          & 13 + CON + WIS & 7 & 3\\            
 14th  & +5 & 1d10  & +25 ft. & Diamond Soul                       & 14 + CON + WIS & 7 & 3\\            
 15th  & +5 & 1d10  & +25 ft. & Timeless Body                      & 15 + CON + WIS & 8 & 3\\            
 16th  & +5 & 1d10  & +25 ft. & Ability Score Improvement          & 16 + CON + WIS & 8 & 3\\            
 17th  & +6 & 1d12 & +25 ft. & Martial Tradition Feature          & 17 + CON + WIS & 9 & 4\\            
 18th  & +6 & 1d12 & +30 ft. & Empty Body                          & 18 + CON + WIS & 9 & 4\\            
 19th  & +6 & 1d12 & +30 ft. & Ability Score Improvement           & 19 + CON + WIS & 10 & 4\\            
 20th  & +6 & 1d12 & +30 ft. & Perfect Self                        & 20 + CON + WIS & 10 & 4\\            
\end{DndTable}

\subsection{Unarmored Defense}

Beginning at 1st level, while you are wearing no armor and not wielding a shield, your AC equals 10 + your Dexterity modifier or your Strength modifier (whichever is higher) + your Wisdom modifier.

\subsection{Martial Arts}

At 1st level, your practice of martial arts gives you mastery of combat styles that use unarmed strikes and brawler weapons, which are shortswords and any simple melee weapons that don't have the two- handed or heavy property.

You gain the following benefits while you are unarmed or wielding only brawler weapons and you aren't wearing armor or wielding a shield:

\begin{itemize}
\item You can use Dexterity instead of Strength for the attack and damage rolls of your unarmed strikes and brawler weapons.
\item You can roll a d6 in place of the normal damage of your unarmed strike or brawler weapon. This die changes as you gain brawler levels, as shown in the Martial Arts column of the Brawler table.
\item When you use the Attack action with an unarmed strike or a brawler weapon on your turn, you can make one unarmed strike as a bonus action. For example, if you take the Attack action and attack with a quarterstaff, you can also make an unarmed strike as a bonus action, assuming you haven't already taken a bonus action this turn.
\item You can grapple and shove creatures that are 2 sizes larger than yourself instead of only 1 size larger.
\end{itemize}

Certain brawler traditions use specialized forms of the brawler weapons. For example, you might use a club that is two lengths of wood connected by a short chain (called a nunchaku) or a sickle with a shorter, straighter blade (called a kama). Whatever name you use for a brawler weapon, you can use the game statistics provided for the weapon.

\subsection{Flurry of Blows}

Immediately after you take the Attack action on your turn, you can spend 1 STA to make two unarmed strikes as a bonus action.

\subsection{Patient Defense}

You can spend 1 STA to take the Dodge action as a bonus action on your turn.

\subsection{Step of the Wind}

You can spend 1 STA to take the Disengage or Dash action as a bonus action on your turn, and your jump distance is doubled for the turn.

\subsection{Conditioning}
Beginning at 2nd level, your understanding of the world around you has made you more efficient. In effect, your Stamina pool is larger than most. You add your Wisdom modifier (if positive) to your Stamina pool. This is included in the Stamina column of the \nameref{tbl:brawler} table. 

\subsection{Unarmored Movement}

Starting at 2nd level, your speed increases by 10 feet while you are not wearing armor or wielding a shield. This bonus increases when you reach certain brawler levels, as shown in the Brawler table.

At 9th level, you gain the ability to move along vertical surfaces, upside down, and across liquids without falling.

\subsection{Martial Tradition}

When you reach 3rd level, you commit yourself to a martial tradition: the Way of the Open Hand, the Way of Shadow, or the Way of the Four Elements, all detailed at the end of the class description. Your tradition grants you features at 3rd level and again at 6th, 11th, and 17th level.

\subsection{Deflect Missiles}

Starting at 3rd level, you can use your reaction to deflect or catch the missile when you are hit by a ranged weapon attack. When you do so, the damage you take from the attack is reduced by 1d10 + your Dexterity or Strength modifier + your brawler level.

If you reduce the damage to 0, you can catch the missile if it is small enough for you to hold in one hand and you have at least one hand free. If you catch a missile in this way, you can spend 1 STA to make a ranged attack with the weapon or piece of ammunition you just caught, as part of the same reaction. You make this attack with proficiency, regardless of your weapon proficiencies, and the missile counts as a brawler weapon for the attack, which has a normal range of 20 feet and a long range of 60 feet. You can expend extra STA to increase the damage dealt by 2 points for every extra STA expended.

\subsection{Ability Score Improvement}

When you reach 4th level, and again at 8th, 12th, 16th, and 19th level, you can increase one ability score of your choice by 1. As normal, you can't increase an ability score above +5 using this feature.

You can also pick a Skill Trick but you must meet the prerequisites for skill tricks learned in this way. See \nameref{ch:skill-tricks} for that list. You can swap out a known skill trick for another you can learn when you gain another skill trick.

\subsection{Slow Fall}

Beginning at 4th level, you can use your reaction when you fall to reduce any falling damage you take by an amount equal to five times your brawler level.

\subsection{Extra Attack}

Beginning at 5th level, you can attack twice, instead of once, whenever you take the Attack action on your turn.

\subsection{Stunning Strike}

Starting at 5th level, you can interfere with the flow of energy in an opponent's body. When you hit another creature with a melee weapon attack, you can spend 2 STA to attempt a stunning strike. The target must succeed on a Constitution saving throw or be stunned until the end of your next turn.

\subsection{Empowered Strikes}

Starting at 6th level, your unarmed strikes count as magical for the purpose of overcoming resistance and immunity to nonmagical attacks and damage. In addition, you can expend 1 AET to increase the damage dealt by your Martial Arts die.

\subsection{Evasion}

At 7th level, your instinctive agility lets you dodge out of the way of certain area effects, such as a blue dragon's lightning breath or a \textit{fireball} spell. When you are subjected to an effect that allows you to make a Dexterity saving throw to take only half damage, you instead take no damage if you succeed on the saving throw, and only half damage if you fail.

\subsection{Stillness of Mind}

Starting at 7th level, you can use your action to end one effect on yourself that is causing you to be charmed or frightened.

Additionally, on your turn you can end an effect on yourself that causes you to be incapacitated by spending 5 STA without taking an action.

\subsection{Purity of Body}

At 10th level, your mastery of the energy flowing through you makes you immune to disease and poison.

\subsection{Diamond Soul}

Beginning at 14th level, your self-mastery grants you proficiency in all saving throws.

Additionally, whenever you make a saving throw and fail, you can spend 1 STA to reroll it and take the second result.

\subsection{Empty Body}

Beginning at 18th level, you can use your action to spend 4 STA to become invisible for 1 minute. During that time, you also have resistance to all damage but force damage. This invisibility does not end when you make an attack roll, but it does end early if you are incapacitated at the start of your turn.

\subsection{Perfect Self}

At 20th level, if you are reduced to zero hit points, you can spend 10 STA as a reaction to instead heal to full health.

\subsection{Martial Traditions}

Three traditions of martial pursuit are common in the monasteries scattered across the multiverse. Most monasteries practice one tradition exclusively, but a few honor the three traditions and instruct each brawler according to his or her aptitude and interest. All three traditions rely on the same basic techniques, diverging as the student grows more adept. Thus, a brawler need choose a tradition only upon reaching 3rd level.

\subsection{Way of the Open Hand}

Brawlers of the Way of the Open Hand are the ultimate masters of martial arts combat, whether armed or unarmed. They learn techniques to push and trip their opponents, manipulate ki to heal damage to their bodies, and practice advanced meditation that can protect them from harm.

\subsection{Open Hand Technique}

Starting when you choose this tradition at 3rd level, you can manipulate your enemy's ki when you harness your own. Whenever you hit a creature with one of the attacks granted by your Flurry of Blows, you can impose one of the following effects on that target:
\begin{itemize}
\item It must succeed on a Dexterity saving throw or be knocked prone.
\item It must make a Strength saving throw. If it fails, you can push it up to 15 feet away from you.
\item It can't take reactions until the end of your next turn.
\end{itemize}

\subsection{Wholeness of Body}

At 6th level, you gain the ability to heal yourself. As an action, you can spend 2 AET to regain hit points equal to three times your brawler level. By expending extra AET, you increase the amount healed by your brawler level per extra AET spent.

\subsection{Tranquility}


\subsection{Quivering Palm}

At 17th level, you gain the ability to set up lethal vibrations in someone's body. When you hit a creature with an unarmed strike, you can spend 3 AET to start these imperceptible vibrations, which last for a number of days equal to your brawler level. The vibrations are harmless unless you use your action to end them. To do so, you and the target must be on the same plane of existence. When you use this action, the creature must make a Constitution saving throw. If it fails, it is reduced to 0 hit points. If it succeeds, it takes 10d10 necrotic damage.

You can have only one creature under the effect of this feature at a time. You can choose to end the vibrations harmlessly without using an action.