\section{Warden}

Design Discussion: Barbarian replacement. Overtly fueled by primal power. Meat tank. Big heavy hits. SDCT 2/7/4/7.

\subsection{Class Features}

As a warden, you gain the following class features.

\subsubsection{Hit Points}

\textbf{Hit Dice:} 1d12 per warden level

\textbf{Hit Points at 1st Level:} 12 + your Constitution modifier

\textbf{Hit Points at Higher Levels:} 1d12 (or 7) + your Constitution modifier per warden level after 1st

\subsubsection{Proficiencies}

\textbf{Armor:} Light armor, medium armor, shields

\textbf{Weapons:} Simple weapons, martial weapons

\textbf{Tools:} None

\textbf{Saving Throws:} Strength, Constitution

\textbf{Skills:} Choose two from Animal Handling, Athletics, Intimidation, Nature, Perception, and Survival

\subsubsection{Equipment}

You start with the following equipment, in addition to the equipment granted by your background:

\begin{itemize}
\item  (\textit{a}) a greataxe or (\textit{b}) any martial melee weapon
\item  (\textit{a}) two handaxes or (\textit{b}) any simple weapon
\item  An explorer’s pack and four javelins
\end{itemize}

\textbf{The Warden (table)}
\onecolumn
\begin{DndTable}[header=The Warden\label{tbl:warden}]{XXXXXX}
 \textbf{Level}  & \textbf{Proficiency Bonus} & \textbf{Features}       & \textbf{Stamina} & \textbf{Aether} & \textbf{Aether Limit} \\
 1st    & +2 & Exceptional Stamina, Furious Blow, Rage, Unarmored Defense    & 1 + CON         & 1     & 1 \\
 2nd    & +2 & Reckless Attack, Danger Sense & 2 + CON         & 1     & 1 \\
 3rd    & +2 & Primal Path                   & 3 + CON         & 2     & 1 \\
 4th    & +2 & Ability Score Improvement     & 4 + CON         & 2     & 1 \\
 5th    & +3 & Extra Attack, Fast Movement   & 5 + CON         & 3     & 2 \\
 6th    & +3 & Path feature                  & 6 + CON         & 3     & 2 \\
 7th    & +3 & Feral Instinct                & 7 + CON         & 4     & 2 \\
 8th    & +3 & Ability Score Improvement     & 8 + CON         & 4     & 2 \\
 9th    & +4 & Brutal Critical (1 die, 19)   & 9 + CON         & 5     & 2 \\
 10th   & +4 & Path feature, Retaliation     & 10 + CON        & 5     & 3 \\
 11th   & +4 & Relentless                    & 11 + CON        & 6     & 3 \\
 12th   & +4 & Ability Score Improvement     & 12 + CON        & 6     & 3 \\
 13th   & +5 & Brutal Critical (2 dice, 18)  & 13 + CON        & 7     & 3 \\
 14th   & +5 & Path feature                  & 14 + CON        & 7     & 3 \\
 15th   & +5 & Spellbreaker               & 15 + CON        & 8     & 3 \\
 16th   & +5 & Ability Score Improvement     & 16 + CON        & 8     & 3 \\ 
 17th   & +6 & Brutal Critical (3 dice)      & 17 + CON        & 9     & 4 \\
 18th   & +6 & Magebane             & 18 + CON        & 9     & 4 \\
 19th   & +6 & Ability Score Improvement     & 19 + CON        & 10     & 4 \\
 20th   & +6 & Primal Champion               & 20 + CON        & 10     & 4 \\
\end{DndTable}
\twocolumn

\subsection{Exceptional Stamina}
Wardens pull on the vitality of the earth, fueling their stamina. Your stamina at each level is equal to your level + your Constitution modifier.

In addition, you can spend an action to regain stamina equal to twice your Constitution modifier. Once you do so, you cannot use this ability again until you finish a long rest.

\subsection{Furious Blow}
Once per turn when you hit with a melee attack, you can choose to expend 1 or more STA to deal additional damage. If you do so, roll an additional weapon damage die for the first STA you spend and each additional 2 stamina expended and add it to the total rolled. You cannot spend more than 7 Stamina this way in a single attack.

When you deal this extra damage, you gain Focus, which can be spent to fuel other abilities.

Starting at 8th level, when you hit a creature that is Concentrating and deal the additional damage, they have disadvantage on the concentration check.

\subsection{Rage}

You can summon primal strength in dire circumstances. As a bonus action on your turn or as a reaction when you take damage, you can spend 2 STA to enter a heightened state, commonly called Rage.

While raging, you gain the following benefits if you aren’t wearing heavy armor:

\begin{itemize}
\item You have advantage on checks and saving throws involving Strength, Constitution, or Dexterity.
\item You have resistance to bludgeoning, piercing, and slashing damage.
\end{itemize}

Your rage lasts until the end of your next turn. If you take damage or make an attack while raging, your rage is extended for another turn. Additionally, you can sustain your rage for another turn with a bonus action on your turn.

\subsection{Unarmored Defense}

While you are not wearing any armor, your Armor Class equals 10 + your Dexterity modifier + your Constitution modifier. You can use a shield and still gain this benefit.

\subsection{Reckless Attack}

Starting at 2nd level, you can throw aside all concern for defense to attack with fierce desperation. When you make your first attack on your turn, you can decide to attack recklessly. Doing so gives you advantage on melee weapon attack rolls using Strength during this turn, but attack rolls against you have advantage until your next turn.

\subsection{Danger Sense}

At 2nd level, you gain an uncanny sense of when things nearby aren’t as they should be, giving you an edge when you dodge away from danger.

You have advantage on Dexterity saving throws against effects that you can see, such as traps and spells. To gain this benefit, you can’t be blinded, deafened, or incapacitated.

Alternatively, you can expend Focus to gain advantage on a Wisdom saving throw against an effect that would impose the charmed, frightened, or incapacitated condition.

\subsection{Primal Path}

At 3rd level, you choose a path that shapes the nature of your rage. Choose the Juggernaut Path or the Whirlwind Path, both detailed at the end of the class description. Your choice grants you features at 3rd level and again at 6th, 10th, and 14th levels.

\subsection{Ability Score Improvement}

When you reach 4th level, and again at 8th, 12th, 16th, and 19th level, you can increase one ability score of your choice by 2, or you can increase two ability scores of your choice by 1. As normal, you can’t increase an ability score above 20 using this feature.

\subsection{Extra Attack}

Beginning at 5th level, you can attack twice, instead of once, whenever you take the Attack action on your turn.

\subsection{Fast Movement}

Starting at 5th level, your speed increases by 10 feet while you aren’t wearing heavy armor. You can also always jump as if you had a running start.

\subsection{Feral Instinct}

By 7th level, your instincts are so honed that you have advantage on initiative rolls.

Additionally, if you are surprised at the beginning of combat and aren’t incapacitated, you can act normally on your first turn, but only if you enter your rage before doing anything else on that turn.

\subsection{Brutal Critical}

Beginning at 9th level, you can roll one additional weapon damage die when determining the extra damage for a critical hit with a melee attack. You also score a critical hit on a 19 as well as a 20 on the d20.

This increases to two additional dice at 13th level and three additional dice at 17th level. At 13th level you also score a critical hit on an 18.

\subsection{Retaliation}
Starting at 10th level, when you take damage from a creature that is within your reach, you can use your reaction to make a melee weapon attack against that creature. You can spend 2 STA while doing so to attack at advantage.

\subsection{Relentless Rage}

Starting at 11th level, your rage can keep you fighting despite grievous wounds. If you drop to 0 hit points while you’re raging and don’t die outright, you can make a DC 10 Constitution saving throw. If you succeed, you drop to 1 hit point instead.

Each time you use this feature after the first, the DC increases by 5. When you finish a short or long rest, the DC resets to 10.

\subsection{Spell Breaker}
Starting at 15th level, the sheer weight of your primal power inhibits your enemies' magical abilities. While you have Focus, you make saving throws against magical effects at advantage. Additionally, you can expend 3 AET to end any spell effect within 5 feet of you.

\subsection{Magebane}

Beginning at 18th level, if an enemy within your reach attempts to cast a spell or use a magical ability, you can spend your reaction and 2 AET to force them to make a Constitution saving throw against a DC of 8 + your Strength + your proficiency bonus. On a failure, the spell fails and the action is wasted but any resource (aether or daily uses) is not expended.

\subsection{Primal Champion}

At 20th level, you embody the power of the wilds. Your Strength and Constitution scores each increase by +2. Your maximum for those scores is now +7.

\subsection{Warden Paths}

\subsubsection{Juggernaut Path}

Those who embrace the Juggernaut Path embrace the inevitability of their opponents' doom. It may not come fast, but it will certainly come. A Juggernaut, once unleashed, is very hard to stop.

\subsubsection{Punishing Stance}

Starting at 3rd level when you pick this Path, when an enemy within your reach makes an attack against anyone but you or targets any creature other than you with a spell or saving throw, you can expend 2 STA to impose disadvantage on the attack roll or grant the target advantage on the saving throw. Alternatively, you can expend Focus. If you do so, this costs no STA.

\subsubsection{Warding Rage}

Beginning at 6th level, you can’t be charmed or frightened while raging. If you are charmed or frightened when you enter your rage, the effect is suspended for the duration of the rage.

\subsubsection{Inevitability}

Starting at 10th level, the inevitability of your approach cannot be denied by normal means as long as you still have strength. You can expend 5 STA as an action to do one of the following:
\begin{itemize}
\item  Automatically succeed on an otherwise possible attempt to intimidate someone into giving you access or information.
\item  Break any lock or door bar, including those protected by the Arcane Lock incantation. Certain legendary effects may still protect the lock.
\item  Open a naturally-occurring portal or Way that would normally require special circumstances.
\item  Jump to a point you can see within 150 ft of you.
\end{itemize}

\subsubsection{Primal Ward}

Starting at 14th level, your primal presence is such that you can ward out all non-legendary spell effects. As an action and by expending 8 STA, you can invoke a Primal Ward, which takes the form of a swirling barrier of energy in a 10-foot radius around you which remains for one minute or unless you are incapacitated.

Any non-legendary magical effect cast from outside the barrier can't affect creatures or objects within the barrier. Such a spell can target creatures and objects within the barrier, but the spell has no effect on them. Similarly, the area within the barrier is excluded from the areas affected by such spells.

\subsection{Whirlwind Path}
Those who follow the Whirlwind Path delight in carving a bloody swath through their foes and anything that gets in their way. Nimble and unpredictable as a tornado, they leave a trail of wreckage behind them as they dance the deadly dance.

\subsubsection{Dance of Death}
Starting at 3rd level when you choose this Path, you can spend 1 STA as a reaction to ignore any difficult terrain until the end of your turn or break free from a grapple or from restraint due to a creature or spell effect.

\subsubsection{Whirling Fury}
Starting at 3rd level when you choose this Path, you can spend 2 STA to trigger Furious Blow when you hit with an a subsequent attack and have already triggered Furious Blow this turn. When you do so, you cannot spend addtional STA to increase the effect and do not gain Focus.

\subsubsection{Raging Leaps}
Starting at 6th level, while you are raging you gain a fly speed equal to half your speed. You must end your turn in contact with a solid object (including a larger creature if you have succeeded at Climbing on a Larger Creature). Your speed also increases by an additional 10 feet.

This fly speed increases to your full speed at 14th level.

\subsubsection{Unpredictable Nature}
Starting at 10th level, your path has become as unpredictable as a tornado. You are immune to divination effects that would predict your actions, read your thoughts, or determine whether you are lying, as well as the charmed condition while you are raging.

In addition, you can expend Focus to automatically succeed on an otherwise possible attempt to deceive someone about your intent or into giving you information.

\subsubsection{Flash Step}
Starting at 14th level, your speed has increased enough so that it appears you can teleport short distances. When you move on your turn, you can choose to instead teleport to the chosen location as long as you have a clear path to the target location and it is no further than your speed would allow. The clear path to the target does not have to be in a straight line, but you cannot pass through areas too small to squeeze through.

Additionally, you no longer need to end your turns in contact with a solid surface while using Raging Leaps, as you can balance on the air itself.