\section{Cultures}
Choose a culture that fits the origin of your character.  You gain the features listed below. If the ability score increase would take your total increase from lineage and background above +2, choose a different one (subject to the same limitation).

\subsection{Generic (multi-region)}
\subsubsection{Cosmopolitan}

\subparagraph*{Region} Any major city

\subparagraph*{Ability Score Increase} +1 Wisdom

\subparagraph*{Languages} any one common language

\subparagraph*{Proficiencies} Insight, one musical instrument and one game set

\subparagraph*{Coin Counter} You can buy for a 10\% discount and sell for a 10\% markup (ie 55\% of stock price for normal gear or 110\% for treasures and trade goods sold at full value normally).

\subparagraph*{Babyl-dweller} You know the *lesser gift of tongues* incantation and can perform it without a ritual scroll and without expending the component.

\subsubsection{Maritime}

\subparagraph*{Region} Byssia, Serpent Dominion, Southshore

\subparagraph*{Ability Score Increase} +1 Dexterity OR Strength

\subparagraph*{Languages} sena'ka OR Metsae (ihmisi) OR Yonwach (gwerin)

\subparagraph*{Proficiencies} water vehicles, Survival, cartographer's tools

\subparagraph*{Like a Fish} you have a swim speed equal to half your speed.

\subsubsection{Tribal}

\subparagraph*{Region} Fiatch Wood, Kotimaa, Byssian Highlands, Jungle of Fangs, and many others

\subparagraph*{Ability Score Increase} +1 Strength, Constitution, or Wisdom

\subparagraph*{Languages} Ngyon Toi (goblin), Ard-teang (orc), or Metsae (ihmisi)

\subparagraph*{Proficiencies} Survival, woodcarvers and leatherworkers tools

\subparagraph*{Hardy} Once per long rest, you can reduce your Exhaustion level by one when you finish a short rest.

\subsubsection{Religious}

\subparagraph*{Region} Any

\subparagraph*{Ability Score Increase} +1 Wisdom OR +1 Charisma

\subparagraph*{Languages} Old Imperial OR Lucian (infernal/celestial...ish)

\subparagraph*{Proficiencies} Religion, Insight, one artisan's tools of your choice

\subparagraph*{Ascendant Guidance} You learn the *augury* incantation and can cast it without a Ritual Scroll.

\subparagraph*{Ritualist} When you perform an incantation that has a performance time greater than 1 minute, you can reduce it by one step (to a minimum of one minute).

\subsubsection{Rural}
\subparagraph*{Region} Any

\subparagraph*{Ability Score Increase} +1 Strength OR +1 Wisdom

\subparagraph*{Languages} Metsae (ihmisi) OR Ard-teang (orc)

\subparagraph*{Proficiencies} Animal Handling, Survival, Nature

\subparagraph*{Animal Whisperer} You can use your action to attempt to soothe a hostile animal or tame a wild one that isn't hostile to you. Choose a beast or monstrosity you can see that can hear you. Make a Wisdom (Animal Handling) check against a DC of 10 + the creature's CR (rounding up), with disadvantage if it has an Intelligence above 3. On a success, the creature is charmed by you for one hour or until you or your companions take a hostile action against it. If you successfully repeat this check every hour for 4 hours, the creature will follow you for 24 hours and will defend you. It remains an NPC and you cannot command it directly. You can only have one creature tamed at a time, and once you successfully use this ability, you cannot use it targeting any other creature until you finish a long rest.

\subsection{Specific}
\subsubsection{Wall-builder}

\subparagraph*{Region} Sea of Grass

\subparagraph*{Ability Score Increase} + 1 Wisdom

\subparagraph*{Languages} Old Imperial

\subparagraph*{Proficiencies} Religion, Insight, a musical instrument

\subparagraph*{Consensus Builder} You have advantage on any Charisma (Persuasion) check made to improve someone's attitude toward you.

\subparagraph*{Children of the Congregation} You learn a cantrip of your choice from the Priest list. Wisdom is the casting ability for this cantrip.

\subsubsection{Wyrmhold Clan}

\subparagraph*{Region} Nocthian Caldera, Safehold, Byarmarsh, Kiren Mountains, Fiach Woods

\subparagraph*{Ability Score Increase} +1 Strength

\subparagraph*{Languages} Draconic OR Ard-teang OR Ngyon Toi

\subparagraph*{Proficiencies} One martial weapon, History, Survival

\subparagraph*{Trained for War} When you spend a hit die to heal, if the die result is below half of its maximum, the hit die is not expended.

\subsubsection{Child of Night}
\subparagraph*{Region} Nocthian Caldera, Byssia

\subparagraph*{Ability Score Increase} +1 Wisdom

\subparagraph*{Languages} Metsae

\subparagraph*{Proficiencies} Animal Handling, Nature, Woodcarver's OR Carpenter's Tools

\subparagraph*{Delver into Mysteries} The culture of Byssia is big on secrets--keeping them and discovering them. You have advantage on any Wisdom (Perception), Intelligence (Investigation), or Wisdom (Insight) check made to determine a hidden location, object, or motive.

\subsubsection{Fang-kin}
\subparagraph*{Region} Jungle of Fangs

\subparagraph*{Ability Score Increase} +1 Dexterity OR +1 Charisma

\subparagraph*{Languages} Iath Neidr OR Old Imperial

\subparagraph*{Proficiencies} Performance, Deception, Disguise Kit

\subparagraph*{Serpentine Grace} The fang-kin prize grace and dancing. Most who grow up here are more flexible than those of other cultures, no matter their lineage. You can squeeze through spaces one size smaller than yourself without penalty. In addition, you can move through enemies spaces as long as they are one size larger or smaller than yourself instead of two, although the spaces are still difficult terrain. 

\subsubsection{Uulani}
\subparagraph*{Region} Uulan Confederacy, Shinevog

\subparagraph*{Ability Score Increase} +1 Constitution

\subparagraph*{Languages} Tumni

\subparagraph*{Proficiencies} History, two crafters tools of your choice

\subparagraph*{Famous Artisans} The Uulani are known far and wide for their crafting prowess. All crafting costs half as much time and money. In addition, you start with the schema for one Common item of your choice. At level 6, you discover a schema of your choice for an Uncommon item.