\section{Cultures}\label{sec:cultures}
Culture represents all the socio-political aspects of your upbringing. In some cases, a culture and a nation of origin are coterminous; in many cases the culture is extends between nations and may only be a minority in any of them. People of many lineages have similar (enough for game purposes) cultures, so you can pick any culture regardless of your lineage. Different lineages tend to express shared cultures slightly differently, but that's between you and your GM.

Choose a culture that fits the origin of your character.  You gain the features listed below. If the ability score increase is the same as the ability score increase from your lineage, choose a different one that does not overlap. You cannot have more than a +1 to any individual ability score from your combined lineage and culture.

\begin{DndComment}{Gender and Sexuality}
	Many of the lineages and cultures of the Federated Nations have "traditional" attitudes toward gender roles. But adventurers are known to be weird. So they get a pass. You can play any gender, including non-binary. Be warned that some NPCs will react differently depending on their perceptions of your gender and how it meets the culture's expectations.

	A few specifics:
	\begin{itemize}
		\item The dwarves of Fuar Uulan are very strictly heterosexual and monogamous. All regular adults are married and stay married to one person. They react badly to adultery or divorce, and look at unmarried/non-traditional dwarves with confusion and pity.
		\item The high-ranking gwerin tend to do arranged marriages--adults only gain stature in the community if they've provided at least one child as part of a contract marriage. Some of the higher-society houses are still quite male-dominated, but the lower ranks and the more progressive houses are egalitarian. In either case, no one cares who (if anyone) you are attracted to as long as you provide your requisite children to the house.
		\item The ihmisi, especially the more tribal folks, have very relaxed gender roles. Anyone can fulfil any societal role. Similarly, goblins have virtually no concept of gender and very little grasp of individuality within the tribes.
		\item The orcs of Wyrmhold tend to hold their women in regard--women tend to be the "glue" that holds the clans and tribes together. They don't care who you are attracted to, although refusing to provide children is considered weird.
		\item Halflings are generally deeply matriarchal, treating men as some combination of decorative, protected, and/or weak. Almost obnoxiously so. Female sexuality is unconstrained, but men are expected to be tied to their wives. Most male halfling adventurers break this expectation, often fleeing from forced marraiges.
		\item The the Tuura Adam tend to place women in the gate-keeping role for relationships. Men are expected to prove themselves worthy. Non-binary or "non-traditional" gender roles are fully accepted...as long as you're good at whatever you do. Incompetence is the only thing not respected. 
	\end{itemize} 
\end{DndComment}

\subsection{Generic (multi-region)}
\subsubsection{Cosmopolitan}
You grew up in one of the major cities dotting Noefra, surrounded by people of many lineages and cultures. The constant rush of commerce, the babel of different languages, all these shaped you.

\subparagraph*{Region} Any major city

\subparagraph*{Ability Score Increase} +1 Wisdom

\subparagraph*{Languages} any one common language

\subparagraph*{Proficiencies} Insight, one musical instrument and one game set

\subparagraph*{Coin Counter} You gain the \nameref{st:haggler} skill trick even if you wouldn't otherwise qualify.

\subparagraph*{Babyl-dweller} You know the \textit{lesser gift of tongues} incantation and can perform it without a ritual scroll and without expending the component.

\subsubsection{Maritime}
You grew up by of the of the major bodies of water (inland or oceanic) and spent your early life in and around the water, even if you didn't sail as a profession.

\subparagraph*{Region} Byssia, Serpent Dominion, Southshore

\subparagraph*{Ability Score Increase} +1 Dexterity OR  +1 Strength

\subparagraph*{Languages} sena'ka OR Metsae (ihmisi) OR Yonwach (gwerin)

\subparagraph*{Proficiencies} water vehicles, Survival, cartographer's tools

\subparagraph*{Like a Fish} you have a swim speed equal to half your speed.

\subsubsection{Tribal}
You grew up in a close-knit family, tribe, or clan, whether settled or migratory. Unlike most rural dwellers, your clan came first, rather than any nation or regional allegiance. Survival in the wilderness was often a top concern.

\subparagraph*{Region} Fiatch Wood, Kotimaa, Byssian Highlands, Jungle of Fangs, and many others

\subparagraph*{Ability Score Increase} +1 Strength OR +1 Constitution OR +1 Wisdom

\subparagraph*{Languages} Ngyon Toi (goblin) OR Ard-teang (orc) OR Metsae (ihmisi)

\subparagraph*{Proficiencies} Survival, woodcarvers and leatherworkers tools

\subparagraph*{Hardy} Once per long rest, you can reduce your Exhaustion level by one when you finish a short rest.

\subsubsection{Religious}
Your upbringing was full of religion, faith, ceremony, and worship. Whether of one of the gods or some other Ascendant, this pattern of worship dominated your early life even though you were not necessarily a priest or acolyte at any temple.

\subparagraph*{Region} Any

\subparagraph*{Ability Score Increase} +1 Wisdom OR +1 Charisma

\subparagraph*{Languages} Old Imperial OR Lucian (infernal/celestial...ish)

\subparagraph*{Proficiencies} Religion, Insight, one artisan's tools of your choice

\subparagraph*{Ascendant Guidance} You learn the \nameref{inc:augury} incantation and can cast it without a Ritual Scroll.

\subparagraph*{Ritualist} When you perform an incantation that has a performance time greater than 1 minute, you can reduce it by one step (to a minimum of one minute).

\subsubsection{Rural}
You grew up part of a "civilized" nation, but like the majority of the population, you grew up outside the major cities and towns. Farming, ranching, keeping sheep or cows, and hunting made up most of your days. 

\subparagraph*{Region} Any

\subparagraph*{Ability Score Increase} +1 Strength OR +1 Wisdom

\subparagraph*{Languages} Metsae (ihmisi) OR Ard-teang (orc)

\subparagraph*{Proficiencies} Animal Handling, Survival, Nature

\subparagraph*{Animal Whisperer} You can use your action to attempt to soothe a hostile animal or tame a wild one that isn't hostile to you. Choose a beast or monstrosity you can see that can hear you. Make a Wisdom (Animal Handling) check against a DC of 10 + the creature's CR (rounding up), with disadvantage if it has an Intelligence above 3. On a success, the creature is charmed by you for one hour or until you or your companions take a hostile action against it. If you successfully repeat this check every hour for 4 hours, the creature will follow you for 24 hours without any further checks and will defend you. It remains an NPC and you cannot command it directly. You can only have one creature tamed at a time, and once you successfully use this ability, you cannot use it targeting any other creature until you finish a long rest.

\subsection{Specific}
These cultures tie you to a specific region of the world.
\subsubsection{Wall-builder}
The wall-builders are the dominant culture of those whose ancestors fled the Cataclysm into the Sea of Grass in central-western Noefra. Their name comes from their historical pattern of always building a wall around every settlement, no matter how minor. Even isolated hamlets and farms have at least a palisade or berm. Highly religious and consensus-driven, they believe in community and finding your expected role in society without making waves.

\subparagraph*{Region} Sea of Grass

\subparagraph*{Ability Score Increase} + 1 Wisdom

\subparagraph*{Languages} Old Imperial

\subparagraph*{Proficiencies} Religion, Insight, a musical instrument

\subparagraph*{Consensus Builder} You have advantage on any Charisma (Persuasion) check made to improve someone's attitude toward you.

\subparagraph*{Children of the Congregation} You learn a cantrip of your choice from the Priest list. Wisdom is the casting ability for this cantrip.

\subsubsection{Wyrmhold Clan}
Wyrmhold, the nation of orcs, dragonborn, and goblins on the eastern edge of the Nocthian Caldera, is a very martial culture while also being clannish (although the clan-based structure is fading in more recent years). Everyone is trained in wielding a weapon and strength is prized. The three lineages each play their own parts and leave the internal affairs of the others alone, which is both a strength and a weakness. 

\subparagraph*{Region} Nocthian Caldera, Safehold, Byarmarsh, Kiren Mountains, Fiach Woods

\subparagraph*{Ability Score Increase} +1 Strength

\subparagraph*{Languages} Draconic OR Ard-teang OR Ngyon Toi

\subparagraph*{Proficiencies} One martial weapon, History, Survival

\subparagraph*{Trained for War} When you spend a hit die to heal, if the die result is below half of its maximum, the hit die is not expended.

\subsubsection{Child of Night}
Byssia, located to the south of the Nocthian Caldera, is a land culturally obsessed with mysteries. Keeping them from others and finding them out. They also are notable for not being conventionally religious--they worship either the nature spirits (kami) or the Ascendant Nocthis, patroness of the moons and of mystery. Their society is rural and decentralized, with much of it making conscious efforts to live in harmony with nature.
\subparagraph*{Region} Nocthian Caldera, Byssia

\subparagraph*{Ability Score Increase} +1 Wisdom

\subparagraph*{Languages} Metsae

\subparagraph*{Proficiencies} Animal Handling, Nature, Woodcarver's OR Carpenter's Tools

\subparagraph*{Delver into Mysteries} The culture of Byssia is big on secrets--keeping them and discovering them. You have advantage on any Wisdom (Perception), Intelligence (Investigation), or Wisdom (Insight) check made to determine a hidden location, object, or motive.

\subsubsection{Fang-kin}
The Jungle of Fangs is home to several nations who share a common culture, history, and religion. Most of the natives of the region are descendents of those who served the Old Masters (who disappeared about 400 years ago) and were changed by them. They dominantly worship the Queen Ascendant, a recently-ascended (about 45 years ago) "eternal queen" figure, doing so via dance and music. They prize the arts and commonly revere snakes. The color blue is reserved for nobles by long-standing tradition. Culturally, they are very caste oriented, with strong boundaries between nobles, tradesmen, and commoners.

\subparagraph*{Region} Jungle of Fangs

\subparagraph*{Ability Score Increase} +1 Dexterity OR +1 Charisma

\subparagraph*{Languages} Iath Neidr OR Old Imperial

\subparagraph*{Proficiencies} Performance, Deception, Disguise Kit

\subparagraph*{Serpentine Grace} The fang-kin prize grace and dancing. Most who grow up here are more flexible than those of other cultures, no matter their lineage. You can squeeze through spaces one size smaller than yourself without penalty. In addition, you can move through enemies spaces as long as they are one size larger or smaller than yourself instead of two, although the spaces are still difficult terrain. 

\subsubsection{Uulani}
The Uulani are mostly dwarves, but some humans and halflings who have adopted their ways. Unlike the Wyrmhold clans who prioritize \textit{family}, the uulani prioritize \textit{ideology}. Each clan is a shared mind-set, and switching clans as one grows up is not unheard of or unwelcome. Like the wall-builders, they are consensus-driven; unlike them they have no problems making waves and questioning things. They prize hard work and good workmanship in everything--idleness is one of their sins. Not incredibly religious, they revere specific ancestor figures (called the Paragons) as examples of what a good Uulani should be and pay homage to the God of Mountains and Silent Endurance, Korokonolkolm.

\subparagraph*{Region} Uulan Confederacy, Shinevog

\subparagraph*{Ability Score Increase} +1 Constitution

\subparagraph*{Languages} Tumni

\subparagraph*{Proficiencies} History, two crafters tools of your choice

\subparagraph*{Famous Artisans} The Uulani are known far and wide for their crafting prowess. All crafting costs half as much time and money. In addition, you start with the schema for one Common item of your choice. At level 6, you discover a schema of your choice for an Uncommon item.