\subsection{Spells (A)}

\DndSpellHeader{Acid Arrow \label{spell:acid-arrow}}
{3 AET, 1T, Acid, Attack}
{1 action}
{90 feet}
{V, S, M (powdered rhubarb leaf and an adder's stomach)}
{Instantaneous}

A shimmering green arrow streaks toward a target within range and bursts in a spray of acid. Make a ranged spell attack against the target. On a hit, the target takes 4d6 acid damage immediately and 2d6 acid damage at the end of its next turn. On a miss, the arrow splashes the target with acid for half as much of the initial damage and no damage at the end of its next turn.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage (both intial and subsequent) increases by 1d6 for every 2 AET you spend above 3.

\DndSpellHeader{Acid Burst \label{spell:acid-burst}}
{Cantrip, 1T, Acid, CON}
{1 action}
{60 feet}
{V, S}
{Instantaneous}

Acidic liquid envelops a target you can see within range. The target must make a Constitution saving throw. On a failed save, the target takes 1d8 damage immediately and half as much at the beginning of their next turn.

This spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Aid \label{spell:aid}}
{3 AET, 3T, Defense}
{1 action}
{30 feet}
{V, S, M (a tiny strip of white cloth)}
{8 hours}

Your spell bolsters your allies with toughness and resolve. Choose up to three creatures within range. Each target's hit point maximum and current hit points increase by 5 for the duration.

\subparagraph*{Overcast} When you cast this spell using 4 or more AET, a target's hit points increase by an additional 5 for each 2 AET above 3.


\DndSpellHeader{Alter Self \label{spell:alter-self}}
{3 AET, Self, Shapechanging}
{1 action}
{Self}
{V, S}
{Concentration, up to 1 hour}

You assume a different form. When you cast the spell, choose one of the following options, the effects of which last for the duration of the spell. While the spell lasts, you can end one option as an action to gain the benefits of a different one.

\subparagraph*{Aquatic Adaptation} You adapt your body to an aquatic environment, sprouting gills and growing webbing between your fingers. You can breathe underwater and gain a swimming speed equal to your walking speed.

\subparagraph*{Change Appearance} You transform your appearance. You decide what you look like, including your height, weight, facial features, sound of your voice, hair length, coloration, and distinguishing characteristics, if any. You can make yourself appear as a member of another race, though none of your statistics change. You also can't appear as a creature of a different size than you, and your basic shape stays the same; if you're bipedal, you can't use this spell to become quadrupedal, for instance. At any time for the duration of the spell, you can use your action to change your appearance in this way again.

\subparagraph*{Natural Weapons} You grow claws, fangs, spines, horns, or a different natural weapon of your choice. Your unarmed strikes deal 1d6 bludgeoning, piercing, or slashing damage, as appropriate to the natural weapon you chose, and you are proficient with your unarmed strikes. Finally, the natural weapon is magic and you have a +1 bonus to the attack and damage rolls you make using it.

\subparagraph*{Overcast} For every 3 AET above the base cost, the duration extends by 1 hour. If you spend at least 12 AET, it no longer requires concentration.


\DndSpellHeader{Antilife Shell \label{spell:antilife-shell}}
{10 AET, CB2/aura, Defensive}
{1 action}
{Self (10-foot radius)}
{V, S}
{Concentration, up to 1 hour}

A shimmering barrier extends out from you in a 10 foot radius and moves with you, remaining centered on you and hedging out creatures other than undead and constructs. The barrier lasts for the duration.

The barrier prevents an affected creature from passing or reaching through. An affected creature can cast spells or make attacks with ranged or reach weapons through the barrier.

If you move so that an affected creature is forced to pass through the barrier, the spell ends.

\DndSpellHeader{Arcane Binding\label{spell:arcane-binding}}
{4 AET}
{1 action}
{60 feet}
{V, S, M (a small, straight piece of iron)}
{Concentration, up to 1 minute}

Choose a creature that you can see within range. The target must succeed on a Strength saving throw or be restrained for the duration as glowing aetheric chains wrap around it and sap its strength. If the target fails by 5 or more, they count as being paralyzed except that they can take any purely verbal or mental action. At the end of each of its turns, the target can make another Strength saving throw. On a success, the spell ends on the target.

\subparagraph*{Overcast} When you cast this spell using 6 or more AET, you can target a second creature within range. When you cast this spell using 8 or more AET, the targets are paralyzed (with the same exception) if they fail the saving throw at all, rather than by 5 or more.


\DndSpellHeader{Arcane Eye \label{spell:arcane-eye}}
{8 AET, scouting}
{1 action}
{30 feet}
{V, S, M (a bit of bat fur)}
{Concentration, up to 1 hour}

You create an invisible, magical eye within range that hovers in the air for the duration.

You mentally receive visual information from the eye, which has normal vision and darkvision out to 30 feet. The eye can look in every direction.

As an action, you can move the eye up to 30 feet in any direction. There is no limit to how far away from you the eye can move, but it can't enter another plane of existence. A solid barrier blocks the eye's movement, but the eye can pass through an opening as small as 1 inch in diameter.

\subsection{Spells (B)}

\DndSpellHeader{Bane \label{spell:bane}}
{2 AET, 3T, Debuff}
{1 action}
{30 feet}
{V, S, M (a drop of blood)}
{Concentration, up to 1 minute. Save ends.}

Up to three creatures of your choice that you can see within range must make Charisma saving throws. Targets that fail their saving throw have disadvantage on attack rolls and ability checks for the duration. They can attempt the saving throw again at the end of each of their turns, ending the effect on a success.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can target one additional creature for every 2 additional AET you spend.


\DndSpellHeader{Banishment \label{spell:banishment}}
{8 AET, 1T, CHA, action loss}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 1 minute. Save ends.}

You attempt to send one creature that you can see within range to another plane of existence. The target must succeed on a Charisma saving throw or be banished.

If the target is native to the plane of existence you're on, you banish the target to a harmless demiplane. While there, the target is incapacitated. The target remains there until the spell ends, at which point the target reappears in the space it left or in the nearest unoccupied space if that space is occupied. Native targets can make the saving throw again at the end of each of their turns, ending the effect and returning back on a success.

If the target is native to a different plane of existence than the one you're on, the target is banished with a faint popping noise, returning to its home plane. If the spell ends before 1 minute has passed, the target reappears in the space it left or in the nearest unoccupied space if that space is occupied. Otherwise, the target doesn't return. Non-native targets do not get additional saving throws past the initial one.

\subparagraph*{Overcast} When you cast this spell using more than 8 AET, you can target one additional creature for every 3 additional AET you spend.


\DndSpellHeader{Barkskin \label{spell:barkskin}}
{3 AET, S, Defense}
{1 action}
{Touch}
{V, S, M (a handful of oak bark)}
{1 hour}

You touch a willing creature. Until the spell ends, the target's skin has a rough, bark-like appearance, and the target's AC can't be less than 16, regardless of what kind of armor it is wearing.

In addition, the target gains temporary hit points equal to your spellcasting modifier at the beginning of each of their turns. These temporary hit points fade at the end of the duration.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the target's minimum AC increases by 1 for ever 3 AET you spend, up to a maximum AC of 20.


\DndSpellHeader{Beacon of Hope \label{spell:beacon-of-hope}}
{5 AET, CB, buff}
{1 action}
{30 feet}
{V, S}
{Concentration, up to 1 minute}

This spell bestows hope and vitality. Choose any number of creatures within range. For the duration, each target has advantage on Wisdom saving throws and death saving throws, and regains the maximum number of hit points possible from any healing.


\DndSpellHeader{Bestow Curse \label{spell:bestow-curse}}
{5 AET, ST, WIS, Debuff}
{1 action}
{Touch}
{V, S}
{Concentration, up to 1 minute}

You touch a creature, and that creature must succeed on a Wisdom saving throw or become cursed for the duration of the spell. When you cast this spell, choose the nature of the curse from the following options:

\begin{itemize}
\item Choose one ability score. While cursed, the target has disadvantage on ability checks and saving throws made with that ability score.
\item While cursed, the target has disadvantage on attack rolls against you.
\item While cursed, the target must make a Wisdom saving throw at the start of each of its turns. If it fails, it wastes its action that turn doing nothing.
\item While the target is cursed, your attacks and spells deal an extra 1d8 necrotic damage to the target.
\end{itemize}

A \textit{remove curse} spell ends this effect. At the GM's option, you may choose an alternative curse effect, but it should be no more powerful than those described above. The GM has final say on such a curse's effect.

\subparagraph*{Overcast} If you cast this spell using 8 AET, the duration is concentration, up to 10 minutes. If you use 12 AET,the duration is 8 hours. If you use 16 AET, the duration is 24 hours. Using 12 or more AET grants a duration that doesn't require concentration.


\DndSpellHeader{Black Tentacles\label{spell:black-tentacles}}
{8 AET, zone, DEX, movement + damage}
{1 action}
{90 feet}
{V, S, M (a piece of tentacle from a giant octopus or a giant squid)}
{Concentration, up to 1 minute}

Squirming, ebony tentacles fill a 20-foot square on ground that you can see within range. For the duration, these tentacles turn the ground in the area into difficult terrain.

When a creature enters the affected area for the first time on a turn or starts its turn there, the creature must succeed on a Strength saving throw or take 3d6 bludgeoning damage and be restrained by the tentacles until the spell ends. A creature that starts its turn in the area and is already restrained by the tentacles takes 3d6 bludgeoning damage.

A creature restrained by the tentacles can use its action to make a Strength (Athletics) check against your spell save DC. On a success, it frees itself.
\subparagraph*{Overcast} When you use more than 8 AET to cast this spell, the damage dealt increases by 1d6 for every two AET you spend up to 12 (dealing 5d6).

\DndSpellHeader{Blade Barrier\label{spell:blade-barrier}}
{15 AET}
{1 action}
{90 feet}
{V, S}
{Concentration, up to 10 minutes}

You create a vertical wall of whirling, razor-sharp blades made of magical energy. The wall appears within range and lasts for the duration. You can make a straight wall up to 100 feet long, 20 feet high, and 5 feet thick, or a ringed wall up to 60 feet in diameter, 20 feet high, and 5 feet thick. The wall provides three-quarters cover to creatures behind it, and its space is difficult terrain.

When a creature enters the wall’s area for the first time on a turn or starts its turn there, the creature must make a Dexterity saving throw. On a failed save, the creature takes 6d10 slashing damage. On a successful save, the creature takes half as much damage.


\DndSpellHeader{Bless\label{spell:bless}}
{2 AET}
{1 action}
{30 feet}
{V, S, M (a sprinkling of holy water)}
{Concentration, up to 1 minute}

You bless up to three creatures of your choice within range. Whenever a target makes an attack roll or a saving throw before the spell ends, the target can roll a d4 and add the number rolled to the attack roll or saving throw.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can target one additional creature for every 2 AET you spend.


\DndSpellHeader{Blight\label{spell:blight}}
{6 AET}
{1 action}
{30 feet}
{V, S}
{Instantaneous}

Necromantic energy washes over a creature of your choice that you can see within range, draining moisture and vitality from it. The target must make a Constitution saving throw. The target takes 10d8 necrotic damage on a failed save, or half as much damage on a successful one. This spell has no effect on undead or constructs.

If you target a plant creature or a magical plant, it makes the saving throw with disadvantage, and the spell deals maximum damage to it.

If you target a nonmagical plant that isn't a creature, such as a tree or shrub, it doesn't make a saving throw; it simply withers and dies.

\subparagraph*{Overcast} When you cast this spell using more than 8 AET, the damage increases by 1d8 for every AET you spend up to a maximum of 12 AET (12d8).


\DndSpellHeader{Blindness/Deafness\label{spell:blindness-deafness}}
{3 AET}
{1 action}
{30 feet}
{V}
{1 minute, save ends}

You can blind or deafen a foe. Choose one creature that you can see within range to make a Constitution saving throw. If it fails, the target is either blinded or deafened (your choice) for the duration. At the end of each of its turns, the target can make a Constitution saving throw. On a success, the spell ends.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, you can target one additional creature for every two additional AET spent.


\DndSpellHeader{Blink\label{spell:blink}}
{5 AET}
{1 action}
{Self}
{V, S}
{1 minute}

Roll a d20 at the end of each of your turns for the duration of the spell. On a roll of 11 or higher, you vanish from your current plane of existence and appear in the Border Shadow (the spell fails and the casting is wasted if you were already on that plane). At the start of your next turn, and when the spell ends if you are on the Border Shadow, you return to an unoccupied space of your choice that you can see within 10 feet of the space you vanished from. If no unoccupied space is available within that range, you appear in the nearest unoccupied space (chosen at random if more than one space is equally near). You can dismiss this spell as an action.

While on the Border Shadow, you can see and hear the plane you originated from, which is cast in shades of gray, and you can't see anything there more than 60 feet away. You can only affect and be affected by other creatures on the Border Shadow. Creatures that aren't there can't perceive you or interact with you, unless they have the ability to do so.


\DndSpellHeader{Blur\label{spell:blur}}
{3 AET}
{1 action}
{Self}
{V}
{Concentration, up to 1 minute}

Your body becomes blurred, shifting and wavering to all who can see you. For the duration, any creature has disadvantage on attack rolls against you. An attacker is immune to this effect if it doesn't rely on sight, as with blindsight, or can see through illusions, as with truesight.


\DndSpellHeader{Burning Hands\label{spell:burning-hands}}
{2 AET}
{1 action}
{Self (15-foot cone)}
{V, S}
{Instantaneous}

A thin sheet of flames shoots forth from your outstretched fingertips. Each creature in a 15-foot cone must make a Dexterity saving throw. A creature takes 3d6 fire damage on a failed save, or half as much damage on a successful one.

The fire ignites any flammable objects in the area that aren't being worn or carried.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d6 for every 2 additional AET spent.

\subsection{Spells (C)}


\DndSpellHeader{Call Lightning\label{spell:call-lightning}}
{6 AET}
{1 action}
{120 feet}
{V, S}
{Concentration, up to 10 minutes}

A storm cloud appears in the shape of a cylinder that is 10 feet tall with a 60-foot radius, centered on a point you can see 100 feet directly above you. The spell fails if you can't see a point in the air where the storm cloud could appear (for example, if you are in a room that can't accommodate the cloud).

When you cast the spell, choose a point you can see within range. A bolt of lightning flashes down from the cloud to that point. Each creature within 5 feet of that point must make a Dexterity saving throw. A creature takes 3d10 lightning damage on a failed save, or half as much damage on a successful one. On each of your turns until the spell ends, you can use your action to call down lightning in this way again, targeting the same point or a different one.

If you are outdoors in stormy conditions when you cast this spell, the spell gives you control over the existing storm instead of creating a new one. Under such conditions, the spell's damage increases by 1d10.

\subparagraph*{Overcast} When you cast this spell using more than 4 AET, the damage increases by 1d10 for every 3 additional AET spent.


\DndSpellHeader{Calm Emotions\label{spell:calm-emotions}}
{3 AET}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 1 minute}

You attempt to suppress strong emotions in a group of people. Each humanoid in a 20-foot-radius sphere centered on a point you choose within range must make a Charisma saving throw; a creature can choose to fail this saving throw if it wishes. If a creature fails its saving throw, choose one of the following two effects.

You can suppress any effect causing a target to be charmed or frightened. When this spell ends, any suppressed effect resumes, provided that its duration has not expired in the meantime.

Alternatively, you can make a target indifferent about creatures of your choice that it is hostile toward. This indifference ends if the target is attacked or harmed by a spell or if it witnesses any of its friends being harmed. When the spell ends, the creature becomes hostile again, unless the GM rules otherwise.

\DndSpellHeader{Chain Lightning\label{spell:chain-lightning}}
{13 AET}
{1 action}
{150 feet}
{V, S, M (a bit of fur; a piece of amber, glass, or a crystal rod; and three silver pins)}
{Instantaneous}

You create a bolt of lightning that arcs toward a target of your choice that you can see within range. Three bolts then leap from that target to as many as three other targets, each of which must be within 30 feet of the first target. A target can be a creature or an object and can be targeted by only one of the bolts.

A target must make a Dexterity saving throw. The target takes 10d8 lightning damage on a failed save, or half as much damage on a successful one.

\subparagraph*{Overcast} When you cast this spell using more than 13 AET, one additional bolt leaps from the first target to another target for every 3 additional AET spent.


\DndSpellHeader{Charm Person\label{spell:charm-person}}
{2 AET}
{1 action}
{30 feet}
{V, S}
{1 hour}

You attempt to charm a humanoid you can see within range. It must make a Charisma saving throw, and does so with advantage if you or your companions are fighting it. If it fails the saving throw, it is charmed by you until the spell ends or until you or your companions do anything harmful to it. The charmed creature's attitude toward you shifts to Friendly for the duration.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can target one additional creature for every 2 additional AET spent. The creatures must be within 30 feet of each other when you target them.


\DndSpellHeader{Clairvoyance\label{spell:clairvoyance}}
{5 AET}
{10 minutes}
{1 mile}
{V, S, M (a focus worth at least 100 gp, either a jeweled horn for hearing or a glass eye for seeing)}
{Concentration, up to 10 minutes}

You create an invisible sensor within range in a location familiar to you (a place you have visited or seen before) or in an obvious location that is unfamiliar to you (such as behind a door, around a corner, or in a grove of trees). The sensor remains in place for the duration. A creature that can see the sensor (such as a creature benefiting from \textit{see invisibility} or who has truesight) sees a luminous, intangible orb about the size of your fist. The sensor has an AC of 10 and is destroyed by any damage taken.

When you cast the spell, you choose seeing or hearing. You can use the chosen sense through the sensor as if you were in its space. As your action, you can switch between seeing and hearing.


\DndSpellHeader{Cloudkill\label{spell:cloudkill}}
{14 AET}
{1 action}
{120 feet}
{V, S}
{Concentration, up to 10 minutes}

You create a 20-foot-radius sphere of poisonous, yellow-green fog centered on a point you choose within range. The fog spreads around corners. It lasts for the duration or until strong wind disperses the fog, ending the spell. Its area is heavily obscured.

When a creature enters the spell's area for the first time on a turn or starts its turn there, that creature must make a Constitution saving throw. The creature takes 5d8 poison damage on a failed save, or half as much damage on a successful one. Creatures are affected even if they hold their breath or don't need to breathe.

The fog moves 10 feet away from you at the start of each of your turns, rolling along the surface of the ground. The vapors, being heavier than air, sink to the lowest level of the land, even pouring down openings.

\subparagraph*{Overcast} When you cast this spell using more than 12 AET , the damage increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Color Spray\label{spell:color-spray}}
{2 AET}
{1 action}
{Self (15-foot cone)}
{V, S, M (a pinch of powder or sand that is colored red, yellow, and blue)}
{1 round}

A dazzling array of flashing, colored light springs from your hand. All creatures that can see you in a 15' cone must make a Constitution saving throw. On a failed save, they are blinded until the end of your next turn. 

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the duration increases by one round for every 2 additional AET spent.


\DndSpellHeader{Command\label{spell:command}}
{2 AET}
{1 action}
{60 feet}
{V}
{1 round}

You speak a one-word command to a creature you can see within range. The target must succeed on a Wisdom saving throw or follow the command on its next turn. The spell has no effect if the target is undead, if it doesn't understand your language, or if your command is directly harmful to it.

Some typical commands and their effects follow. You might issue a command other than one described here. If you do so, the GM determines how the target behaves. If the target can't follow your command, the spell ends.

\subparagraph*{Approach} The target moves toward you by the shortest and most direct route, ending its turn if it moves within 5 feet of you.

\subparagraph*{Drop} The target drops whatever it is holding and then ends its turn.

\subparagraph*{Flee} The target spends its turn moving away from you by the fastest available means.

\subparagraph*{Grovel} The target falls prone and then ends its turn.

\subparagraph*{Halt} The target doesn't move and takes no actions.

A flying creature stays aloft, provided that it is able to do so. If it must move to stay aloft, it flies the minimum distance needed to remain in the air.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you can affect one additional creature for every 2 additional AET spent. The creatures must be within 30 feet of each other when you target them.


\DndSpellHeader{Cone of Cold\label{spell:cone-of-cold}}
{10 AET}
{1 action}
{Self (60-foot cone)}
{V, S, M (a small crystal or glass cone)}
{Instantaneous}

A blast of cold air erupts from your hands. Each creature in a 60-foot cone must make a Constitution saving throw. A creature takes 8d8 cold damage on a failed save, or half as much damage on a successful one. Targets that fail their save are stunned until the end of the caster's next turn.

A creature killed by this spell becomes a frozen statue until it thaws.

\subparagraph*{Overcast} When you cast this spell using more than 10 AET, the damage increases by 1d8 for every 2 additional AET spent.


\DndSpellHeader{Confusion\label{spell:confusion}}
{8 AET}
{1 action}
{90 feet}
{V, S, M (three nut shells)}
{Concentration, up to 1 minute}

This spell assaults and twists creatures' minds, spawning delusions and provoking uncontrolled action. Each creature in a 10-foot-radius sphere centered on a point you choose within range must succeed on a Wisdom saving throw when you cast this spell or be affected by it.

An affected target can't take reactions and must roll a d10 at the start of each of its turns to determine its behavior for that turn.

\begin{DndTable}{XX}
    \textbf{d10} & \textbf{Behavior} \\
    1 & The creature uses all its movement to move in a random direction. To determine the direction, roll a d8 and assign a direction to each die face. The creature doesn't take an action this turn. \\
    2-6 & The creature doesn't move or take actions this turn. \\
    7-8 & The creature uses its action to make a melee attack against a randomly determined creature within its reach. If there is no creature within its reach, the creature does nothing this turn. \\
    9-10 & The creature can act and move normally.
\end{DndTable}

At the end of each of its turns, an affected target can make a Wisdom saving throw. If it succeeds, this effect ends for that target.

\subparagraph*{Overcast} When you cast this spell using more than 8 AET, the radius of the sphere increases by 5 feet for every 3 additional AET spent.


\DndSpellHeader{Conjure Animals\label{spell:conjure-animals}}
{5 AET}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 1 hour}

You summon fey spirits that take the form of beasts and appear in unoccupied spaces that you can see within range. Choose one of the following options for what appears:

\begin{itemize}
\item One cave bear, giant boar, giant constrictor snake, hunter shark, or saber-toothed tiger.
\item Two dire wolves, brown bears, giant spiders, giant octopuses, or giant toads.
\end{itemize}

The summoned creatures are friendly to you and your companions. Roll initiative for the summoned creatures as a group, which has its own turns. They obey any verbal commands that you issue to them (no action required by you). If you don't issue any commands to them, they defend themselves from hostile creatures, but otherwise take no actions.

The GM has the creatures' statistics.

\subparagraph*{Overcast} When you cast this spell using more than 5 AET, you can choose higher CR monsters:
\begin{itemize}
\item 8 AET spent: 1 giant scorpion or killer whale or 2 of the base solo options.
\item 11 AET spent: 1 elephant or 2 giant scorpions or killer whales.
\end{itemize}


\DndSpellHeader{Conjure Elemental\label{spell:conjure-elemental}}
{12 AET}
{1 minute}
{90 feet}
{V, S, M (burning incense for air, soft clay for earth, sulfur and phosphorus for fire, or water and sand for water)}
{Concentration, up to 1 hour}

You call forth an elemental servant. Choose an area of air, earth, fire, or water that fills a 10-foot cube within range. An elemental of challenge rating 5 or lower appropriate to the area you chose appears in an unoccupied space within 10 feet of it. For example, a fire elemental emerges from a bonfire, and an earth elemental rises up from the ground. The elemental disappears when it drops to 0 hit points or when the spell ends.

The elemental is friendly to you and your companions for the duration. Roll initiative for the elemental, which has its own turns. It obeys any verbal commands that you issue to it (no action required by you). If you don't issue any commands to the elemental, it defends itself from hostile creatures but otherwise takes no actions.

If your concentration is broken, the elemental doesn't disappear. Instead, you lose control of the elemental, it becomes hostile toward you and your companions, and it might attack. An uncontrolled elemental can't be dismissed by you, and it disappears 1 hour after you summoned it.

The GM has the elemental's statistics.

\subparagraph*{Overcast} When you cast this spell using more than 12 AET, the hit points of the summoned elemental increase by 5 for each additional AET spent and it gains an additional attack for every 3 additional AET spent.


\DndSpellHeader{Conjure Mephits\label{spell:conjure-mephits}}
{8 AET}
{1 minute}
{90 feet}
{V, S}
{Concentration, up to 1 hour}

You summon mephits that appear in unoccupied spaces that you can see within range. You choose one the following options for what appears:
\begin{itemize}
\item 4 of any of the following: Dust, Ice, or Magma mephits
\item 5 of any of the following: Mud, Smoke, or Steam mephits
\end{itemize}

The summoned creatures are friendly to you and your companions. Roll initiative for the summoned creatures as a group, which has its own turns. They are not under your control, but will attack and hamper enemies. When your concentration ends or is broken, they disappear.

The GM has the creatures' statistics.


\DndSpellHeader{Conjure Woodland Beings\label{spell:conjure-woodland-beings}}
{8 AET}
{1 action}
{60 feet}
{V, S, M (one holly berry per creature summoned)}
{Concentration, up to 1 hour}

You summon fey creatures that appear in unoccupied spaces that you can see within range. Choose one of the following options for what appears:

\begin{itemize}
\item One kipuliin
\item Three satyrs or three blink dogs
\end{itemize}

A summoned creature disappears when it drops to 0 hit points or when the spell ends.

The summoned creatures are friendly to you and your companions. Roll initiative for the summoned creatures as a group, which have their own turns. They obey any verbal commands that you issue to them (no action required by you). If you don't issue any commands to them, they defend themselves from hostile creatures, but otherwise take no actions.

The GM has the creatures' statistics.


\DndSpellHeader{Contagion\label{spell:contagion}}
{12 AET}
{1 action}
{Touch}
{V, S}
{7 days}

Your touch inflicts disease. Make a melee spell attack against a creature within your reach. On a hit, you afflict the creature with a disease of your choice from any of the ones described below.

At the end of each of the target's turns, it must make a Constitution saving throw. After failing three of these saving throws, the disease's effects last for the duration, and the creature stops making these saves. After succeeding on three of these saving throws, the creature recovers from the disease, and the spell ends.

Since this spell induces a natural disease in its target, any effect that removes a disease or otherwise ameliorates a disease's effects apply to it.

\subparagraph*{Blinding Sickness} Pain grips the creature's mind, and its eyes turn milky white. The creature has disadvantage on Wisdom checks and Wisdom saving throws and is blinded.

\subparagraph*{Filth Fever} A raging fever sweeps through the creature's body. The creature has disadvantage on Strength checks, Strength saving throws, and attack rolls that use Strength.

\subparagraph*{Flesh Rot} The creature's flesh decays. The creature has disadvantage on Charisma checks and vulnerability to all damage.

\subparagraph*{Mindfire} The creature's mind becomes feverish. The creature has disadvantage on Intelligence checks and Intelligence saving throws, and the creature behaves as if under the effects of the \textit{confusion} spell during combat.

\subparagraph*{Seizure} The creature is overcome with shaking. The creature has disadvantage on Dexterity checks, Dexterity saving throws, and attack rolls that use Dexterity.

\subparagraph*{Slimy Doom} The creature begins to bleed uncontrollably. The creature has disadvantage on Constitution checks and Constitution saving throws. In addition, whenever the creature takes damage, it is stunned until the end of its next turn.


\DndSpellHeader{Control Water\label{spell:control-water}}
{8 AET}
{1 action}
{300 feet}
{V, S, M (a drop of water and a pinch of dust)}
{Concentration, up to 10 minutes}

Until the spell ends, you control any freestanding water inside an area you choose that is a cube up to 100 feet on a side. You can choose from any of the following effects when you cast this spell. As an action on your turn, you can repeat the same effect or choose a different one.

\subparagraph*{Flood} You cause the water level of all standing water in the area to rise by as much as 20 feet. If the area includes a shore, the flooding water spills over onto dry land.

If you choose an area in a large body of water, you instead create a 20-foot tall wave that travels from one side of the area to the other and then crashes down. Any Huge or smaller vehicles in the wave's path are carried with it to the other side. Any Huge or smaller vehicles struck by the wave have a 25 percent chance of capsizing.

The water level remains elevated until the spell ends or you choose a different effect. If this effect produced a wave, the wave repeats on the start of your next turn while the flood effect lasts.

\subparagraph*{Part Water} You cause water in the area to move apart and create a trench. The trench extends across the spell's area, and the separated water forms a wall to either side. The trench remains until the spell ends or you choose a different effect. The water then slowly fills in the trench over the course of the next round until the normal water level is restored.

\subparagraph*{Redirect Flow} You cause flowing water in the area to move in a direction you choose, even if the water has to flow over obstacles, up walls, or in other unlikely directions. The water in the area moves as you direct it, but once it moves beyond the spell's area, it resumes its flow based on the terrain conditions. The water continues to move in the direction you chose until the spell ends or you choose a different effect.

\subparagraph*{Whirlpool} This effect requires a body of water at least 50 feet square and 25 feet deep. You cause a whirlpool to form in the center of the area. The whirlpool forms a vortex that is 5 feet wide at the base, up to 50 feet wide at the top, and 25 feet tall. Any creature or object in the water and within 25 feet of the vortex is pulled 10 feet toward it. A creature can swim away from the vortex by making a Strength (Athletics) check against your spell save DC.

When a creature enters the vortex for the first time on a turn or starts its turn there, it must make a Strength saving throw. On a failed save, the creature takes 2d8 bludgeoning damage and is caught in the vortex until the spell ends. On a successful save, the creature takes half damage, and isn't caught in the vortex. A creature caught in the vortex can use its action to try to swim away from the vortex as described above, but has disadvantage on the Strength (Athletics) check to do so.

The first time each turn that an object enters the vortex, the object takes 2d8 bludgeoning damage; this damage occurs each round it remains in the vortex.


\DndSpellHeader{Create or Destroy Water\label{spell:create-or-destroy-water}}
{2 AET}
{1 action}
{30 feet}
{V, S, M (a drop of water if creating water or a few grains of sand if destroying it)}
{Instantaneous}

You either create or destroy water.

\subparagraph*{Create Water} You create up to 10 gallons of clean water within range in an open container. Alternatively, the water falls as rain in a 30-foot cube within range, extinguishing exposed flames in the area.

\subparagraph*{Destroy Water} You destroy up to 10 gallons of water in an open container within range. Alternatively, you destroy fog in a 30-foot cube within range.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you create or destroy 10 additional gallons of water, or the size of the cube increases by 5 feet, for every 2 additional AET spent.


\DndSpellHeader{Cure Wounds\label{spell:cure-wounds}}
{1 AET}
{1 action}
{Touch}
{V, S}
{Instantaneous}

A creature you touch regains a number of hit points equal to 1d8 + your spellcasting ability modifier. This spell has no effect on undead or constructs.

\subparagraph*{Overcast} When you cast this spell using more than 1 AET, the healing increases by 1d8 for every additional AET spent, up to a maximum of 12d8 + your spellcasting ability modifier.

\subsection{Spells (D)}


\DndSpellHeader{Dancing Lights\label{spell:dancing-lights}}
{cantrip}
{1 action}
{120 feet}
{V, S, M (a bit of phosphorus or wychwood, or a glowworm)}
{Concentration, up to 1 minute}

You create up to four torch-sized lights within range, making them appear as torches, lanterns, or glowing orbs that hover in the air for the duration. You can also combine the four lights into one glowing vaguely humanoid form of Medium size. Whichever form you choose, each light sheds dim light in a 10 foot radius.

As a bonus action on your turn, you can move the lights up to 60 feet to a new spot within range. A light must be within 20 feet of another light created by this spell, and a light winks out if it exceeds the spell's range.


\DndSpellHeader{Darkness\label{spell:darkness}}
{3 AET}
{1 action}
{60 feet}
{V, M (bat fur and a drop of pitch or piece of coal)}
{Concentration, up to 10 minutes}

Opaque magical darkness spreads from a point you choose within range to fill a 15-foot-radius sphere for the duration. The darkness spreads around corners. A creature with darkvision can't see through this darkness, and nonmagical light can't illuminate it.

If the point you choose is on an object you are holding or one that isn't being worn or carried, the darkness emanates from the object and moves with it. Completely covering the source of the darkness with an opaque object, such as a bowl or a helm, blocks the darkness.

If any of this spell's area overlaps with an area of light created by a spell of costing 3 AET or less, the spell that created the light is dispelled.


\DndSpellHeader{Darkvision\label{spell:darkvision}}
{3 AET}
{1 action}
{Touch}
{V, S, M (either a pinch of dried carrot or an agate)}
{8 hours}

You touch a willing creature to grant it the ability to see in the dark. For the duration, that creature has darkvision out to a range of 60 feet. If it already had darkvision, it can see color in the dark.


\DndSpellHeader{Daylight\label{spell:daylight}}
{5 AET}
{1 action}
{60 feet}
{V, S}
{1 hour}

A 60-foot-radius sphere of sunlight spreads out from a point you choose within range. The sphere is bright light and sheds dim light for an additional 60 feet.

If you chose a point on an object you are holding or one that isn't being worn or carried, the light shines from the object and moves with it. Completely covering the affected object with an opaque object, such as a bowl or a helm, blocks the light.

If any of this spell's area overlaps with an area of darkness created by a spell costing 3 or less AET, the spell that created the darkness is dispelled.


\DndSpellHeader{Death Ward\label{spell:death-ward}}
{8 AET}
{1 action}
{Touch}
{V, S}
{8 hours}

You touch a creature and grant it a measure of protection from death.

The first time the target would drop to 0 hit points as a result of taking damage, the target instead drops to 1 hit point, and the spell ends.

If the spell is still in effect when the target is subjected to an effect that would kill it instantaneously without dealing damage, that effect is instead negated against the target, and the spell ends.


\DndSpellHeader{Detect Thoughts\label{spell:detect-thoughts}}
{3 AET}
{1 action}
{Self}
{V, S, M (a copper piece)}
{Concentration, up to 1 minute}

For the duration, you can read the thoughts of certain creatures. When you cast the spell and as your action on each turn until the spell ends, you can focus your mind on any one creature that you can see within 30 feet of you. If the creature you choose has an Intelligence of 3 or lower or doesn't speak any language, the creature is unaffected.

You initially learn the surface thoughts of the creature—what is most on its mind in that moment. As an action, you can either shift your attention to another creature's thoughts or attempt to probe deeper into the same creature's mind. If you probe deeper, the target must make a Wisdom saving throw. If it fails, you gain insight into its reasoning (if any), its emotional state, and something that looms large in its mind (such as something it worries over, loves, or hates). If it succeeds, the spell ends. Either way, the target knows that you are probing into its mind, and unless you shift your attention to another creature's thoughts, the creature can use its action on its turn to make an Intelligence check contested by your Intelligence check; if it succeeds, the spell ends.

Questions verbally directed at the target creature naturally shape the course of its thoughts, so this spell is particularly effective as part of an interrogation.

You can also use this spell to detect the presence of thinking creatures you can't see. When you cast the spell or as your action during the duration, you can search for thoughts within 30 feet of you. The spell can penetrate barriers, but 2 feet of rock, 2 inches of any metal other than lead, or a thin sheet of lead blocks you. You can't detect a creature with an Intelligence of 3 or lower or one that doesn't speak any language.

Once you detect the presence of a creature in this way, you can read its thoughts for the rest of the duration as described above, even if you can't see it, but it must still be within range.


\DndSpellHeader{Dimension Door\label{spell:dimension-door}}
{8 AET}
{1 action}
{500 feet}
{V}
{Instantaneous}

You teleport yourself from your current location to any other spot within range. You arrive at exactly the spot desired. It can be a place you can see, one you can visualize, or one you can describe by stating distance and direction, such as “200 feet straight downward” or “upward to the northwest at a 45-degree angle, 300 feet.”

You can bring along objects as long as their weight doesn't exceed what you can carry. You can also bring one willing creature of your size or smaller (or a medium creature if you are small) who is carrying gear up to its carrying capacity. The creature must be within 5 feet of you when you cast this spell.

If you would arrive in a place already occupied by an object or a creature, you and any creature traveling with you each take 4d6 bludgeoning damage, and the spell fails to teleport you.


\DndSpellHeader{Disguise Self\label{spell:disguise-self}}
{2 AET}
{1 action}
{Self}
{V, S}
{1 hour}

You make yourself—including your clothing, armor, weapons, and other belongings on your person— look different until the spell ends or until you use your action to dismiss it. You can seem 1 foot shorter or taller and can appear thin, fat, or in between. You can't change your body type, so you must adopt a form that has the same basic arrangement of limbs. Otherwise, the extent of the illusion is up to you.

The changes wrought by this spell fail to hold up to physical inspection. For example, if you use this spell to add a hat to your outfit, objects pass through the hat, and anyone who touches it would feel nothing or would feel your head and hair. If you use this spell to appear thinner than you are, the hand of someone who reaches out to touch you would bump into you while it was seemingly still in midair.

To discern that you are disguised, a creature can use its action to inspect your appearance and must succeed on an Intelligence (Investigation) check against your spell save DC. Creatures that physically interact with the disguise in a way that would reveal its falsity as part of a different action can immediate make the check without spending another action.


\DndSpellHeader{Dispel Otherworldly Influence\label{spell:dispel-otherworldly-influence}}
{12 AET}
{1 action}
{Self}
{V, S, M (holy water or powdered silver and iron)}
{Concentration, up to 1 minute}

Shimmering energy surrounds and protects you from fey, undead, and creatures originating from beyond the Material Plane. For the duration, celestials, elementals, fey, fiends, and undead have disadvantage on attack rolls against you.

You can end the spell early by using either of the following special functions.

\subparagraph*{Break Enchantment} As your action, you touch a creature you can reach that is charmed, frightened, or possessed by a celestial, an elemental, a fey, a fiend, or an undead. The creature you touch is no longer charmed, frightened, or possessed by such creatures.

\subparagraph*{Dismissal} As your action, make a melee spell attack against a celestial, an elemental, a fey, a fiend, or an undead you can reach. On a hit, you attempt to drive the creature back to its home plane. The creature must succeed on a Charisma saving throw or be sent back to its home plane (if it isn't there already). If they aren't on their home plane, undead are sent to the Shadowfell, and fey are sent to the Feywild.


\DndSpellHeader{Divine Favor\label{spell:divine-favor}}
{2 AET}
{1 bonus action}
{Self}
{V, S}
{Concentration, up to 1 minute}

Your prayer empowers you with divine radiance. Until the spell ends, your weapon attacks deal an extra 1d4 radiant damage on a hit.

\DndSpellHeader{Divine Wrath\label{spell:divine-wrath}}
{12 AET}
{1 action}
{Self (15 ft radius)}
{V, S}
{Concentration, up to 1 minute}

You call upon holy power and radiate destruction. When you cast this spell and at the start of each of your turns for the duration, a wave of radiant energy pulses out in a 15 foot sphere. All creatures in the area must make a Charisma saving throw, taking 5d8 radiant damage on a failed save or half as much on a success. Fiends, undead, and creatures summoned by magic make this saving throw at disadvantage. When you cast the spell, you can choose a number of targets up to your proficiency bonus. Designated targets are immune to the effect.

\DndSpellHeader{Dominate Beast\label{spell:dominate-beast}}
{8 AET}
{1 action}
{60 feet}
{V, S}
{Concentration, up to 1 minute}

You attempt to beguile a beast that you can see within range. It must succeed on a Wisdom saving throw or be charmed by you for the duration. If you or creatures that are friendly to you are fighting it, it has advantage on the saving throw.

While the beast is charmed, you have a telepathic link with it as long as the two of you are on the same plane of existence. You can use this telepathic link to issue commands to the creature while you are conscious (no action required), which it does its best to obey. You can specify a simple and general course of action, such as “Attack that creature,” “Run over there,” or “Fetch that object.” If the creature completes the order and doesn't receive further direction from you, it defends and preserves itself to the best of its ability.

You can use your action to take total and precise control of the target. Until the end of your next turn, the creature takes only the actions you choose, and doesn't do anything that you don't allow it to do. During this time, you can also cause the creature to use a reaction, but this requires you to use your own reaction as well.

Each time the target takes damage, it makes a new Wisdom saving throw against the spell. If the saving throw succeeds, the spell ends.

\subparagraph*{Overcast} When you cast this spell and spend at least 11 AET, the duration is concentration, up to 10 minutes. When you use at least 14 AET, the duration is concentration, up to 1 hour.

\subsection{Spells (E)}

\DndSpellHeader{Enhance Ability\label{spell:enhance-ability}}
{3 AET}
{1 action}
{Touch}
{V, S, M (fur or a feather from a beast)}
{Concentration, up to 1 hour.}

You touch a creature and bestow upon it a magical enhancement. Choose one of the following effects; the target gains that effect until the spell ends.

\subparagraph*{Bear's Endurance} The target gains 2d6 temporary hit points, which are lost when the spell ends. It also gains advantage on saving throws against being incapacitated or poisoned.

\subparagraph*{Bull's Strength} The target's carrying capacity and jump distance doubles. It also counts as one size larger when grappling or being grappled.

\subparagraph*{Cat's Grace} The target gains a 20 ft climbing speed. It also doesn't take damage from falling 20 feet or less if it isn't incapacitated.

\subparagraph*{Eagle's Splendor} The target has advantage on Charisma saving throws. It also gains proficiency with your choice of Deception, Intimidation, or Persuasion checks. If it already had proficiency with that skill, it can add twice its proficiency bonus instead. If that was already the case, it gains advantage on those checks instead.

\subparagraph*{Fox's Cunning} The target has advantage on Intelligence saving throws as well as Intelligence checks made to discern illusions.

\subparagraph*{Owl's Wisdom} The target has advantage on checks made to find hidden objects and to avoid being surprised.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, you can target one additional creature for every 3 additional AET spent.


\DndSpellHeader{Enlarge/Reduce\label{spell:enlarge-reduce}}
{3 AET}
{1 action}
{30 feet}
{V, S, M (a pinch of powdered iron)}
{Concentration, up to 1 minute}

You cause a creature or an object you can see within range to grow larger or smaller for the duration. Choose either a creature or an object that is neither worn nor carried. If the target is unwilling, it can make a Charisma saving throw. On a success, the spell has no effect.

If the target is a creature, everything it is wearing and carrying changes size with it. Any item dropped by an affected creature returns to normal size at once.

\subparagraph*{Enlarge} The target's size doubles in all dimensions, and its weight is multiplied by eight. This growth increases its size by one category— from Medium to Large, for example. If there isn't enough room for the target to double its size, the creature or object attains the maximum possible size in the space available. Until the spell ends, the target also has advantage on Strength checks and Strength saving throws. The target's weapons also grow to match its new size. While these weapons are enlarged, the target's attacks with them deal 1d4 extra damage.

\subparagraph*{Reduce} The target's size is halved in all dimensions, and its weight is reduced to one-eighth of normal. This reduction decreases its size by one category—from Medium to Small, for example. Until the spell ends, the target also has disadvantage on Strength checks and Strength saving throws. The target's weapons also shrink to match its new size. While these weapons are reduced, the target's attacks with them deal 1d4 less damage (this can't reduce the damage below 1).


\DndSpellHeader{Entangle\label{spell:entangle}}
{2 AET}
{1 action}
{90 feet}
{V, S}
{Concentration, up to 1 minute}

Grasping weeds and vines sprout from the ground in a 20-foot square starting from a point within range. For the duration, these plants turn the ground in the area into difficult terrain.

A creature in the area when you cast the spell must succeed on a Strength saving throw or be restrained by the entangling plants until the spell ends. A creature restrained by the plants can use its action to make a Strength check against your spell save DC. On a success, it frees itself.

When the spell ends, the conjured plants wilt away.


\DndSpellHeader{Enthrall\label{spell:enthrall}}
{4 AET}
{1 action}
{60 ft}
{V, M (a gold pendant worth at least 100 gp)}
{Concentration, up to 1 minute (see text)}

You weave a distracting string of words, causing creatures of your choice that you can see within range and that can hear you to make a Wisdom saving throw. Any creature that can't be charmed succeeds on this saving throw automatically, and if you or your companions are fighting a creature, it has advantage on the save. On a failed save, the target is blinded and deafened to all creatures other than you for 1 minute. If you move out of sight of them or they can no longer hear you, they can attempt the saving throw at the end of each of their turns, ending the effect on a success. Once the effect ends for a target, they cannot be affected by it for 48 hours regardless of the source.

\subparagraph*{Overcast} When you cast this spell and spend at least 6 AET, the duration increases to 10 minutes. If you spend at least 8 AET, it no longer requires concentration and lasts for 10 minutes.

\DndSpellHeader{Expeditious Retreat\label{spell:expeditious-retreat}}
{2 AET}
{1 bonus action}
{Self}
{V, S}
{Concentration, up to 10 minutes}

This spell allows you to move at an incredible pace. When you cast this spell, and then as a bonus action on each of your turns until the spell ends, you can take the Dash action.

\subsection{Spells (F)}
\DndSpellHeader{Faerie Fire\label{spell:faerie-fire}}
{2 AET}
{1 action}
{60 feet}
{V}
{Concentration, up to 1 minute}

Each object in a 20-foot cube within range is outlined in blue, green, or violet light (your choice). Any creature in the area when the spell is cast is also outlined in light if it fails a Dexterity saving throw. For the duration, objects and affected creatures shed dim light in a 10-foot radius.

Any attack roll against an affected creature or object has advantage if the attacker can see it, and the affected creature or object can't benefit from being invisible.


\DndSpellHeader{Faithful Hound\label{spell:faithful-hound}}
{8 AET}
{1 action}
{30 feet}
{V, S, M (a tiny silver whistle, a piece of bone, and a thread)}
{8 hours}

You conjure a phantom watchdog in an unoccupied space that you can see within range, where it remains for the duration, until you dismiss it as an action, or until you move more than 100 feet away from it.

The hound is invisible to all creatures except you and can't be harmed. When a Small or larger creature comes within 30 feet of it without first speaking the password that you specify when you cast this spell, the hound starts barking loudly. The hound sees invisible creatures and can see into the Border Shadow. It ignores illusions.

At the start of each of your turns, the hound attempts to bite one creature within 5 feet of it that is hostile to you. The hound's attack bonus is equal to your spellcasting ability modifier + your proficiency bonus. On a hit, it deals 4d8 piercing damage.


\DndSpellHeader{False Life\label{spell:false-life}}
{2 AET}
{1 action}
{Self}
{V, S, M (a small amount of alcohol or distilled spirits)}
{1 hour}

Bolstering yourself with a necromantic facsimile of life, you gain 1d4 + 4 temporary hit points for the duration.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, you gain 5 additional temporary hit points for every 2 additional AET spent.


\DndSpellHeader{Fear\label{spell:fear}}
{5 AET}
{1 action}
{Self (30-foot cone)}
{V, S, M (a white feather or the heart of a hen)}
{Concentration, up to 1 minute, save ends}

You project a phantasmal image of a creature's worst fears. Each creature in a 30-foot cone must succeed on a Wisdom saving throw or become shaken for the duration. If they were already shaken, they become frightened instead. Shaken creatures can make the saving throw at the end of each of their turns, ending the effect on a success or becoming frightened if they fail one of these subsequent saving throws by 5 or more.

While frightened by this spell, a creature that can see you must take the Dash action and move away from you by the safest available route on each of its turns, unless there is nowhere to move. Frightened creatures can make the saving throw again at the end of any turn if they cannot see you at that point.


\DndSpellHeader{Feather Fall\label{spell:feather-fall}}
{2 AET}
{1 reaction, which you take when you or a creature within 60 feet of you falls}
{60 feet}
{V, M (a small feather or piece of down)}
{1 minute}

Choose up to five falling creatures within range. A falling creature's rate of descent slows to 60 feet per round until the spell ends. If the creature lands before the spell ends, it takes no falling damage and can land on its feet, and the spell ends for that creature.


\DndSpellHeader{Find Traps\label{spell:find-traps}}
{3 AET}
{1 action}
{120 feet}
{V, S}
{Instantaneous}

You sense the presence of any trap within range that is within line of sight. A trap, for the purpose of this spell, includes anything that would inflict a sudden or unexpected effect you consider harmful or undesirable. Thus, the spell would sense an area affected by the \textit{alarm} or \textit{glyph of warding} incantation, or a mechanical pit trap as well as a natural weakness in the floor, an unstable ceiling, or a hidden sinkhole.

You learn the location and the general nature of the danger posed by each trap.


\DndSpellHeader{Fireball\label{spell:fireball}}
{5 AET}
{1 action}
{150 feet}
{V, S, M (a tiny ball of bat guano and sulfur)}
{Instantaneous}

A bright streak flashes from your pointing finger to a point you choose within range and then blossoms with a low roar into an explosion of flame. Each creature in a 20-foot-radius sphere centered on that point must make a Dexterity saving throw. A target takes 6d6 fire damage on a failed save, or half as much damage on a successful one. Targets that fail their saving throw by 5 or more ignite, taking an additional 2d6 fire damage at the beginning of each of their turns until they or someone else spends an action to extinguish the flames. 

The fire spreads around corners. It ignites flammable objects in the area that aren't being worn or carried.

\subparagraph*{Overcast} When you cast this spell using more than 5 AET, the initial damage increases by 1d6 for every 2 additional AET spent.


\DndSpellHeader{Fire Shield\label{spell:fire-shield}}
{8 AET}
{1 action}
{Self}
{V, S, M (a bit of phosphorus or a firefly)}
{10 minutes}

Thin and wispy flames wreathe your body for the duration, shedding bright light in a 10-foot radius and dim light for an additional 10 feet. You can end the spell early by using an action to dismiss it.

The flames provide you with a warm shield or a chill shield, as you choose. The warm shield grants you resistance to cold damage, and the chill shield grants you resistance to fire damage.

In addition, whenever a creature within 5 feet of you hits you with a melee attack, the shield erupts with flame. The attacker takes 2d8 fire damage from a warm shield, or 2d8 cold damage from a cold shield.


\DndSpellHeader{Flame Blade\label{spell:flame-blade}}
{3 AET}
{1 bonus action}
{Self}
{V, S, M (leaf of sumac)}
{Concentration, up to 10 minutes}

You evoke a fiery scimitar in your free hand. The blade is lasts for the duration and you have proficiency in it. If you let go of the blade, it disappears, but you can evoke the blade again as a bonus action.

You can use your action to make a melee spell attack with the fiery blade. On a hit, the target takes 3d6 fire damage.

Alternatively, you can use it as a normal scimitar, except that it deals fire damage.

The flaming blade sheds bright light in a 10-foot radius and dim light for an additional 10 feet.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage increases by 1d6 for every 4 additional AET spent. When used as a normal weapon, the damage dice increase by 1d6 for every 4 additional AET.


\DndSpellHeader{Flame Strike\label{spell:flame-strike}}
{9 AET}
{1 action}
{60 feet}
{V, S, M (pinch of sulfur)}
{Instantaneous}

A vertical column of divine fire roars down from the heavens in a location you specify. Each creature in a 10-foot-radius, 40-foot-high cylinder centered on a point within range must make a Dexterity saving throw. A creature takes 4d6 fire damage and 4d6 radiant damage on a failed save, or half as much damage on a successful one. Undead and fiends have disadvantage on the saving throw.

\subparagraph*{Overcast} When you cast this spell using more than 9 AET, the fire damage and the radiant damage each increase by 1d6 for every 3 additional AET spent.


\DndSpellHeader{Flaming Sphere\label{spell:flaming-sphere}}
{3 AET}
{1 action}
{60 feet}
{V, S, M (a bit of tallow, a pinch of brimstone, and a dusting of powdered iron)}
{Concentration, up to 1 minute}

A 5-foot-diameter sphere of fire appears in an unoccupied space of your choice within range and lasts for the duration. The sphere is a solid object that occupies its space. Any creature that ends its turn within 5 feet of the sphere must make a Dexterity saving throw. The creature takes 2d6 fire damage on a failed save, or half as much damage on a successful one.

As a bonus action, you can move the sphere up to 30 feet. If you ram the sphere into a creature, that creature must make the saving throw against the sphere's damage, and the sphere stops moving this turn.

When you move the sphere, you can direct it over barriers up to 5 feet tall and jump it across pits up to 10 feet wide. The sphere ignites flammable objects not being worn or carried, and it sheds bright light in a 20-foot radius and dim light for an additional 20 feet.

\subparagraph*{Overcast} When you cast this spell using more than 3 AET, the damage increases by 1d6 for every 2 additional AET spent.


\DndSpellHeader{Flash Freeze\label{spell:flash-freeze}}
{2 AET}
{1 action}
{60 feet}
{V, S}
{Instantaneous}

The air around a creature you can see within range takes on the aspect of ice. The target must make a Constitution saving throw. On a failed save, the target takes 4d6 cold damage and their movement is reduced to 0 until the end of their next turn. On a success, they take half damage and are not slowed.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d6 for every 2 additional AET spent.


\DndSpellHeader{Fog Cloud\label{spell:fog-cloud}}
{2 AET}
{1 action}
{120 feet}
{V, S}
{Concentration, up to 1 hour}

You create a 20-foot-radius sphere of fog centered on a point within range. The sphere spreads around corners, and its area is heavily obscured. It lasts for the duration or until a wind of moderate or greater speed (at least 10 miles per hour) disperses it.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the radius of the fog increases by 10 feet for every 3 additional AET spent.


\DndSpellHeader{Freedom of Movement\label{spell:freedom-of-movement}}
{8 AET}
{1 action}
{Touch}
{V, S, M (a leather strap, bound around the arm or a similar appendage)}
{1 hour}

You touch a willing creature. For the duration, the target's movement is unaffected by difficult terrain, and spells and other magical effects can neither reduce the target's speed nor cause the target to be paralyzed or restrained.

The target can also spend 5 feet of movement to automatically escape from nonmagical restraints, such as manacles or a creature that has it grappled. Finally, being underwater imposes no penalties on the target's movement or attacks.

\subsection{Spells (G)}


\DndSpellHeader{Gaseous Form\label{spell:gaseous-form}}
{5 AET}
{1 action}
{Touch}
{V, S, M (a bit of gauze and a wisp of smoke)}
{Concentration, up to 1 hour}

You transform a willing creature you touch, along with everything it's wearing and carrying, into a misty cloud for the duration. The spell ends if the creature drops to 0 hit points. An incorporeal creature isn't affected.

While in this form, the target's only method of movement is a flying speed of 10 feet. The target can enter and occupy the space of another creature. The target has resistance to nonmagical damage, and it has advantage on Strength, Dexterity, and Constitution saving throws. The target can pass through small holes, narrow openings, and even mere cracks, though it treats liquids as though they were solid surfaces. The target can't fall and remains hovering in the air even when stunned or otherwise incapacitated.

While in the form of a misty cloud, the target can't talk or manipulate objects, and any objects it was carrying or holding can't be dropped, used, or otherwise interacted with. The target can't attack or cast spells.


\DndSpellHeader{Giant Insect\label{spell:giant-insect}}
{8 AET}
{1 action}
{30 feet}
{V, S}
{Concentration, up to 10 minutes}

You transform up to ten centipedes, three spiders, five wasps, or one scorpion within range into giant versions of their natural forms for the duration. A centipede becomes a giant centipede, a spider becomes a giant spider, a wasp becomes a giant wasp, and a scorpion becomes a giant scorpion.

Each creature obeys your verbal commands, and in combat, they act on your turn each round. The GM has the statistics for these creatures and resolves their actions and movement.

A creature remains in its giant size for the duration, until it drops to 0 hit points, or until you use an action to dismiss the effect on it.

The GM might allow you to choose different targets. For example, if you transform a bee, its giant version might have the same statistics as a giant wasp.


\DndSpellHeader{Grave Touch\label{spell:grave-touch}}
{cantrip}
{1 action}
{120 feet}
{V, S}
{1 round}

You create a ghostly, skeletal hand in the space of a creature within range. Make a ranged spell attack against the creature to assail it with the chill of the grave. On a hit, the target takes 1d8 necrotic damage, and it can't regain hit points until the start of your next turn. Until then, the hand clings to the target.

If you hit an undead target, it also has disadvantage on attack rolls against you until the end of your next turn.

This spell's damage increases by 1d8 when you reach 5th level (2d8), 11th level (3d8), and 17th level (4d8).


\DndSpellHeader{Grease\label{spell:grease}}
{2 AET}
{1 action}
{60 feet}
{V, S, M (a bit of pork rind or butter)}
{1 minute}

Slick grease covers the ground in a 10-foot square centered on a point within range and turns it into difficult terrain for the duration.

When the grease appears, each creature standing in its area must succeed on a Dexterity saving throw or fall prone. A creature that enters the area or ends its turn there must also succeed on a Dexterity saving throw or fall prone.

This grease is flammable, acting like a flask of oil. Any ignition source ignites the entire area.


\DndSpellHeader{Greater Invisibility\label{spell:greater-invisibility}}
{8 AET}
{1 action}
{Touch}
{V, S}
{Concentration, up to 1 minute}

You or a creature you touch becomes invisible until the spell ends. Anything the target is wearing or carrying is invisible as long as it is on the target's person.

\DndSpellHeader{Guidance\label{spell:guidance}}
{cantrip}
{1 action}
{Touch}
{V, S}
{Concentration, up to 10 minutes}

You touch one willing creature. Once before the spell ends, the target can roll a d4 and add the number rolled to one ability check of its choice. It can roll the die before or after making the ability check. The spell then ends. 

\DndSpellHeader{Guiding Bolt\label{spell:guiding-bolt}}
{2 AET}
{1 action}
{120 feet}
{V, S}
{1 round}

A flash of light streaks toward a creature of your choice within range. Make a ranged spell attack against the target. On a hit, the target takes 4d6 radiant damage, and the next attack roll made against this target before the end of your next turn has advantage, thanks to the mystical dim light glittering on the target until then.

\subparagraph*{Overcast} When you cast this spell using more than 2 AET, the damage increases by 1d6 for every 2 additional AET spent.

\DndSpellHeader{Gust of Wind\label{spell:gust-of-wind}}
{3 AET}
{1 action}
{Self (60-foot line)}
{V, S, M (a legume seed)}
{Concentration, up to 1 minute}

A line of strong wind 60 feet long and 10 feet wide blasts from you in a direction you choose for the spell's duration. Each creature that starts its turn in the line must succeed on a Strength saving throw or be pushed 15 feet away from you in a direction following the line.

Any creature in the line must spend 2 feet of movement for every 1 foot it moves when moving closer to you.

The gust disperses gas or vapor, and it extinguishes candles, torches, and similar unprotected flames in the area. It causes protected flames, such as those of lanterns, to dance wildly and has a 50 percent chance to extinguish them.

As a bonus action on each of your turns before the spell ends, you can change the direction in which the line blasts from you.