\subsection{By Cost}
Abbreviations and notes: MOD is your spellcasting modifier (ie INT for arcanists and spellblades, etc.). (C) indicates a concentration spell. ATK means "requires a spell attack". "Save for half" means that on a successful save the target takes half damage and no other effect. "Save ends" means that the target can re-attempt the saving throw at the end of each of their turns, ending the effect on a success.

\subsubsection{0 (Cantrips)}
\begin{itemize}
	\item \nameref{spell:acid-burst} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): acid damage 60 ft, CON, 1d8. Damage over time.
	\item \nameref{spell:dancing-lights} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): utility (C, 1 minute). Creates several mobile lights.
	\item \nameref{spell:grave-touch} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:spellblade}): necrotic damage 90 ft, ATK, 1d8. Prevents healing 1 rnd.
	\item \nameref{spell:guidance} (\nameref{class:priest}): utility (C, 10 minutes), touch. Assists with one check.
	\item \nameref{spell:light} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:shaman}, \nameref{class:spellblade}): utility. Creates a torch-like light.
	\item \nameref{spell:mage-hand} (\nameref{class:arcanist}, \nameref{class:spellblade}): utility (C, 1 minute), 30 ft. Limited third hand.
	\item \nameref{spell:message} (\nameref{class:arcanist}, \nameref{class:priest}): communications. Limited-range outgoing telepathy.
	\item \nameref{spell:minor-illusion} (\nameref{class:arcanist}): utility (C, 1 minute). Creates small static illusion.
	\item \nameref{spell:prestidigitation} (\nameref{class:arcanist}, \nameref{class:spellblade}): utility. Creates various effects.
	\item \nameref{spell:produce-flame} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): fire damage 90 ft, ATK, 1d8. Can be held for light.
	\item \nameref{spell:ray-of-frost} (\nameref{class:arcanist}, \nameref{class:spellblade}): cold damage 90 ft, ATK, 1d8. Reduces speed.
	\item \nameref{spell:resistance} (\nameref{class:priest}, \nameref{class:spellblade}): Defense (C, 1 minute), self. Assists with one save.
	\item \nameref{spell:sacred-flame} (\nameref{class:priest}): radiant damage, 60 ft. DEX, 1d8. Ignores cover.
	\item \nameref{spell:shillelagh} (\nameref{class:priest}, \nameref{class:shaman}): utility (C, 1 hour), touch. Makes wooden weapon magical.
	\item \nameref{spell:shocking-grasp} (\nameref{class:arcanist}, \nameref{class:spellblade}): lightning damage, touch. ATK, 1d8. Prevents reactions 1 rnd.
	\item \nameref{spell:thaumaturgy} (\nameref{class:priest}): utility. Creates various effects.
\end{itemize}

\subsubsection{1-2 Aether}
\begin{itemize}
	\item \nameref{spell:bane} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): debuff, 3 targets, 90 ft. wis. \textminus 1d4 to rolls.
	\item \nameref{spell:bless} (\nameref{class:oathbound}, \nameref{class:priest}): 3 targets, 90 ft. +1d4 to rolls.
	\item \nameref{spell:burning-hands} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 15 ft cone, DEX, 2d6 fire (save for half).
	\item \nameref{spell:charm-person} (\nameref{class:priest}): social (C, 1 minute), WIS, 60ft. change attitude to friendly. 
	\item \nameref{spell:color-spray} (\nameref{class:arcanist}, \nameref{class:spellblade}): 15 ft cone, WIS, blinded (save ends).
	\item \nameref{spell:command} (\nameref{class:oathbound}, \nameref{class:priest}): 60 ft., 1 target. WIS, 1 word command
	\item \nameref{spell:create-or-destroy-water} (\nameref{class:priest}, \nameref{class:shaman}): 30 ft. Create or destroy up to 10 gal. water.
	\item \nameref{spell:cure-wounds} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}): Touch, heal 1d8 + MOD.
	\item \nameref{spell:disguise-self} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:spellblade}): Self, not concentration, illusionary change to self.
	\item \nameref{spell:divine-favor} (\nameref{class:oathbound}): Self (C, 1 min), weapon attacks deal +1d4 radiant.
	\item \nameref{spell:entangle} (\nameref{class:ranger}, \nameref{class:shaman}): 20ft cube (C, 1 min), 90ft. STR or restrained.
	\item \nameref{spell:expeditious-retreat} (\nameref{class:arcanist}, \nameref{class:spellblade}): Self (C, 10 min), Dash as bonus action.
	\item \nameref{spell:faerie-fire} (\nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 20ft cube (C, 1 minute), 60 ft. DEX or glow + attacks are at advantage.
	\item \nameref{spell:false-life} (\nameref{class:arcanist}, \nameref{class:spellblade}): Self, 1 hr. 1d4 + 4 THP.
	\item \nameref{spell:feather-fall} (\nameref{class:arcanist}, \nameref{class:priest}): Reaction, up to 5 falling creatures. 1 min. Fall at 60ft./rnd, no fall damage.
	\item \nameref{spell:flash-freeze} (\nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 1 creature, 60 ft. CON. 4d6 cold + no movement 1 rnd. Save for half.
	\item \nameref{spell:fog-cloud} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}): 20-ft radius (C, 1 hour), 120 ft. Heavily obscured. Dispersed by wind.
	\item \nameref{spell:grease} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}): 10 ft square, 60 ft. 1 min. DEX or fall prone. Flammable.
	\item \nameref{spell:guiding-bolt} (\nameref{class:priest}, \nameref{class:spellblade}): 120 ft. ATK. 4d6 radiant and next attack has advantage.
	\item \nameref{spell:healing-word} (\nameref{class:priest}): 60 ft. 1d4+MOD healing.
	\item \nameref{spell:heroism} (\nameref{class:oathbound}): Touch (C, 1 min). Immune to fear + MOD THP at start of each turn.
	\item \nameref{spell:hideous-laughter} (\nameref{class:spellblade}): 30 ft (CON, 1 minute), INT > 4. WIS or prone and incapacitated. Save ends (new save on damage + end of turn).
	\item \nameref{spell:inflict-wounds} (\nameref{class:priest}, \nameref{class:spellblade}): Touch, ATK. 3d10 necrotic.
	\item \nameref{spell:jump} (): Touch, 1 min. Jump distance tripled until spell ends.
	\item \nameref{spell:longstrider} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:ranger}): Touch, 1 hr. +10 ft. movement.
	\item \nameref{spell:mage-armor} (\nameref{class:arcanist}): Touch, 8 hrs. AC = 12 + DEX and can use deflect action. Incompatible with armor.
	\item \nameref{spell:magic-missile} (\nameref{class:arcanist}): 120 ft, 1-3 targets. 3 auto-hit darts, separately targeted. 1d4 + 1 piercing each.
	\item \nameref{spell:protection-from-otherworldly-influence} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:shaman}): Touch (C, 10 min). Protected (attacks at disadvantage, no charm/fear/possession) by aberrations, celestials, fey, fiends, undead.
	\item \nameref{spell:sanctuary} (\nameref{class:oathbound}, \nameref{class:priest}): 30ft. Attacks against creature must pass a WIS save. Ends if target attacks, deals damage, or casts a hostile spell.
	\item \nameref{spell:shield} (\nameref{class:arcanist}, \nameref{class:spellblade}): Self, 10 minutes. Barrier with 10 HP absorbs damage until depleted. Immune to magic missile.
	\item \nameref{spell:shield-of-faith} (\nameref{class:oathbound}, \nameref{class:priest}): Bonus action. 60 ft (C, 10 min). +2 AC. Does not stack with a physical shield. 
	\item \nameref{spell:silent-image} (\nameref{class:arcanist}): 15-ft cube (C, 10 min). 60 ft. Illusion (visual only). Can move image with action. 
	\item \nameref{spell:sleep} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 5 creatures in 20ft radius (C, 1 min), 90 ft. WIS or fall asleep until take damage or are awoken.
	\item \nameref{spell:thunderwave} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 15ft cube, Self. CON. 2d8 thunder and pushed 10 ft (save for half). Loud (300 ft).
	\item \nameref{spell:true-strike} (\nameref{class:oathbound}, \nameref{class:spellblade}, \nameref{class:ranger}): Bonus action. 60 ft. 1 round. Your next attack has advantage and crits on 19 or 20.
\end{itemize}

\subsubsection{3-4 Aether}
\begin{itemize}
	\item \nameref{spell:acid-arrow} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): 90 ft. Attack. 4d6 acid + 2d6 acid next turn. Half of 4d6 on miss.
	\item \nameref{spell:aid} (\nameref{class:oathbound}, \nameref{class:priest}): 3 creatures, 30 ft. 8 hrs. Current and max HP increases by 5.
	\item \nameref{spell:alter-self} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}): Self (C, 1 hr). Actual physical changes to body.
	\item \nameref{spell:barkskin} (\nameref{class:ranger}, \nameref{class:shaman}): Touch, 1 hr. AC is minimum 16 and MOD THP at the start of each of their turns.
	\item \nameref{spell:blindness-deafness} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): 30 ft, 1 minute. CON or blinded/deafened (your choice). Save ends.
	\item \nameref{spell:blur} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): Self (C, 1 min). All attacks against you have disadvantage unless target can see through illusions or doesn't need sight.
	\item \nameref{spell:calm-emotions} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:shaman}): Humanoids in 20 ft radius (C, 1 minute). 60 ft. Suppress charmed/frightened or make target indifferent unless attacked/friends harmed.
	\item \nameref{spell:darkness} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:spellblade}): 15 ft radius (C, 10 min), 60 ft. Opaque sphere of blackness, extinguishes spell-based light.
	\item \nameref{spell:darkvision} (\nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}): Touch, 8 hrs. Target gains darkvision (60 ft).
	\item \nameref{spell:detect-thoughts} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:ranger}): Self (C, 1 min), 30 ft. Can perceive intelligent (INT > 3 and speaks a language) minds and read surface thoughts. Can probe deeper, but might get detected.
	\item \nameref{spell:enhance-ability} (\nameref{class:spellblade}): Touch (C, 1 hour). One of several abilities (one per ability score).
	\item \nameref{spell:enlarge-reduce} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:shaman}, \nameref{class:spellblade}): 30 ft (C, 1 min). Creature or object grows or shrinks one size. CON negates if unwilling.
	\item \nameref{spell:find-traps} (\nameref{class:priest}, \nameref{class:ranger}): 120 ft radius, Self. Detect location and nature of traps and hazards in line of sight. 
	\item \nameref{spell:flame-blade} (\nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}): Self (C, 10 min). Bonus action. Create scimitar of flame. Can also melee spell attack (3d6 fire). Grants temporary proficiency.
	\item \nameref{spell:flaming-sphere} (\nameref{class:shaman}, \nameref{class:spellblade}): 60 ft (C, 1 min). Creates 5-ft diameter solid sphere of fire. 2d6 fire (DEX for half). Mobile as bonus action.
	\item \nameref{spell:gust-of-wind} (\nameref{class:shaman}): 60x10 line (C, 1 min), self. Line of strong wind, STR or pushed 15 feet. Re-aimable.
	\item \nameref{spell:heat-metal} (\nameref{class:shaman}, \nameref{class:spellblade}): 60 ft (C, 1 min). Targets metal object. Action to cause 2d8 fire damage, no save.
	\item \nameref{spell:hold-person} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:spellblade}): 60 ft (C, 1 min). Humanoid target, WIS or paralyzed (save ends).
	\item \nameref{spell:invisibility} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:spellblade}): Touch (C, 1 hr). Target becomes invisible until they attack, cast spell, or deal damage.
	\item \nameref{spell:levitate} (\nameref{class:arcanist}, \nameref{class:spellblade}): 60 ft (C, 10 min). Object or creature. Moves vertically, can't move normally. Can be moved as action.
	\item \nameref{spell:magic-weapon} (\nameref{class:oathbound}, \nameref{class:priest}): Touch (C, 1 hr). Bonus action. One non-magical weapon becomes magic +1.
	\item \nameref{spell:mirror-image} (\nameref{class:arcanist}): Self. 1 min. Potentially absorbs up to 3 attacks.
	\item \nameref{spell:misty-step} (\nameref{class:arcanist}, \nameref{class:spellblade}): Self. Bonus action. Teleport up to 30 ft.
	\item \nameref{spell:moonbeam} (\nameref{class:shaman}): 5x40 cylinder (C, 1 min), 120 ft. 2d10 radiant (CON for half). Can move as action. Shapechangers are penalized.
	\item \nameref{spell:pass-without-trace} (\nameref{class:shaman}, \nameref{class:ranger}): 30ft radius (C, 1 hr), self. Min Stealth roll of 10, cannot be nonmagically tracked.
	\item \nameref{spell:prayer-of-healing} (\nameref{class:oathbound}, \nameref{class:priest}): 6 targets, 30 ft. 10 min cast time. 2d8+MOD healing each.
	\item \nameref{spell:protection-from-poison} (\nameref{class:oathbound}, \nameref{class:priest}): Touch, 1 hr. Neutralize one poison, adv. on saves vs poison + resist poison damage.
	\item \nameref{spell:ray-of-enfeeblement} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:spellblade}): 60 ft (C, 1 min). ATK. 3d8 necrotic and half damage with weapon attacks. CON ends.
	\item \nameref{spell:scorching-ray} (\nameref{class:arcanist}, \nameref{class:spellblade}): 3 attacks, 120 ft. 2d6 fire per attack.
	\item \nameref{spell:see-invisibility} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:ranger}): Self (1 hr). See invisible and ethereal creatures.
	\item \nameref{spell:shatter} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:shaman}, \nameref{class:spellblade}): 10 ft radius, 60 ft. 3d8 (CON for half). Inorganic creatures have disadvantage.
	\item \nameref{spell:silence} (\nameref{class:priest}, \nameref{class:ranger}): 20 ft radius (C, 10 min), 120 ft. Area of silence, stops V components + immune to thunder.
	\item \nameref{spell:spike-growth} (\nameref{class:shaman}, \nameref{class:ranger}): 20 ft circle (C, 10 min), 150 ft. Area is difficult terrain and causes 2d4 piercing damage per 5 feet traveled (max 16d4/rnd).
	\item \nameref{spell:spiritual-weapon} (\nameref{class:priest}): 60 ft, 1 min. Bonus action. Spectral weapon appears, makes spell attacks as bonus action. 1d10 + MOD bludgeoning, piercing, or slashing. Can move up to 20 ft as part of bonus action.
	\item \nameref{spell:suggestion} (\nameref{class:priest}): 30 ft (C, 8 hr). Short, seemingly reasonable suggestion. WIS or must obey. Cannot be obviously harmful.
	\item \nameref{spell:vampiric-touch} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): Self (C, 1 min). Necrotic melee spell attack. 3d6 necrotic and heal for half. Can repeat action.
	\item \nameref{spell:warding-bond} (\nameref{class:oathbound}, \nameref{class:priest}): Touch + self (1 hr). While target in 60 ft, target gains bonus to AC and saves and has resistance to damage. You take same amount of damage it takes. 
	\item \nameref{spell:web} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:shaman}): 20 ft cube (C, 1 hr), 60 ft. DEX or restrained, STR to break out. Flammable.
\end{itemize}

\subsubsection{5-7 Aether}
\begin{itemize}
	\item \nameref{spell:beacon-of-hope} (\nameref{class:oathbound}, \nameref{class:priest}): 30 ft radius (C 1 min), self. Chosen targets have advantage on WIS saves and death saves and maximizes all healing received.
	\item \nameref{spell:bestow-curse} (\nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}): Touch (C, 1 min). WIS save or cursed with one of a few options.
	\item \nameref{spell:blight} (\nameref{class:arcanist}, \nameref{class:ranger}, \nameref{class:spellblade}): 30 ft. 10d8 necrotic (CON for half). Extra effect on plants; no effect on undead or constructs.
	\item \nameref{spell:blink} (\nameref{class:arcanist}, \nameref{class:spellblade}): Self (1 min). 50\% chance of being ethereal between turns. Dismissable.
	\item \nameref{spell:call-lightning} (\nameref{class:ranger}): 60x10 cylinder, 120 ft (C, 10 min). Requires enough space. Action to do 3d10 (Dex for half) lightning in 5 ft radius. Extra effect when stormy.
	\item \nameref{spell:clairvoyance} (\nameref{class:priest}, \nameref{class:ranger}): 1 mile (C, 10 min). 100 gp focus. Create invisible, immobile sensor at familiar or obvious unfamiliar location. Hearing or sight, can switch.
	\item \nameref{spell:conjure-animals} (\nameref{class:shaman}, \nameref{class:ranger}): 60 ft. (C, 1 hr). Create 1 OR 2 or 2 OR 1 from list. They obey verbal commands.
	\item \nameref{spell:daylight} (\nameref{class:oathbound}, \nameref{class:priest}): 60 ft, 1 hr. 60 bright (sunlight) + 60 ft dim light. Dispels magical darkness.
	\item \nameref{spell:fear} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:spellblade}): 30 ft cone (C, 1 min), self. WIS or frightened, drop objects, must dash away (safely).
	\item \nameref{spell:fireball} (\nameref{class:arcanist}, \nameref{class:spellblade}): 20 ft sphere, 150 ft. 6d6 fire (DEX for half). Ignites unattended objects.
	\item \nameref{spell:gaseous-form} (\nameref{class:shaman}): Touch (C, 1 hr). Willing target. Becomes slowly flying gas.
	\item \nameref{spell:haste} (\nameref{class:arcanist}, \nameref{class:priest}): 30 ft (C, 1 min). Willing creature moves faster, gains AC, advantage on DEX saves, and can Dash, Disengate, Hide, or Use an Object as an extra action.
	\item \nameref{spell:hypnotic-pattern} (\nameref{class:arcanist}, \nameref{class:spellblade}): 30 ft cube (C, 1 min), 120 ft. WIS or charmed and incapacitated, speed zero. WIS ends, as does damage or anyone's action.
	\item \nameref{spell:ice-storm} (\nameref{class:shaman}): 20x40 cylinder, 300 ft. 2d8 bludgeoning and 4d6 cold, DEX for half. Area is difficult terrain 1 rnd.
	\item \nameref{spell:lightning-bolt} (\nameref{class:arcanist}, \nameref{class:shaman}): 100x5 line. 6d6 lightning, DEX for half. Ignites unattended objects.
	\item \nameref{spell:major-image} (\nameref{class:arcanist}): 20 ft cube (C, 10 min), 120 ft. Multi-sensory illusion, can move as action. No negative effects.
	\item \nameref{spell:meld-into-stone} (\nameref{class:arcanist}): Touch, Self. 8 hrs. Become one with a chunk of stone touched.
	\item \nameref{spell:plant-growth} (\nameref{class:ranger}, \nameref{class:shaman}): 100 ft radius, 150 ft. Extra difficult terrain. Can exclude areas.
	\item \nameref{spell:protection-from-energy} (\nameref{class:arcanist}, \nameref{class:priest}): Touch (C, 1 hr). Target resists one energy type.
	\item \nameref{spell:remove-curse} (\nameref{class:oathbound}, \nameref{class:priest}): Touch. End all curses on target creature or break attunement to cursed object.
	\item \nameref{spell:revivify} (\nameref{class:priest}): Touch. 300 gp consumed component. Raise dead (1 min or less).
	\item \nameref{spell:sleet-storm} (\nameref{class:shaman}): 40x20 cylinder (C, 1 min), 150 ft. Difficult terrain. 1/turn DEX or prone. Provokes concentration saves every turn for targets in area.
	\item \nameref{spell:slow} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): 6 creatures in 40-ft cube (C, 1 min), 120 ft. WIS or slowed. Speed halved, lower AC/DEX saves, 1 attack max. Casting penalties.
	\item \nameref{spell:stinking-cloud} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}): 20-ft sphere (C, 1 min), 90 ft. CON or waste action. Heavy obscuration.
	\item \nameref{spell:unbind} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:shaman}): 120 ft. Attempt to end one spell. If base cost 5 or lower, automatic success.
	\item \nameref{spell:wall-of-thorns} (\nameref{class:ranger}, \nameref{class:shaman}): Special, 120 ft (C, 10 min). Shapeable wall. Extra dfficult terrain + damage to move through. Provides total cover.
	\item \nameref{spell:wind-wall} (\nameref{class:priest}, \nameref{class:shaman}): Special (C, 1 min), 120 ft. 3d8 bludgeoning for creatures in area (STR for half). Blocks small or smaller flying creatures and projectiles.
\end{itemize}

\subsubsection{8-11 Aether}
\begin{itemize}
	\item \nameref{spell:arcane-eye} (\nameref{class:arcanist}, \nameref{class:spellblade}):
	\item \nameref{spell:banishment} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:spellblade}):
	\item \nameref{spell:black-tentacles} (\nameref{class:arcanist}, \nameref{class:spellblade}):
	\item \nameref{spell:confusion} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:spellblade}):
	\item \nameref{spell:conjure-mephits} (\nameref{class:arcanist}, \nameref{class:shaman}):
	\item \nameref{spell:conjure-woodland-beings} (\nameref{class:ranger}, \nameref{class:shaman}):
	\item \nameref{spell:control-water} (\nameref{class:priest}, \nameref{class:shaman}):
	\item \nameref{spell:death-ward} (\nameref{class:oathbound}, \nameref{class:priest}):
	\item \nameref{spell:dimension-door} (\nameref{class:arcanist}, \nameref{class:spellblade}):
	\item \nameref{spell:dominate-beast} (\nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}):
	\item \nameref{spell:faithful-hound} (\nameref{class:arcanist}):
	\item \nameref{spell:fire-shield} (\nameref{class:priest}, \nameref{class:shaman}):
	\item \nameref{spell:freedom-of-movement} (\nameref{class:oathbound}, \nameref{class:priest}):
	\item \nameref{spell:giant-insect} (\nameref{class:ranger}, \nameref{class:shaman}):
	\item \nameref{spell:greater-invisibility} (\nameref{class:arcanist}, \nameref{class:ranger}):
	\item \nameref{spell:hallucinatory-terrain} (\nameref{class:arcanist}):
	\item \nameref{spell:mass-cure-wounds} (\nameref{class:oathbound}, \nameref{class:priest}):
	\item \nameref{spell:phantasmal-killer} (\nameref{class:arcanist}, \nameref{class:spellblade}):
	\item \nameref{spell:polymorph} (\nameref{class:shaman}):
	\item \nameref{spell:resilient-sphere} (\nameref{class:priest}):
	\item \nameref{spell:spirit-guardians} (\nameref{class:priest}):
	\item \nameref{spell:stoneskin} (\nameref{class:priest}, \nameref{class:shaman}):
	\item \nameref{spell:stone-shape} (\nameref{class:shaman}):
	\item \nameref{spell:wall-of-fire} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:shaman}):
	
\end{itemize}

\subsubsection{12+ Aether}
\begin{itemize}
	\item \nameref{spell:antilife-shell} (\nameref{class:spellblade}):
	\item \nameref{spell:blade-barrier} (\nameref{class:priest}):
	\item \nameref{spell:chain-lightning} (\nameref{class:ranger}):
	\item \nameref{spell:cloudkill} (\nameref{class:arcanist}):
	\item \nameref{spell:cone-of-cold} (\nameref{class:arcanist}, \nameref{class:shaman}, \nameref{class:spellblade}):
	\item \nameref{spell:conjure-elemental} (\nameref{class:arcanist}, \nameref{class:shaman}):
	\item \nameref{spell:contagion} (\nameref{class:priest}, \nameref{class:ranger}, \nameref{class:shaman}, \nameref{class:spellblade}):
	\item \nameref{spell:dispel-otherworldly-influence} (\nameref{class:oathbound}, \nameref{class:priest}, \nameref{class:shaman}):
	\item \nameref{spell:divine-wrath} (\nameref{class:oathbound}, \nameref{class:priest}):
	\item \nameref{spell:flame-strike} (\nameref{class:priest}):
	\item \nameref{spell:hold-monster} (\nameref{class:arcanist}, \nameref{class:priest}, \nameref{class:ranger}, \nameref{class:spellblade}):
	\item \nameref{spell:insect-plague} (\nameref{class:shaman}):
	\item \nameref{spell:mislead} (\nameref{class:priest}, \nameref{class:spellblade}):
	\item \nameref{spell:move-earth} (\nameref{class:shaman}):
	\item \nameref{spell:passwall} (\nameref{class:shaman}):
	\item \nameref{spell:reincarnate} (\nameref{class:shaman}):
	\item \nameref{spell:seeming} ():
	\item \nameref{spell:sunbeam} \nameref{class:priest}:
	\item \nameref{spell:telekinesis} (\nameref{class:arcanist}, \nameref{class:spellblade}):
	\item \nameref{spell:true-seeing} (\nameref{class:oathbound}, \nameref{class:priest}):
	\item \nameref{spell:wall-of-ice} (\nameref{class:shaman}):
	\item \nameref{spell:wall-of-force} (\nameref{class:arcanist}):
	\item \nameref{spell:wall-of-stone} (\nameref{class:priest}, \nameref{class:shaman}):
\end{itemize}

\subsubsection{Legendary}
\begin{itemize}
	\item \nameref{spell:abjure-aether-manipulation} (Generic, Greater)
	\item \nameref{spell:animal-shapes} (Primal, Greater)
	\item \nameref{spell:anyspell} (Generic, Greater)
	\item \nameref{spell:arcane-hand} (Arcane, Lesser)
	\item \nameref{spell:arcane-sword} (Arcane, Lesser)
	\item \nameref{spell:circle-of-death} (Arcane or Divine, Greater)
	\item \nameref{spell:call-divine-servant} (Divine, Lesser)
	\item \nameref{spell:call-fey-ally} (Primal, Lesser)
	\item \nameref{spell:contingency} (Arcane, Greater)
	\item \nameref{spell:control-weather} (Primal, Greater)
	\item \nameref{spell:disintegrate} (Arcane, Lesser)
	\item \nameref{spell:divine-word} (Divine, Lesser)
	\item \nameref{spell:domination} (Generic, Greater)
	\item \nameref{spell:etherealness} (Arcane, Lesser)
	\item \nameref{spell:flesh-to-stone} (Primal, Lesser)
	\item \nameref{spell:glibness} (Generic, Greater)
	\item \nameref{spell:globe-of-invulnerability} (Arcane, Lesser)
	\item \nameref{spell:harm} (Arcane or Divine, Lesser)
	\item \nameref{spell:heal} (Divine, Lesser)
	\item \nameref{spell:heroes-feast} (Divine, Lesser)
	\item \nameref{spell:holy-aura} (Divine, Greater)
	\item \nameref{spell:irresistible-dance} (Arcane, Lesser)
	\item \nameref{spell:mass-heal} (Divine, Greater)
	\item \nameref{spell:maze} (Arcane, Greater)
	\item \nameref{spell:meteor-swarm} (Arcane or Primal, Greater)
	\item \nameref{spell:mind-blank} (Arcane or Divine, Greater)
	\item \nameref{spell:mirage-arcane} (Arcane, Lesser)
	\item \nameref{spell:modify-memory} (Arcane, Greater)
	\item \nameref{spell:power-word-kill} (Arcane or Divine, Greater)
	\item \nameref{spell:power-word-stun} (Divine, Greater)
	\item \nameref{spell:regenerate} (Primal, Lesser)
	\item \nameref{spell:reverse-gravity} (Arcane or Primal, Lesser)
	\item \nameref{spell:rip-soul} (Arcane, Lesser)
	\item \nameref{spell:shapechange} (Primal, Greater)
	\item \nameref{spell:sunburst} (Divine, Greater)
	\item \nameref{spell:symbol} (Generic, Lesser)
	\item \nameref{spell:time-stop} (Generic, Greater)
	\item \nameref{spell:weird} (Arcane, Greater)
	\item \nameref{spell:wind-walk} (Primal, Lesser)
	\item \nameref{spell:word-of-recall} (Divine, Lesser)
\end{itemize}




