\documentclass[10pt,twoside,openany,bg=none]{dndbook}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{import}
\usepackage{mdframed}
\usepackage{caption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage[section]{placeins}

\hfuzz=1pt

\extrafloats{1000}

\hypersetup {
    colorlinks = true,
    urlcolor = blue,
    linkcolor = blue,
    citecolor = red
}

\makeatletter
\newcommand{\smartnameref}[2]{%
    \@ifundefined{r@#1}%
      {#2}%
      {\nameref{#1}}%
}
\makeatother

\title{NIH GM Guidance}
\author{Ben Hall}
\date{}
\begin{document}
\maketitle
\section{How to use this document}
This is an auxiliary document for Game Masters (GMs) of the NIH TTRPG system. It's not needed to play the game, but is designed to answer questions GMs might have and to provide additional helps, including guidance on treasure distribution, encounter building, setting DCs for checks, etc. There is no expectation that players will or will not read it. It does not contain \textit{rules}; everything here is couched as \textit{guidance and suggestions}. The tone is intended to be much less formal than the main play document, as it's friendly advice from one long-running GM to another, as well as a frank discussion of how I, as the designer, was thinking the game would be run. Feel free to deviate from this guidance if your table needs something different. That's a core principle of NIH: no one can know better what your table needs than you do. You're the final authority on that subject. As a game designer, all I can do is point out pitfalls I've seen and solutions that have worked at other tables. Consider this document to be a toolkit.

\subsection{Core Principles}
NIH is built with several principles in mind. These are described below. None of these are hard and fast rules, but they're where the rules were intended to point, as well as preconditions to make the rules work well.
\begin{itemize}
    \item[] \textbf{he focus is the PCs.} This is the story of the player characters and their heroics (or heroic failures). They are the ones shaping the narrative by their actions and the consequences of those actions. This means that it's not \textit{your} story that's being told, although you're a critical part of getting it told. Planning for a good NIH adventure works best when focused on scenarios ("The princess has gone missing!"), factions (the evil vizier and his minions, the guards, etc), and their interactions, rather than on specific story beats or events. 
    \item[] \textbf{You're in charge of the rules.} The designers have worked hard to provide rules that work together well in as many circumstances as they could. But there will always be cases where the rules just don't fit the needs of your party. When you find the need to alter or disregard a rule, you should feel comfortable doing so, as you're the one closest to the action. There are no rules police that will come after you, after all. That said, most rule changes go best when everyone discusses and agrees on what needs to change and why. Sometimes you'll need to make a ruling and move on--don't forget to follow up with helping everyone feel confident in that decision after the session (or earlier). Guidance about what rules are "load-bearing" and will have large consequences if disregarded or modified will be provided later in this document.
    \item[] \textbf{You are a player too and so deserve to have fun.} A GM is a critical moving piece of the NIH system. Very few things kill games faster than the GM getting burnt out or losing interest or feeling like they're just going through the motions. Work with your party if you start to feel that way. Talk out of character. Make whatever changes are necessary. Don't feel like you need to sacrifice yourself on the altar of their fun--come to a mutually beneficial solution or stop playing.
    \item[] \textbf{Player agency is important.} Agency is the feeling that your decisions matter. When players don't feel like they have agency, they tend to check out or start acting badly, trying to break things to regain a sense of agency (or as a form of protest). And this is generally bad. Player agency is maintained when players make informed decisions with actual consequences. It doesn't mean they have to have \textit{total} freedom to choose anything, and they certainly don't get to choose the consequences. But throughout the game, they have to be able to make choices that matter. If you're going to do a set piece that requires them to fail, get them to agree to it out of character first. Or choose a different path that doesn't require them to fail to move the story along. Otherwise you'll probably end up with a broken story (players are notorious for \textit{not} generally wanting to fail and doing all sorts of things to get out of it) and you'll hurt your relationship with the players if you try to force the issue.
    \item[] \textbf{The PCs are working together against the enemies, not each other.} NIH is not designed for player-vs-player action. It's designed for the party working as a (mostly) cohesive group against monstrous foes and challenges. But remember that many abilities do affect everyone and friendly fire is turned on by default. I generally add an out of character breakpoint when someone is going to, for example, hit another PC with a \textit{fireball} spell: I check with the threatened PC's player. If they're not ok, I ask the other player to choose a different target or different spell. PvP actions only resolve if all affected players are ok with it. Stealing from the group or getting the group in serious trouble (such as by attacking a king in his court) are, in my mind, pvp actions. This doesn't mean the players or their characters can't disagree about a course of action; that's totally normal and expected, but in-game actions that harm other members of the party are a recipe for hurt feelings and broken games.
    \item[] \textbf{The monsters are the enemy, not you.} You will be called on to portray antagonists. That does \textit{not} mean you yourself should be antagonistic to the player or the PCs. Monsters may lie to the PCs. You, acting as players' window to the world, shouldn't lie to the players. The characters may not be observing the real truth (due to failed rolls or illusions or whatever), but your job is to accurately portray what their characters experience. Similarly, you shouldn't act out of malice or desire to hurt the characters or the players. The monsters certainly want to hurt the characters, but that's their problem, and something the rules handle.
    \item[] \textbf{The monsters generally lose.} NIH is a game about heroes. And the basic assumption is that the heroes will win, eventually. Not all the time, but most fights should end up going their way. When they lose, it should be because they messed up, made bad choices, didn't do their research, etc. Not because you threw a "rocks fall, everyone dies" no-win scenario. And loss may or may not mean \textit{death}. An average party will face dozens if not hundreds of fights over their campaign. If they had an average fatality rate of even 5\%, the chances of having no deaths is under 75\% by the 6th fight and under 35\% by the 20th. And many campaigns will have 6 fights \textit{within the first few days of adventuring}. NIH is not designed as meat-grinder.
    \item[] \textbf{Out of character problems need out of character solutions.} If someone is causing issues, talk to them. Seriously. And kindly. Not accusing, not angry. Trying to "punish" their character for the player being bad rarely, if ever, works. It usually just makes things worse for everyone via escalation.  Solve out of character problems out of character. The game system isn't designed to help you solve people problems. No rules can stop a malicious actor. In the end, it comes down to only playing with people you trust and trusting the people you play with.
    \item[] \textbf{NIH is a game about heroes, not villains.} This is a corollary to the principle above about the monsters generally losing. NIH presumes that the player characters, in their role as protagonists, are trying to make the world generally a better place. Nothing in it is designed for evil parties. This is why you won't see things like player necromancers (or in fact any player-facing way to create undead), and why player-facing charm and domination effects have been toned down compared to similar games. This is mostly a tone thing. The game won't actively \textit{break} if people are bad. But you'll get tonal dissonance. The intended setting is one where there are major threats, but where heroes working together can genuinely make the world a better place (even if only temporarily until the next threat). Monsters exist (including ones in humanoid shape), but the PCs shouldn't \textit{act} like monsters.
    \item[] \textbf{Player characters and monsters are built different but interact the same ways.} NIH is built around player-non-player asymmetry. This means that monsters (including humanoid NPCs) are built along different lines. Monsters don't have classes, official lineages, backgrounds, skill tricks, spell lists, etc. Each one only has the pieces necessary to fit that character's role. And you can mix and match these to your heart's content. In fact, building monsters out of PC pieces generally doesn't work very well and is not recommended. However, monsters and PCs do interact using the same rules for action economy (what actions they can take and how many), hit points, armor class, etc. They generally don't have aether or stamina and instead have fixed recharges for the sake of simplicity.
\end{itemize}

\subsection{Layout of this document}
This document is divided into several sections that each cover a specific chunk of running the game.
\begin{itemize}
    \item This current section discusses how to use this document.
    \item \nameref{sec:resolving-checks} discusses the role of dice in resolving checks (attack rolls, ability checks, or saving throws). It discusses different ways to set DCs, and what the DC values are intended to mean, as well as gives a few more examples as to when one proficiency might be more appropriate than others. It lightly touches on non-binary resolutions (aka degrees of success/failure) and when you'd want (or not want) to use those.
    \item \nameref{sec:encounter-design} discusses different models for designing both individual encounters or challenges as well as "adventures" (strings of connected challenges). It discusses particular gotchas commonly encountered, and gives some guidance on manipulating the difficulty to be what you want it to be.
    \item \nameref{sec:modding-the-rules} discussion some considerations for making the rules fit your table better. Both rules that have large consequences if altered as well as some variants for specific cases.
    \item \nameref{sec:treasure} discusses the expectations for treasure distribution, as well as presenting ways to randomly generate "appropriate" treasure.
    \item \nameref{sec:worldbuilding} discusses the system's expectations around the world you play in. There is a default setting for NIH, but it's not binding.
    \item \nameref{sec:sample-adventure} presents a sample (short) adventure for level 1-3 characters, designed to quickly introduce people to the system, especially for new GMs. Although it takes place in Quartus, it is designed to be "low lore" and transplantable into any setting.
\end{itemize}

\section{Resolving Checks}\label{sec:resolving-checks}
The mechanical core of NIH is using a d20 to resolve \textit{uncertainty}. If it's already clear from the fiction what should happen, adding dice rolls is at best fluff and at worse confusing. Now, rolling dice is fun for many people. And I don't want to discount that. But in general, the safest guiding principle behind any attack roll, ability check, or saving throw is "is this a situation where both success and failure are plausible outcomes and, more importantly, \textit{interesting, acceptable} outcomes with interesting consequences that move the narrative along." If the answer to either of those questions is no, \textit{don't roll. Just narrate.} This takes some finesse on the GM's part. There are differences between the different kinds of checks, so let's look at those in more detail.

\subsection{Attack Rolls}
NIH is a game about heroes and villains worthy of heroes. So most of the time, if the task is "can I damage that creature", there is meaningful uncertainty. Even a completely unaware enemy might flinch out of the way at a critical moment. Even if asleep. And even attacking an object under stress can fail. And with attack rolls, a natural 1 is a miss and a natural 20 is a hit, regardless of the AC. So generally, a player asking to attack something will be resolved with an attack roll.

There are exceptions. A party or NPC might decide to kill creatures that, in the fiction, pose them no threat and that they should be able to defeat without problem (such as a villain kicking a puppy). In this case, there's little reason to spend the extra time to roll the attack; we can just simply move along to the consequences of them hitting. Similarly, a character trying to break an enchanted adamantine wall with a wooden toothpick doesn't require an attack roll; he simply cannot do that. Simply narrate the failure of the attack. There is one situation where rolling might be useful in cases like this: to decide \textit{how messy} the successful attack is or \textit{how bad} the failure is. That's covered in \nameref{subsec:degrees-of-success}.

\subsection{Saving Throws}
Nobody likes feeling powerless. It's a rare circumstance that a PC will not be able to make a saving throw against an effect. They will be able to at least attempt the saving throw unless they are subjected to a condition that imposes automatic failures (such as \smartnameref{condition:paralyzed}{the Paralyzed condition} against Dexterity saving throws). And the consequences are almost always meaningful: death, damage, or a disabling condition. As a result, saving throws are almost always useful to roll and play out.

There are a few exceptions, discussed below. In all of them, the core principle is \textit{to keep the focus on the party and keep the narrative moving.} NIH is not a physics engine; it's a toolkit designed to help you bring narratives to life and figure out what happens to the party and their goals.
\begin{itemize}
    \item If there isn't anything threatening the party and they're not under time pressure and the condition naturally resolves given time. For example, a slippery slope where the only consequence is falling prone, but no other threat looms. Since all this does is slightly slow down the party, rolling might not be all that important. Just narrate how they slip around and have to move carefully and move on to more important things. You can even throw in narrative that emphasizes that it's harder for the less dextrous party members.
    \item If a damaging effect will do more than the monster's hit points even if halved, don't bother rolling saving throws for them. They can't survive even if they pass the saving throw. This is particularly common at higher levels facing off against a horde of small creatures such as bandits targeted by \textit{fireball}.
    \item If it's just set dressing and the party isn't directly involved. A dragon doing a drive-by flaming of a camp of guards and commoners just kills however many of them as is thematically appropriate. Important NPCs may get saves if narratively appropriate.
\end{itemize}

\subsection{Ability Checks}
Unlike attack rolls and saving throws, ability checks are rarely the result of a fixed ability and thus require more GM adjudication. The same basic principles (only roll when the outcome is uncertain, both results are acceptable both in the narrative and to the table's culture, and both results are interesting) apply, but there's the additional need to set a DC. At the end of the day, the exact value you decide for the DC isn't \textit{that} important, as long as it's reasonable. Don't worry too much about distinguishing a 16 from a 17; it's totally fine to only ever use one of four options (and the fourth very rarely): 10, 15, 20, and 25 (discussed below).

All actions requre some information to resolve. These are generally summarized as Intent, Approach, and Details. The first two are always required, the last is included for supplimentary clarification. The acting character's Intent is what they're trying to get by doing the action. For most actions, this is fairly obvious on its face. The character wants to jump over a gap, wants to force the door open, wants to remember information about a particular environmental cue, etc. For others, especially social interactions, the Intent is much less clear on its face. Is the character trying to improve their reputation? Get a particular piece of information, and if so which one? Get past the guard? Etc. The Intent anchors the outcome if the character succeeds on the check. The Approach is \textit{how} they're doing it. Again, for many actions, there just isn't much ambiguity here. But it informs things like "does the character have a running start? Are they using any resources? Are they trying to sweet-talk the guard? Or threaten them? Or maybe trick them?". The Approach anchors the requisite ability scores and proficiencies used in the check (if any), as well as materially affecting the DC. It can also affect whether the check is needed at all, as some Intents cannot be achieved by certain Approaches...and some Intents are automatically achieved if the right Approach is used. Details provide any necessary backing information. If you're trying to bribe the guard, how much are you paying? Are you using equipment? Do you have a book on the subject? Etc. These generally go into setting the DC and possibly providing advantage or disadvantage.

For social scenarios in particular, how much you need the players to talk in first person and provide their actual wording is up to you. All the system and this advice require is a statement of Intent and Approach and any relevant Details. This could be third person "Grog wants guard to move aside; he gets all in his face and tries to scare him off". Or it could be first person, with full dialog, like "I walk up to the guard and loom in his face, staring down with narrowed eyes. 'Move', I rumble, trying to scare him off". Which convention you use is up to you and your table. Both have advantages and disadvantages. The only cautionary note is that giving players benefits (lower DC or advantage) for being more descriptive tends to allow people to cheese the system by having low scores but getting the benefit of higher ones, while also making people who aren't as comfortable with ad lib dialog not wanting to play high Charisma "face" characters. Two actions with the same Intent, Approach, and Details should be treated the same.

\subsubsection{Resolving Actions Without Fixed Features}
When there isn't a fixed character or monster feature involved in a player or monster's actions, there's a short process you'll have to go through to resolve the action and decide what happens next.
\begin{enumerate}
    \item Decide if any mechanical resolution is actually necessary. Use the core decision tree below as a starting point.
    \begin{enumerate}
        \item Is the proposed action's Intent impossible? If so, and if the character would reasonably \textit{know} that (even if the player doesn't), I recommend informing the player of that fact and letting them do something else without wasting their action or resources. If the character wouldn't know, simply resolve the attempt as a failure. This should be used sparingly. For things that aren't logical or physical impossibilities (and are merely unlikely), it's often better to let them try and set a high DC. Note that if there is a range of outcomes for failure that could depend on how well you do at an otherwise impossible task, it might be better to use a non-binary check (described below).
        \item Is the proposed action unlikely to fail, given the narrative circumstances? Alternatively, would it be a breach of established characterization for that routine action to fail? For example, a professional chef will rarely, if ever, have a routine dish (one they cook frequently) just fail. And it would be weird if even 5\% of the time a trained sailor couldn't balance on the mast during normal weather conditions. In these cases, simply resolve the action as a success and move on. This one can be used liberally. Players love feeling important. Especially if narrated that this is something most people would have to roll for, but this character is just that awesome. Similarly to above, if there are a range of success options that depend on how well you do (but no failure case), it might be better to use a non-binary check (in this case degrees of success, rather than degrees of failure).
        \item Are there actual consequences that matter and are interesting for both success and failure? This combines a few factors. If there isn't any factor preventing the person from trying again until they succeed and time \textit{doesn't} matter...simply let them succeed. If time \textit{does} matter, but there aren't any other meaningful consequences of failing other than having to try again, a simple ruling is to allow them to automatically succeed at the cost of taking 10 times as much time as normal (defaulting to taking 1 minute, or 10 rounds).
        \item Are you ok with the outcomes of both success and failure? If not, simply narrate whichever one you \textit{are} ok with or change the outcome to be something you can be ok with. In this case, we don't want to let the dice or the rules override our own knowledge of what is fun. \textbf{CAUTION:} Using this to force your preferred narrative denies players their agency and is one form of railroading. This lever is best only pulled when something critically un-fun (as judged by your party) would happen. For many parties, an example might be an ignominious, pointless death for something simple. Instead of a party member falling to their death on a failed Dexterity (Acrobatics) check to land safely from a risky jump (narratively slipping off the edge of a huge cliff), instead they might succeed at balancing (success on the check) at a cost: maybe they lose something from their pack, take an injury (mechanically implemented by a level of exhaustion, etc), or something like that. Similarly, if a party would get deadlocked on a quest if they failed an Intelligence (Investigation) check and not be able to make any more progress, don't have them make that check. They simply find the needed information. Find a different way of testing them.
        \item Is this something the person is doing routinely, rather than a focused effort? Use a passive check (10 + the relevant modifiers) instead of an active one. The most common use of this is Wisdom (Perception) to avoid being surprised.
    \end{enumerate}
    \item Once you've decided that a check is, after all, necessary, you need to decide \textit{what kind of check} is needed. This involves picking an ability score and possibly one or more sources of proficiency to add. There is no extra bonus for a character having proficiency from more than one source on a given check, generally.
    \item Now it's time to pick a DC.
    \item Now determine whether there are other circumstances, not rolled into the DC, that either give the PC a boost or a penalty (ie decide whether the check is at advantage or disadvantage). Examples include having someone else Help (granting advantage) or wearing certain types of armor while trying to sneak (disadvantage on the Dexterity (Stealth) check).
    \item After that, you resolve the check by having the player roll and comparing the result to the DC, as usual.
    \item Finish up by narrating the results and handling any mechanical consequences (lost hit points, death, consumed resources, etc.)
\end{enumerate}

Note that a natural 1 or a natural 20 have no meaning for ability checks \textit{other} than the worst or best possible result. If the DC ends up being set to 25 (a possible but very difficult task), a character with a +4 cannot succeed on the check. Conversely, if the DC ends up being set to 5 (a task that could plausibly fail but that is quite easy), a character with a +4 cannot fail (4 + 1 = 5) on the check. And that's normal.

Remember that a number of things are listed as not generally requiring checks at all, such as jumping, riding a normal mount, climbing, or swimming. As a rule of thumb, if it's something that a normal (normally abled, normally in-shape, but not specifically trained for the task) person can reliably do in real life, it shouldn't require a check from an adventurer. Adventurers are above the norm and shouldn't have to roll for mundane things.

\subsubsection{Ability Scores and Checks}
Ability checks only optionally include proficiency in a skill, tool, or something else. Those that allow proficiency are usually stated following the pattern of Strength (Athletics), with the ability score coming first (Strength) and any applicable sources of proficiency in parentheses. Proficiency is not a gate to making the check, merely an added bonus if present. A Strength check, without the listed proficiency, wouldn't add any proficiency (and thus would not gain any benefit from features like a rogue's Expertise). That said, \textit{most} checks you make on a day-to-day basis will involve some source of proficiency, often from a skill. While skills are tied to an ability score by default (such as Athletics defaulting to Strength), it is totally fine to apply the proficiency to a check for a different ability score if that makes more sense. The canonical example is using Intimidation (normally a Charisma skill relying on verbal threats) with Strength. This might represent a \textit{wordless} intimidation check, such as by bending the target's weapon in half with your bare hands or casually hefting a huge boulder over their head. There are a few traps to avoid, however. Among them are the following.
\begin{enumerate}
    \item Letting characters always use their focused ability score. For example, letting the rogue jump longer distances using Dexterity (Acrobatics) instead of Strength (Athletics). This cheapens the game by not having consequences for different choices. Some ability scores are used more than others during normal play, often because they're tied to more features and skills. For example Dexterity is referred to more than Strength. By letting people do things with Dexterity that should be done with Strength, it provides incentives to use Dexterity and leave Strength out, which harms those classes that are focused on Strength.
    \item Being too rigid about what applies. The GM's role is to decide what checks are called for. Players do not call for checks, they describe actions. On the flip side, you should be willing to negotiate with players if they want to, for example, substitute one source of proficiency for a different one. A tactic to use here is to change the narrated result depending on which proficiency they rely on. This is easiest for Intelligence-based "lore checks": give them different information for an Intelligence (History) check than an Intelligence (Religion) check. On both, a success should give meaningful (in the context of the check) information, but the exact information may be different depending on what they're relying on.
\end{enumerate}

\subsubsection{Deciding the DC}
The higher the DC, the harder the check. In general, DCs in NIH are context dependent, including who is making the check. This does not mean that the DC is higher for a character with a low ability score (or vice versa). But a noble-born character interacting with nobility may be the only one who \textit{can even possibly make the attempt}, even if they have a low Charisma. Or, if that particular character has annoyed the NPC in question, the DC to get their desired result may be higher. And DCs are generally not constant because it's really rare that the party attempts to perform the same task under the same exact circumstances multiple times. That said, it is generally bad play to adjust the DC depending on the modifier of the person attempting it. The rogue having a +17 to Dexterity (Sleight of Hand) checks shouldn't imply that all such checks should be much harder to "provide a challenge" to the rogue. It's totally ok for tasks to be trivial for one person and difficult for another. That's what class features are for.

Note that setting the DC comes \textit{after} you've decided that a check is necessary. Attempting to do an impossible task has a DC of SUCCESS NOT POSSIBLE, not any specific numeric DC. Similarly, attempting to do a task that has no risk or is otherwise an automatic success has a DC of FAILURE NOT POSSIBLE.

The system expects that DCs will be in a range between about 5 and 30, although most of the time, useful DCs are between 10 and 20. DCs below 10 usually indicate that a task should probably have been an automatic success \textit{unless} the consequences of failing it are so severe that even that small chance of failure is enough. And usually that is only even a possibility if someone in the party has a low or negative modifier. The three most common DCs will probably be 10, 15, and 20. Remember that DCs for tasks without proficiency are between 10 percentage points and 60 percentage points (accounting for possible expertise) more difficult than those without proficiency.

Generally, the system works best when the probability of success on most checks is approximately 50\% ($\plusminus$ 10-20\%). Too much higher and things like advantage provide less value. Too much lower and you're rolling in vain, which can feel disheartening. These other ranges of success chance still happen, but if it's happening regularly, you might want to reconsider how you're setting DCs.

The following discussion and tables assumes as a baseline three different "personas" to help give intuition on what the various DCs mean. One is a "commoner", meaning they have +0 to all ability scores and no proficiency, leading to a consistent +0 modifier. This represents both the unspecialized PC and the "common man" (although see below for using checks for NPCs). The next is someone with total modifier of +4, whether from proficiency, ability score, or a combination of both. This represents the character of "normal skill," as well as a task that does not involve proficiency being performed by someone specialized in that attribute. The third has a modifier of +11, the maximum possible without expertise or other magic. This represents a high-level specialized character.

\begin{DndTable}[DCs and Chance of Success]{Xccccc}
    Modifier & DC 5 & DC 10 & DC 15 & DC 20 & DC 25 \\
    +0 (Disadvantage) & 64 & 30 & 9 & 0.25 & 0 \\
    +0 (Normal) & 80 & 55 & 30 & 5 & 0 \\
    +0 (Advantage) & 96 & 79.75 & 51 & 9.75 & 0\\
    +4 (Disadvantage) & 100 & 56.25 & 25 & 6.25 & 0 \\
    +4 (Normal) & 100 & 75 & 50 & 25 & 0 \\
    +4 (Advantage) & 100 & 93.75 & 75 & 43.75 & 0 \\
    +11 (Disadvantage) & 100 & 100 & 72.25 & 36 & 12.25 \\
    +11 (Normal) & 100 & 100 & 85 & 60 & 35 \\
    +11 (Advantage) & 100 & 100 & 97.75 & 84 & 57.75
\end{DndTable}

Note that advantage and disadvantage don't make impossible things possible (or vice versa), they merely make possible things more likely/unlikely (respectively). Higher modifiers make a dramatic difference; someone with a +4 at disadvantage has about the same chance of success on a DC 10 check as someone at +0 rolling normally and only an approximately 5\% lower chance on a DC 15 or 20. Effectively, "disadvantage" is worth a variable amount between -4 and -5, except that something that cannot fail for a higher modifier isn't now failable with disadvantage. Similarly, advantage is worth a variable +4 or +5, \textit{except} that it doesn't actually change 0\% success chances to something doable.

Most people aren't sensitive to small percentage-point changes in probability. As a result, it's usually not worth trying to be more specific than about steps of 5. It is totally fine to pick from the following (all assuming that people are going to do things they're good at, so using the "normal skill" persona):
\begin{itemize}
    \item A very easy check that a commoner might fail but a normally-skilled person shouldn't has \textbf{DC 5}. Use these sparingly; prefer to just give automatic success instead where that makes any sense.
    \item An easy check that a normal-skill person should only fail rarely has \textbf{DC 10}. These probably the most common.
    \item A check that a normal-skill person (\textit{not a commoner}) should fail about 50\% of the time has \textbf{DC 15}. Usually the second-most-common DC.
    \item A check that should be difficult for a normal-skill person (in the absence of advantage or magical help) has \textbf{DC 20}. Less common than DC 10 or 15, but still quite common.
    \item A check that should require magical help to even be possible for a normal-skill person (but isn't fundamentally impossible given that assistance) has \textbf{DC 25}. Use these sparingly, although there's generally much more call for these than for DC 5 checks.
\end{itemize}

\subsection{Passive vs Active Checks}
Some of the time, the GM needs to determine whether an adventurer succeeds at something without the \textit{player} saying that they're particularly doing anything. Most often, this is in connection to one of two cases: routine actions or secret information. Routine actions are things the adventurer is just doing all the time, effectively standing orders. People are presumed to be normally watchful of their surroundings while not distracted or actively doing something that requires their focus. As a result, there's a large number of things they might be able to see. Instead of constantly rolling Wisdom (Perception) checks, you can summarize this by comparing the character's Passive Perception score (calculated as 10 + the character's Wisdom modifier + their proficiency if proficient in Perception) to the DC of the trap or item (or especially hiding creature).

Similarly, if there's a question about something that the player shouldn't necessarily know, where the act of asking for a check might give away information and change behavior (a type of metagaming), the DM can choose to use a passive check for a different ability score or proficiency source (calculated similarly to passive Perception).

Note that these are passive \textit{from the perspective of the player}, not from the perspective of the character. The character is actively looking around; if they're \textit{not} (because they're unconscious or busy doing something else), they don't get the benefit of any check and automatically fail any necessary check. But the player isn't actively making a check.

Other than passive Perception, these checks come up infrequently at most tables. Passive Perception is the exception because it's used to determine whether a creature is surprised when initiative begins.

\subsection{Non-binary Check Results}
Also called "degrees of success or failure", non-binary checks represent cases where the uncertainty isn't in whether the action succeeds or not, but \textit{how good or bad the results are}. For example, a PC asking a king to give up their throne (without something like a victorious army backing them up or similar \textit{force majure}) is always a failure. But someone who is particularly charming \textit{might} avoid having the guards throw them and the party into jail or kill them and instead merely be ejected from court and/or laughed at. Here, the Charisma (Persuasion) check might have a DC of 10 to soften the consequences at all (anything below a 10 is "guards try to take them to jail") and for every 5 above they get slightly softened consequences. To be clear, no check result can end with them in control of the kingdom. At best, they might get the king to find them amusing and throw a few copper their direction (ie no significant consequence). On the other side (degrees of success), many checks about "what do I know about \textit{topic}" will never result in "I don't know anything." Characters can be presumed to have heard rumors, legends, etc. about various monster types, for example. In this case, a higher result might result in either more information or more relevant information. A check result below 10 might be "trolls are very hungry all the time" (a true, but not very useful fact), while a 15 might be "trolls are known to fear fire and acid", while a result of 20 might be more like "trolls will regenerate from anything unless their wounds are cauterized with fire or acid. They are faster and smarter than they look."

I prefer to err on the side of more information rather than less, so I tend to be generous with checks made to gain knowledge. But each table can decide what level of mystery they're shooting for.

\subsection{Group Checks and Piling On}
One very common question is "who rolls this check?" Alongside that, a common temptation for parties is to do "pile in" checks, where one person attempts, rolls low, and then others go "can I roll as well"? Letting everyone roll, just like always letting the specialist roll, both result in a distinct decrease in difficulty. If each person has a 50\% chance of success, two people rolling is like rolling at advantage. Three makes it almost a guaranteed success.

Two tactics to tame this tendency are group checks and the Help action. If the task in question is something where more assistance isn't always better (such as a group of people debating a topic) but the party has the opportunity to work together and cover for each other, you can use a Group Check. Here, have everyone roll, and the check is a success if half the party or more succeeds. Examples include decoding inscriptions on a wall as part of solving a puzzle (the party can discuss and come to consensus) or helping each other scale obstacles that they'd need to roll for. Here, the party succeeds or fails as a group. Another path is to have the others Help the person actually doing the check, granting them advantage. Holding the light, handing them tools, etc.

In general, well-run checks are not repeatable because there are consequences that happen at the end of the check for both a success and a failure. The PC has already given it their best attempt, and trying again with that method is futile. Either the circumstances must change or the party must change their approach.

\section{Combat Encounter And Adventuring Day Design}\label{sec:encounter-design}
There are several major factors in designing satisfying combat encounters. A lot of this requires finding what your particular party feels to be a fun balance---are they seeking to be challenged and to run the ragged edge of difficulty? Or are they mostly about enjoying the narrative and/or feeling powerful? Or somewhere in between? It's fine to have very hard encounters. It's fine to have very easy encounters. "Balanced" encounters (where each one is the same difficulty) are generally \textit{worse} for table morale and enjoyment than having variation. Variety is the spice of life. And almost everyone wants to feel powerful \textit{sometimes}. If the universe seems to scale with you, it can feel like you're on a treadmill; numbers go up but you never feel more powerful.

\subsection{The Adventuring Day}
The time between long rests while out adventuring is called one adventuring day. Because you can't benefit from a long rest again for 24 hours after you finish the first, one adventuring day is \textit{at least} one calendar day. But some days you'll go longer without a solid long rest (due to the needs of the campaign). And some adventuring days are empty. You may spend 3-4 days travelling along safe roads between two cities without any significant threats. That's normal.

NIH assumes that attrition will occur. It expects that characters will go into many fights without all their resources (whether those are HP, daily uses of abilities, STA, or AET). This only comes by having multiple fights between long rests (per adventuring day). How many? Each table will vary, but a good baseline expectation is to have ~10-12 combat rounds split between 2-4 combats, with at least one short rest available between some of the fights. Note that this is an \textit{average} expectation, not a hard system requirement. Some days may only have one or no fights for a total of 2-4 combat rounds. Others may have 20+ combat rounds. The only \textit{pathological} case is only having one combat regularly enough that the party can count on always having their best abilities ready. There isn't a good way of balancing the system's expectations under those conditions. An encounter difficult enough to pose a siginficant challenge to a party who knows that this is the only thing they have to spend resources on that day is \textit{also} generally difficult enough to wipe out the party if they roll badly (or the enemies roll well). That results in very swingy, difficult to predict encounters. It also strongly weights the scale in favor of spell-casters whose primary resources use AET over those who use STA, since STA users tend to have fewer resources at any given time but get them back faster...which doesn't matter if you only ever have one fight.

This does require designing scenarios that don't regularly allow them to retreat and rest somewhere safe for a day after every fight before moving further. Options for doing so include explicit "doom clocks" (e.g. "the princess will be sacrificed in three days, at the height of the new moon") that require the party to push beyond their comfort zone to fulfil their mission, interrupting rests (ie not having a safe place to rest), reinforcements (so that if they do back off after every fight, they've alerted the enemy and made the situation worse, not better), etc. 

At the same time, the game also expects the PCs to take short rests once or twice per adventuring day (on average). By default, a short rest lasts an hour, which can seem like a lot while in the middle of a dungeon or time-bound quest. If this is often a problem, there are a few workarounds.
\begin{itemize}
    \item If you're playing a particularly high-power game or can give narrative justification (such as a divine blessing), simply reduce the duration of a short rest to 5 or 10 minutes while keeping the length of a long rest the same. This does strengthen the STA-focused classes over the AET-focused ones, so it has to be done with care.
    \item If it's a limited-duration problem (such that most of the campaign will act normally), make sure the party has access to the Apprentice's Friend potion, which grants a PC the benefits of a short rest when consumed over the course of a minute once per day (per dose). This mimics the WWII habit of both sides of dosing soldiers with amphetamines...
    \item Recognize that things aren't quite as fast as they seem. Searching a room thoroughoughly might take 20 minutes if it's small and simple or as much as a few hours if it's full of stuff or being searched in detail. A cursory search might take 1-2 minutes. And remember that they can't long rest again for 16 hours after their last one finishes (so that the next one ends exactly 24 hours later), so it's not as simple as holing up for 8 hours, it's holing up undetected for an entire day.
    \item Explicitly provide limited safe-havens for the party to find, but not safe enough for a long rest. Maybe an area that's only patrolled once per day.
\end{itemize}

\subsection{Designing Good Encounters}
Every encounter has the same core pieces: (a) opponents, (b) a location, and (c) a win condition for each side. Interesting encounters vary all three pieces. Fighting in a bare white room where the only win conditions are "kill the other side" gets old fast. The same goes for always fighting the same types of enemies. An interesting location with terrain that can be interacted with (or even simplly hidden behind) goes a long way to spice things up. And win conditions that aren't simply "kill the other side" require players to think beyond pushing the biggest damage buttons. Things such as "protect an NPC", "amuse a crowd", "stop a ritual", "kill one particular leader", "protect a location", "hold off the enemy until you can escape", etc. work well. I'll talk more about encounter \textit{composition} in a bit, but it's often the least impactful part. 

Another trick is to ham it up. This may not mean narrating every blow, but consider the spectacle that is pro wrestling. If the wrestlers simply got in the ring and fought, \textit{even if they didn't change the fight itself} but didn't have their scenery-chewing personas, it'd be a lot less captivating of a show. Give your monsters personalities. That goblin bumbles and does obviously dumb things, that one has anger issues and attacks people who make them mad, while the other is smart and thinks tactically. But make it more obvious and blatant than you think. Subtlety can get lost in a TTRPG medium. A good bombastic villain or a chilling, sadistic one goes a long way to setting the scene even for a mechanically simple fight.

Also, give your players the reason to care about your enemies. This can be done in multiple ways. For example, people that you want them to double down on and kill with extreme predjudice might steal from (or attempt to steal from) the party or threaten those they care about. That, in my experience, is almost guaranteed to make the party move heaven and earth to take out that NPC. On the flip side, you can make some of them complex, such that the party has to decide whether they \textit{really} want to fight them. And if your party tries to talk to enemies...sometimes let that actually work.

Good environments generally involve variety, with multiple interactable objects or terrain pieces that fit the scene being played out. Cliffs to toss the enemy off (or be tossed off of yourself), pits of lava or acid, ways for the party (or the enemies) to create choke points, things to hide behind (both for stealth and just for cover), etc. This is a great place to use obvious traps; not so much to \textit{hurt} the party, but to force people to make interesting decisions and take risks. And when they capitalize on those environmental factors, let it actually work and work better than just standing and slinging a spell or an attack, but also not be regularly repeatable. Dynamic environments, where walls break down, vents erupt, etc. can be a lot of fun, but also take more management. It's fine to have simple backgrounds some of the time, but as with most things, variety makes for a much better game.

\subsubsection{Difficulty}
The following discussion uses a few labels for different levels of difficulty. All assume a fairly baseline, low-optimization party. Those with higher levels of optimization, tactics, or combat-relevant magic items will punch above their listed weight by at least one step. Being surprised increases difficulty at least one step; being able to surprise the enemy drops the difficulty by at least one, sometimes two steps. Extremely favorable or disfavorable terrain, encounter composition, or environments (including the presence of traps) can lower or raise difficulty by one step.

An \textit{Easy} encounter is unlikely to use any resources at all other than a few hit points or maybe a few Stamina or Aether. Higher-power parties consider these trivial and it might be best to simply narrate the outcome. For example, a mid-level Adventuring Tier party against a few wolves or bandits. It's rarely useful to include these in your adventuring day balance.

A \textit{Medium} encounter will generally take 2-3 rounds and take significant hp from the party's overall pool OR will consume significant (relative to their available pool) resources, whether per-day uses, Stamina, Aether, or consumable items. They are unlikely to take a PC to zero hit points unless they were already wounded, even if the party is fairly unlucky. Luck or tactical skills may significantly reduce the resource consumption.

A \textit{Hard} encounter will generally take 3-5 rounds and runs a significant risk of reducing even a full-health party member to zero hit points. It's unlikely to outright kill them unless the monster gets very lucky or the party gets very unlucky, however, giving a chance to get them back in the fight. Luck or tactical skill can reduce the threat significantly.

A \textit{Deadly} encounter will generally take 3-7 rounds and runs a significant risk of reducing multiple party members to zero and possibly outright killing one or more of them \textit{unless} the party is either lucky or tactically skilled. Even tactical skill cannot entirely remove the risk, however.

\subsubsection{Encounter Composition}
There are a few different general "modes" of combat. At one end, you have a cinematic "boss fight" against a single monster who single-handedly (or clawedly) poses a threat to the party. At the other end, you have waves of disposable minions, each of which poses little threat. A few prototypes, each discussed in more detail below, might include
\begin{enumerate}
    \item The mass minion fight. None of them stronger than the others, but lots of them. Often 2-4\texttimes as many monsters as party members. Fighting a group of goblins or bandits might be an example.
    \item The mirror fight. About as many enemies as party members, each one more complex than a minion but not quite boss level. This one relies on group tactics. Fighting a rival adventuring party is a good example.
    \item The commander and squad fight. Here you have one "boss" and several weaker minions. Generally the boss is much stronger than the minions but not as strong as a true solo fight. A squad of demons with a stronger one in charge is a good example. Or a pirate crew with a captain or warlord at the lead.
    \item The deadly duo (or trio). Fewer enemies than PCs, each of which is strong. A mated pair of manticores against a low-level party might be an example.
    \item The true solo. Here you only have one monster. Generally, this will be a legendary monster (with legendary actions).
\end{enumerate}

In each case, action economy is king. Everyone on the field gets one main action, plus movement and a possible bonus action. Generally, taking two actions is better than taking one action that's twice as strong, in part because attacks can miss and abilities be nullified. Unless a monster is capable of taking down a party member reliably on each action and surviving for multiple rounds under the party's focused fire, it will likely end up being an underwhelming spectacle as a solo monster. This is what legendary and lair actions (off-turn actions that certain monsters can take each round) are for: to partially balance the playing field when outnumbered 4-5:1. However, it doesn't always work, and solo monsters are a risky endeavor; make them strong enough to be a fun fight and you risk it being much more lethal than intended, especially at low-to-mid levels. 

\subparagraph*{The Mass Minion Fight}
This is a standard kind of fight. The sweet spot is between 1.5 and 3 times as many monsters as PCs, each substantially weaker. For example, a level 2 party of four might face 6-10 bandits (offensive CR 1/8, defensive CR 1/8). Each minion dies quickly, but there are a lot of them. This style makes those characters capable of hitting multiple targets feel really special (because they can wipe bunches of enemies per action), but can be quite threatening to the squishier party members who get swarmed. Careful tactics (such as hit-and-run or attack-and-hide) can turn a bunch of goblins or bandits into a threat even against much higher level parties.

If you're going to use a homogenous group (all the same kind of monster), make sure that the monster chosen either has extra movement abilities OR some way of attacking at range (such as a bow or javelin). While it's fun occasionally for the party's ranged characters to kite the enemies around shooting them to death without them getting to act meaningfully...too much of that can be boring for the melee types. Mixed attack modes solves a lot of this. Groups composed of multiple types (such as bow-wielding bandits mixed in with melee bruisers) also help the party make interesting choices.

\textbf{Encounter Difficulty with Mass Minion Fights:} Medium fights have the median offensive CR between 1/4 (at low tiers) and 1/2 the median level of the party members, and the sum of the offensive CRs should be somewhere between 1/3 and 1/2 of the total of the class levels of the party. Hard mass minion fights have no monster at offensive CR $>$ minimum level of the party, but a total between 0.5x and 1x the total of the class levels of the party and total defensive CR around the total party level.

\subparagraph*{The Mirror Fight} This is a hard one to get right. Generally you'll want to have monsters that are roughly on par with party members \textit{in their weakened, partially depleted state}. Because the monsters come in fresh and have all their good stuff available, while the party probably won't. They need to have a mix of capabilities that at least resemble those of the party (even though they don't have class levels): a spellcaster or two, a ranged weapon user, and at least one melee-focused individual. This is a good one to ham up, with plenty of opportunity for quips, barbed comments, and flashy displays of prowess.

\textbf{Encounter Difficulty with Mirror Fights:} If there an equal number of party members and opponents, a Medium encounter has average offensive and defensive CR somewhere about 1/2 the party's average level. A Hard encounter has average CRs around 3/4 of the party's average level.

\subparagraph*{The Commander and Squad Fight} This is a classic "boss fight" that works quite well. Better, in fact, than most solo fights. The boss is stronger than a party member (but not an outright one-round-kill threat), while the minions are weaker. The minions are there to provide breathing room for the commander, who has a few special flashy abilities. This one also provides lots of room for scenery chewing, and the boss's commands, demands, and reactions can also clue party members into what to expect the next round from the minions! If your boss is a spell-caster, melee minions are a great backup. Conversely, if your commander is a heavy melee brute, having support or artillery minion (including those that can heal, counterspell, buff, or hamper the party) makes for more interesting combat.

\textbf{Encounter Difficulty with Commander/Squad Fights:} A Commander whose offensive CR is about the same as the party's average level and 4-6 minions of about 1/8 that CR makes for a Medium fight against 4 PCs. A hard fight keeps the commander the same but either increases the number or power of the minions, but not beyond about 1/4 the party's level. Increasing the commander's offensive CR runs risks of knocking out a party member before they can do anything, which puts the party at an extreme disadvantage.

\subparagraph*{The Deadly Duo/Trio} This is like the Mirror Fight, but with fewer, bigger monsters. Here it's very important to have interesting environmental factors, because otherwise they can get locked down and dead real fast (and anticlimactically).

\textbf{Encounter Difficulty with Deadly Duo Fights:} Two monsters with offensive CR at about the party's average level + 1 or + 2 make for an interesting Hard fight. It's fairly difficult to make an engaging Medium duo fight.

\subparagraph*{The True Solo} Here you have something like a lich or an older, legendary dragon. These can be totally cinematic fights...if done well. Or they can totally fall flat. You'll need a monster that's much more durable than you expect (since the party can focus fire) and with off-turn actions (or the ability to take multiple turns). You'll especially need a way (such as legendary resistances) to overcome hard control abilities (such as a monk's Stunning Strike); if it can't act at all, the party can just gang up and beat it down without ceremony. Especially in the early game (Journeyman or Adventurer tiers), use these rarely if at all. The party doesn't have a good way of bouncing back after one bad turn (or a really good turn from the enemy's perspective). These fights should be aimed to last for just long enough for the monster to get to do its cool flashy things but not long enough to feel like a damage sponge.

\textbf{Encounter Difficulty with True Solo Fights:} The monster's offensive CR should be at least the party's average level + 2-4 (+5 for deadly), and the defensive CR should similarly be significantly above the party's average level. The monster should have legendary actions and resistances. Honestly, a Medium or Easy True Solo fight is a waste of everyone's time.

\section{Modding the Rules}\label{sec:modding-the-rules}
asdfasdf

\section{Treasure}\label{sec:treasure}
A large part of many adventures is the treasure gained. At the same time, it is possible to unbalance the game (or at least make it harder to reach a fun balance) by giving too much loot too soon. This section is dedicated to explaining the expected treasure distribution and laying out some tools to decide on a distribution that fits your needs.

As a starting point, there is no system requirement that characters ever upgrade their weapons and armor. The game math is built around the idea that you can keep your starting weapon and not run into things that significantly penalize you for doing so. Your starting AC is, for the most part (especially for those most dependent on armor), "ok" as far as the system math is concerned. Weapons and armor (and other items) can give bonuses and benefits, but aren't required. This does \textit{not} mean that giving people magic items is bad or unwelcome. The system math should also tolerate giving items fitting the adventurers' tier (plus or minus one, really) just fine. Giving a legendary item during the journeyman tier can cause problems until they "grow into" the item, but overall the math should be fine.

\subsection{Expectations and Tiers}
Each of the four tiers of play matches the tiers of magic items other than Artifact items (which are always plot-coupons rather than "normal" items). The expectation is that the majority of items acquired during each tier are of that tier of rarity, with the remainder being either one tier higher or one tier lower. Exceptions may exist when the party defeats a substantially higher foe (such as defeating a Legendary tier foe as an Adventuring tier party). These should be kept fairly low.

\subsection{Total Items Awarded}
There are two different large-scale categories of items: major and minor items. Major items include all combat-relevant items, permanent items that grant significant new capabilities, and generally anything that requires attunement. Minor items are mostly consumables, but also some items that are purely role-playing or have extremely minor effects. Major items materially affect the gameplay and accumulate over time, while minor items either don't materially affect the gameplay loop or only do so for a limited number of times. This means you can be much more comfortable giving minor items freely without disruption, while major items should carry narrative weight.

\subsubsection{Minor Items}
Generally consumable items, especially potions of healing, are fine to give out as much as you want. They're limited more by their situational nature and the action economy needed to use them than by their power. If you average 4-6 minor items per character per tier (1-2 per character per session using normal leveling speed), that's a nice amount. But don't feel bad about giving them extra, especially if the party is lacking on spell-based healing. Many of these can be made available to purchase, especially at the Journeyman tier.

\subsubsection{Major Items}
An average party of four characters will generally get 2-3 major items per tier, with a few more during the Adventurer tier. If you're hand-picking the items, they should be items that the party can use and shouldn't be too slanted to any individual character. If you're choosing randomly (see below), all bets are off.

\subsection{Coin}
asdfasdf

\subsection{Other Rewards}
Boons, renown, lands/titles, access, training, skill tricks, spells

\section{Worldbuilding}\label{sec:worldbuilding}
adsfadsf

\section{Sample Adventure}\label{sec:sample-adventure}
asdfasdf

\end{document}